---
title: "Analyzing Data From Atmar 2008"
author: "Kayla Zhang"
date: "2024-02-18"
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We use the R package metaDigitise to extract data points from figure 1C-E in
Atmar *et al*., 2008.
```{r}
require(metaDigitise)
 
atmardata <- metaDigitise("Process Images", summary=FALSE)
atmardata <- rbind(atmardata$scatterplot$atmarplotc.jpg, 
                   rbind(atmardata$scatterplot$atmarplotd.jpg,
                         atmardata$scatterplot$atmarplote.jpg))
```

We then fit a statistical model (Li & Handel, 2014, Holder & Beauchemin, 2011) 
that estimates viral load values over time to the shedding data we just 
extracted using the least squares method. There is a typo in 
Li & Handel, 2014--in the denominator, they have mistakenly written the term 
exp(p4 * (t-p3)) as exp(-p4 * (t-p3)). We make the assumption that viral load is 
a 1-to-1 mapping to viral shedding. We also subtract 19 hours each time we 
convert from days to hours to set time point = 0 as time of inoculation,
since Atmar et al., 2008 defined study day 1 as beginning around 5-6 hours after 
inoculation.
```{r}
generate.estimates <- function(t, pars) {
  
  p1 <- exp(pars[1])
  p2 <- exp(pars[2])
  p3 <- exp(pars[3])
  p4 <- exp(pars[4])
  
  estimate <- (2 * p1) / (exp(-p2*(t-p3)) + exp(p4*(t-p3)))
  
  return(log10(estimate))
  
} 

sse.viralLoad <- function(pars, data) {
  vl <- log10(data$y)
  t <- data$x * 24 - 19
  
  estimates <- generate.estimates(t, pars)
  
  sse <- sum((estimates - vl)^2)
  
  if (sse > 1e20) {
    return(1e20)
  } else {
    return(sse)
  }
  
}

generate.possible.params <- function(data) {
  numguess <- 1e4
  
  fits <- data.frame(matrix(NA, nrow = numguess, ncol = 6))
  
  colnames(fits) <- c('p1','p2','p3','p4','convergence','sse')
  
  # These are ln(param) values
  for (g in 1:numguess) {
    p1 <- runif(1, -1, log(1e12))
    p2 <- runif(1, -1, 1)
    p3 <- runif(1, -1, log(4*24))
    p4 <- runif(1, -5, 1)
    
    params <- c(p1, p2, p3, p4)
    
    fit <- optim(params, sse.viralLoad, data=data)
    
    fits[g, ] <- c(fit$par, fit$convergence, fit$value)
  }
  
  return(fits)
}

get.params <- function(data) {
  
  fits <- generate.possible.params(data)
  
  minsse <- min(fits[fits$convergence==0, ]$sse)
  
  pars <- unlist((fits[fits$sse == minsse, ])[1:4])
  
  return (c(pars, minsse))
}

```

We define two endpoints for the fit: 1) stopping at the first local minimum
in viral load after the peak, and 2) stopping at the first viral load value 
after the peak that is equal to or less than the inoculum dose/initial viral 
load.
```{r}
make.param.df <- function(data, maxt) {
  
  participants <- unique(data$participant)
  
  participantparams <- matrix(nrow=length(participants), ncol=5)
  
  for (i in 1:length(participants)) {
    pars <- get.params(data[data$participant==participants[i],]
                         [1:(maxt[maxt$participant==participants[i],]$max.point),])
    
    participantparams[i,] <- c(participants[i], pars[1:4])
  }
  
  participantparams <- as.data.frame(participantparams)
  colnames(participantparams) <- c("participant", "p1", "p2",
                               "p3", "p4")
  
  return(participantparams)
}

participants <- unique(atmardata$participant)

# Method 1: Stopping at the first local minimum
maxt.firstmin <- data.frame(participant=participants, 
                            max.point=rep(NA, length(participants)))

for (p in participants) {
  
  vldata <- atmardata[atmardata$participant==p,]
  
  t.peak <- which(vldata$y == max(vldata$y))
  
  points.after.peak <- vldata$y[-1:-t.peak]
  
  next.points <- append(vldata$y[-1:-(t.peak+1)], NA)
  
  local.min <- which(points.after.peak <= next.points)[1] + t.peak
  
  if (is.na(local.min)) {
    
    maxt.firstmin[maxt.firstmin$participant==p,]$max.point <- length(vldata$y)
    
  } else {
    
    maxt.firstmin[maxt.firstmin$participant==p,]$max.point <- local.min
    
  }
}

# Method 2: Stopping when viral load gets back down to the initial viral load
maxt.backtoinitial <- data.frame(participant=participants, 
                                 max.point=rep(NA, length(participants)))

for (p in participants) {
  vldata <- atmardata[atmardata$participant==p,]
  
  t.peak <- which(vldata$y == max(vldata$y))
  
  points.after.peak <- vldata$y[-1:-t.peak]
  
  maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point <-  
    which(points.after.peak <= vldata$y[1])[1] + t.peak
}

params.localmin <- make.param.df(atmardata, maxt.firstmin)

params.backtoinitial <- make.param.df(atmardata, maxt.backtoinitial)
```

Now we plot the trajectories of three participants (703, 722, 724) as an example.
```{r}
plot.viralload <- function(data, pars, sympdata, max.modeltime, yhigh, cex,
                           xaxisLabs = T, yaxisLabs = T, 
                           cex_labels, xlabel, ylabel, example) {
  
    
  if(unique(participantdata$dose) == '4800') {
    col <- "#E69F00"
    if (example) {title <- "A) High dose"}
  } else if (unique(participantdata$dose) == '48') {
    col <- "#871a6e" 
    if (example) {title <- "B) Medium dose"}
  } else if (unique(participantdata$dose) == '4.8') {
    col <- "#009E73"
    if (example) {title <- "C) Low dose"}
  }
  
  t <- data$x * 24 - 19
    
  vl <- data$y

  # Plotting real VL
  plot(t, log10(vl), col=col, xlim=c(0, 350), ylim=c(7, yhigh),
     xlab="", ylab="", xaxt="n", yaxt="n", cex.lab=cex, cex.axis=cex, 
     pch=5, font.lab=2, cex=cex, bty='n')
  
  # Plotting simulated VL 
  linet <- seq(min(data$x)*24-19, data[max.modeltime,]$x*24-19, by=1)
  predictedvl <- generate.estimates(linet, pars)
  
  symp.onset <- sympdata$symp.onset * 24 - 19
  symp.end <- symp.onset + sympdata$symp.dur * 24

  lines(linet, predictedvl, col=col)
  
  abline(v = symp.onset, col=col, lty=2)
  abline(v = symp.end, col=col, lty=2)
  
  t.peak <- linet[which(predictedvl == max(predictedvl))]
  
  if (t.peak < symp.onset) {
    redpointstyle <- 19
    symp.timing <- "Pre-symptomatic"
  } else {
    redpointstyle <- 1
    symp.timing <- "Post-symptomatic"
  }
  
  if (!example) {
    title <- paste("ID:", unique(data$participant))
  }
  
  points(t.peak, max(predictedvl), pch=redpointstyle, 
         col="red", lwd=cex, cex=cex+0.5)
  
  xLabs = NA
  if(xaxisLabs==T){
        xLabs = c(expression(0),
                  expression(100),
                  expression(200),
                  expression(300),
                  expression(400))}
  axis(1, cex.axis=cex+0.25, at=seq(0, 400, by=100), labels = xLabs,
       line=1.25,
       mgp=c(2, 0.75, 0), tck=-0.04)
  
  yLabs = NA
  if(yaxisLabs==T){
      yLabs = c(expression(10^7),
                expression(10^8),
                expression(10^9),
                expression(10^10),
                expression(10^11),
                expression(10^12),
                expression(10^13),
                expression(10^14))}
  axis(2, cex.axis=cex, at=seq(7, 14, by=1), 
       labels=yLabs,
       line=1.25,
       las=2, 
       mgp=c(2, 1, 0), tck= -0.04)

  text(x=300, y=13, labels=title, font=2)

  mtext(text=ylabel, side=2, line=4.5, cex=cex_labels)
  
  mtext(text=xlabel, side=1, line=3.5, cex=cex_labels)
  
}

atmarsymptoms <- read.csv('atmarsymptoms.csv')

participantexamples <- c('703', '722', '724')

tiff(filename="Example Atmar.tiff", height=7.5, width=3.5, units = "in",
    res=500)
par(mfrow=c(3, 1), mai=c(0.52, 0.82, 0.42, 0.1), oma=c(2, 3, 0, 0))

for(p in participantexamples) {

  participantdata <- atmardata[atmardata$participant==p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  params <- unlist(params.localmin[params.localmin$participant==p,][2:5])
  
  if(p==participantexamples[2]) {
    ylabel <- "Viral shedding"
    xlabel <- ""
  } else if (p == participantexamples[3]) {
    xlabel <- "Hours after inoculation"
    ylabel <- ""
  } else {
    xlabel <- ""
    ylabel <- ""
  }
  
  plot.viralload(data=participantdata, 
                 pars=params, 
                 sympdata=sympdata, 
                 max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
                 yhigh=13,
                 cex=1.33, 
                 cex_labels=1,
                 xlabel=xlabel,
                 ylabel=ylabel,
                 example=TRUE)
}
dev.off()
```

We then plot the trajectories of each participant using both sets of endpoints.
Since they produce qualitatively similar fits, we arbitrarily use the first 
local minimum in pathogen load as the endpoint going forward.
```{r}
participants.by.dose <- unique(atmardata$participant[order(atmardata$dose)])

tiff(filename="Participant Trajectories.tiff", height=7.5, width=6.5,
    units="in", res=500)
par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(participants.by.dose)) {
  
  p <- participants.by.dose[i]
  participantdata <- atmardata[atmardata$participant==p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant==p, ]
  params <- unlist(params.localmin[params.localmin$participant==p,][2:5])
  dose <- unique(participantdata$dose)
  
  if (i %in% c(1, 4, 7, 10, 13)) {
    plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      xaxisLabs=F,
      cex_labels=0.8,
      xlabel="",
      ylabel="Viral shedding",
      example=FALSE)
  } else if (i %in% c(14, 15)) {
     plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      yaxisLabs=F,
      cex_labels=0.8,
      xlabel="Hours after inoculation",
      ylabel="",
      example=FALSE) 
  } else if (i==16) {
    plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      cex_labels=0.8,
      xlabel="Hours after inoculation",
      ylabel="Viral shedding",
      example=FALSE)
  } else {
    plot.viralload(
    data=participantdata, 
    pars=params, 
    sympdata=sympdata, 
    max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
    yhigh=14,
    cex=1,
    xaxisLabs=F,
    yaxisLabs=F,
    cex_labels=0.8,
    xlabel="",
    ylabel="",
    example=FALSE)
  }
}

par(mar=c(0, 0, 0, 0))
plot.new()
legend(x = "bottom",
       legend = c("4800 RT-PCR Units",
                  "48 RT-PCR Units",
                  "4.8 RT-PCR Units"),
       col = c("#E69F00",
               "#871a6e",
               "#009E73"),
       lty = 1, bty='n')
plot.new()
dev.off()

tiff(filename="Participant Trajectories ver 2.tiff", height=7.5, width=6.5,
    units="in", res=500)
par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(participants.by.dose)) {
  
  p <- participants.by.dose[i]
  participantdata <- atmardata[atmardata$participant==p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant==p, ]
  params <- unlist(params.backtoinitial[params.backtoinitial$participant==p,][2:5])
  dose <- unique(participantdata$dose)
  
  if (i %in% c(1, 4, 7, 10, 13)) {
    plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      xaxisLabs=F,
      cex_labels=0.8,
      xlabel="",
      ylabel="Viral shedding",
      example=FALSE)
  } else if (i %in% c(14, 15)) {
     plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      yaxisLabs=F,
      cex_labels=0.8,
      xlabel="Hours after inoculation",
      ylabel="",
      example=FALSE) 
  } else if (i==16) {
    plot.viralload(
      data=participantdata, 
      pars=params, 
      sympdata=sympdata, 
      max.modeltime=maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point,
      yhigh=14,
      cex=1,
      cex_labels=0.8,
      xlabel="Hours after inoculation",
      ylabel="Viral shedding",
      example=FALSE)
  } else {
    plot.viralload(
    data=participantdata, 
    pars=params, 
    sympdata=sympdata, 
    max.modeltime=maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point,
    yhigh=14,
    cex=1,
    xaxisLabs=F,
    yaxisLabs=F,
    cex_labels=0.8,
    xlabel="",
    ylabel="",
    example=FALSE)
  }
}

par(mar=c(0, 0, 0, 0))
plot.new()
legend(x = "bottom",
       legend = c("4800 RT-PCR Units",
                  "48 RT-PCR Units",
                  "4.8 RT-PCR Units"),
       col = c("#E69F00",
               "#871a6e",
               "#009E73"),
       lty = 1, bty='n')
plot.new()
dev.off()
```
Because participants 723 and 731 do not have clear peaks in pathogen load,
we exclude them from any future calculations using our conservative measure
of pre-symptomatic transmission, which is that peak pathogen load is reached
before symptom onset.

```{r}
excluded.participants <- c('723', '731')

tiff(filename="Atmar Disqualified.tiff",
           height=6.5, width=3.5, units="in", res=500)
par(mfrow=c(2, 1), mai=c(0.75, 0.75, 0.82, 0.25), oma=c(1, 1, 0, 0))

for(i in 1:length(excluded.participants)) {

  p <- excluded.participants[i]
  participantdata <- atmardata[atmardata$participant==p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  params <- unlist(params.localmin[params.localmin$participant==p,][2:5])
  
  if(i == 2) {
    ylabel <- "Viral shedding"
    xlabel <- "Hours after inoculation"
  } else {
    xlabel <- ""
    ylabel <- ""
  }
  
  plot.viralload(data=participantdata, 
                 pars=params, 
                 sympdata=sympdata, 
                 max.modeltime=maxt.firstmin[maxt.firstmin$participant==p,]$max.point,
                 yhigh=13,
                 cex=0.9, 
                 cex_labels=0.9,
                 xlabel=xlabel,
                 ylabel=ylabel,
                 example=FALSE,
                 letter = toupper(letters)[i])
  if(i==2) {
    legend(x = "topright",
       legend = c("48 RT-PCR Units",
                  "4.8 RT-PCR Units"),
       col = c("#871a6e",
               "#009E73"),
       lty = 1, bty='n', cex=0.8)
  } 
}

dev.off()

```


Here we calculate the delay between symptom onset and peak pathogen load for
each participant in the Atmar *et al*., 2008 study.
```{r}
participants.conservative <- participants[participants != 723 & 
                                            participants != 731]

delays.conservative <- data.frame("participant"=participants.conservative, 
                    "delay"=rep(0, length(participants.conservative)), 
                    "dose"=rep(0, length(participants.conservative)))

for (p in participants.conservative) {
  
  vldata <- atmardata[atmardata$participant == p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  dose <- unique(vldata$dose)
  
  params <- unlist(params.localmin[params.localmin$participant==p, 2:5])

  t <- seq(min(vldata$x)*24-19, 
        vldata[maxt.firstmin[maxt.firstmin$participant==p,]$max.point,]$x*24-19, 
        by=1)
  
  predictedvl <- generate.estimates(t, params)
  
  peak.vl.time <- t[which(predictedvl==max(predictedvl))]
  
  participantdelay <- peak.vl.time - (sympdata$symp.onset * 24 - 19)
  
  delays.conservative[delays.conservative$participant==p, ]$delay <- participantdelay
   
  delays.conservative[delays.conservative$participant==p, ]$dose <- dose

}
```

We run a linear regression between the delay between symptom onset and peak
pathogen load and each of the 4 parameters from the statistical model as well
as between viral growth and decay rate. Only p2 and p4, which represent pathogen 
replication rate and time of peak pathogen load, respectively, were found to 
have significant correlations with delay. They are also significantly correlated 
 to each other.
```{r}
params.and.delays.conservative <- merge(params.localmin, delays.conservative)

lm.fit.p1.delay.conservative <- 
  lm(delay/24 ~ p1, data=params.and.delays.conservative)
p1.delay.conservative.pval <- 
  summary(lm.fit.p1.delay.conservative)$coefficients[2,4]
p1.delay.conservative.pval

lm.fit.p2.delay.conservative <- 
  lm(delay/24 ~ p2, data=params.and.delays.conservative)
p2.delay.conservative.pval <- 
  summary(lm.fit.p2.delay.conservative)$coefficients[2,4]
p2.delay.conservative.pval

lm.fit.p3.delay.conservative <- 
  lm(delay/24 ~ p3, data=params.and.delays.conservative)
p3.delay.conservative.pval <- 
  summary(lm.fit.p3.delay.conservative)$coefficients[2,4]
p3.delay.conservative.pval

lm.fit.p4.delay.conservative <- 
  lm(delay/24 ~ p4, data=params.and.delays.conservative)
p4.delay.conservative.pval <- 
  summary(lm.fit.p4.delay.conservative)$coefficients[2,4]
p4.delay.conservative.pval

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative <- 
  lm(p4 ~ p2, data=params.and.delays.conservative)
p4.p2.conservative.pval <- 
  summary(lm.fit.p4.p2.conservative)$coefficients[2,4]
p4.p2.conservative.pval

# Test for collinearity
lm.fit.p2.p3.conservative <-
  lm(p2 ~ p3, data=params.and.delays.conservative)
p2.p3.conservative.pval <-
  summary(lm.fit.p2.p3.conservative)$coefficients[2,4]
p2.p3.conservative.pval
```

Using a non-conservative measure of pre-symptomatic transmission, ie shedding
onset values as reported in the study, we calculate the delay between
shedding onset and symptom onset (the length of the latency period). For 
simplicity's sake, a negative delay still indicates pre-symptomatic transmission.

```{r}
atmarshedding <- read.csv("atmarshedding.csv")

delays.nonconservative <- data.frame("participant"=participants, 
                    "delay"=rep(0, length(participants)), 
                    "dose"=rep(0, length(participants)))

for (p in participants) {
  
  sheddata <- atmarshedding[atmarshedding$participant == p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  dose <- unique(sheddata$dose)
  
  participantdelay <- sheddata$shed.onset - sympdata$symp.onset
  
  delays.nonconservative[delays.nonconservative$participant==p, ]$delay <- participantdelay
   
  delays.nonconservative[delays.nonconservative$participant==p, ]$dose <- dose

}
```

As before, we test for any significant correlations between the statistical 
model parameters and the duration of the latency period as well as
between viral growth and decay rate. Just as with the
conservative measure of pre-symptomatic transmission, there is a significant
correlation between both viral replication rate and time of peak pathogen load
with the duration of the latency period. We test for collinearity with
another linear regression, which returns positive.
```{r}
params.and.delays.nonconservative <- merge(params.localmin, delays.nonconservative)

lm.fit.p1.delay.nonconservative <- 
  lm(delay ~ p1, data=params.and.delays.nonconservative)
p1.delay.nonconservative.pval <- 
  summary(lm.fit.p1.delay.nonconservative)$coefficients[2,4]
p1.delay.nonconservative.pval

lm.fit.p2.delay.nonconservative <- 
  lm(delay ~ p2, data=params.and.delays.nonconservative)
p2.delay.nonconservative.pval <- 
  summary(lm.fit.p2.delay.nonconservative)$coefficients[2,4]
p2.delay.nonconservative.pval

lm.fit.p3.delay.nonconservative <- 
  lm(delay ~ p3, data=params.and.delays.nonconservative)
p3.delay.nonconservative.pval <- 
  summary(lm.fit.p3.delay.nonconservative)$coefficients[2,4]
p3.delay.nonconservative.pval

lm.fit.p4.delay.nonconservative <- 
  lm(delay ~ p4, data=params.and.delays.nonconservative)
p4.delay.nonconservative.pval <- 
  summary(lm.fit.p4.delay.nonconservative)$coefficients[2,4]
p4.delay.nonconservative.pval

# Viral growth vs. decay rate
lm.fit.p4.p2.nonconservative <- 
  lm(p4 ~ p2, data=params.and.delays.nonconservative)
p4.p2.nonconservative.pval <- 
  summary(lm.fit.p4.p2.nonconservative)$coefficients[2,4]
p4.p2.nonconservative.pval

# Test for collinearity
lm.fit.p2.p3.nonconservative <-
  lm(p2 ~ p3, data=params.and.delays.nonconservative)
p2.p3.nonconservative.pval <-
  summary(lm.fit.p2.p3.nonconservative)$coefficients[2,4]
p2.p3.nonconservative.pval
```

We also run a linear regression between pre-symptomatic transmission potential,
measured as average shedding prior to symptoms, and time of symptom onset. 
There is no correlation.
```{r}
transmission.latent <- data.frame("participant"=participants.conservative, 
                                  "dose"=rep(0, length(participants.conservative)),
                    "transmission_potential"=rep(0, length(participants.conservative)), 
                    "symptom_onset"=rep(0, length(participants.conservative)))

for (i in 1:length(participants.conservative)) {
  p <- participants.conservative[i]
  
  transmission.latent[i,]$participant <- p
  
  symp.onset <- atmarsymptoms[atmarsymptoms$participant == p,]$symp.onset
  
  transmission.latent[i,]$symptom_onset <- symp.onset
  
  params <- unlist(params.localmin[params.localmin$participant == p,][2:5])
  
  vldata <- atmardata[atmardata$participant == p,]
  
  transmission.latent[i,]$dose <- unique(vldata$dose)
  
  shed.onset <- min(vldata$x) * 24 - 19
  
  if(shed.onset > symp.onset * 24 - 19) {
    avg.shedding.latent <- 0
  } else {
    t <- seq(min(vldata$x)*24-19, symp.onset*24-19, by=1)
    vl.latent <- 10^generate.estimates(t, params)
    avg.shedding.latent <- log10(mean(vl.latent))
  }
  
  transmission.latent[i,]$transmission_potential <- avg.shedding.latent
}

transmission.latent.fit <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent)
transmission.latent.pval <- 
  summary(transmission.latent.fit)$coefficients[2,4]

transmission.latent.pval
```

We plot delay vs. inoculum dose, delay vs. viral growth rate, and pre-symptomatic
transmission potential vs. symptom onset in a 3-panel figure.
```{r}

id <- delays.conservative$dose
id[id==4.8] <- 1
id[id==48] <- 2
id[id==4800] <- 3

participant.params <- params.localmin[params.localmin$participant != 723 & 
                                        params.localmin$participant != 731,]
transmission.latent.delays.data <- merge(transmission.latent, delays.conservative)

colors <- c("#009E73", "#871a6e","#E69F00")
cex <- 1.33
cex_labels <- .95
panellabelratio <- .975

{
  tiff(filename="Atmar Presymptomatic Transmission.tiff", 
             height=7.5, width=3.5, units="in", res=500)
  par(mfrow=c(3, 1), oma=c(0.5, 4, 0, 0), mai=c(0.6, 0.6, 0.2, 0.1))
  
  # Panel A: Delay between symptom onset and peak viral shedding vs. inoculum
  # dose
  
  plot(jitter(id), delays.conservative$delay/24,
          col=colors[factor(id)], xlim=c(.5, 3.5),
          ylim=c(-0.5, 6), cex.axis=cex, cex=cex,
          pch=ifelse(delays.conservative$delay < 0, 16, 1),
          xaxt="n",
          ylab = "",
          xlab = "",
       mgp=c(0, 1.5, 0), tck=-0.04, las=2, bty='l')
  axis(1, at=seq(1,3), labels=c("4.8", "48", "4800"), cex.axis=cex, 
       mgp=c(0, 1.5, 0), tck=-0.04)
  mtext(text="Day of peak viral shedding \n minus day of symptom onset",
        side=2, line=-1, cex=cex_labels, outer=TRUE, adj=0.75)
  mtext(text="Inoculum dose", side=1, line=3.5, cex=cex_labels)
  
  #Average delays per infectious dose
  segments(0.7, 
           mean(delays.conservative[delays.conservative$dose == 4.8,]$delay/24), 
           1.3, col=colors[1], lwd=cex)
  segments(1.7,
           mean(delays.conservative[delays.conservative$dose == 48,]$delay/24), 
           2.3, col=colors[2], lwd=cex)
  segments(2.7,
           mean(delays.conservative[delays.conservative$dose == 4800,]$delay/24), 
           3.3, col=colors[3], lwd=cex)
  
  #Pre-symptomatic vs post-symptomatic line
  abline(h=0, lty=3, lwd=cex)
  
  text(panellabelratio * 3 + .5, panellabelratio * 7 - 1, labels="A", cex=cex, font=2)
  
   legend(0.5, 6, legend=c("Pre-symptomatic", "Post-symptomatic"), 
         pch=c(19, 1, NA), col=c("black", "black"), cex=cex_labels, bty="n")
 
  # Panel B: Delay between symptom onset and peak viral shedding vs. viral
  # growth rate
  
  plot(participant.params[order(participant.params$participant),]$p2,
     delays.conservative[order(delays.conservative$participant),]$delay/24, 
     col=colors[factor(delays.conservative[order(delays.conservative$participant),]$dose)],
     pch=ifelse(delays.conservative[order(delays.conservative$participant),]$delay < 0, 16, 1),
     cex=cex, cex.axis=cex,
     ylab="", xlab="", xaxt="n",
     ylim=c(-0.5, 6),
     mgp=c(0, 1.5, 0), tck=-0.04, las=2, bty='l')
  
  axis(1, at=seq(-4, 1), labels=c(expression(10^-4), 
                                             expression(10^-3), 
                                             expression(10^-2),
                                             expression(10^-1),
                                             expression(10^0),
                                             expression(10^1)), cex.axis=cex, 
       mgp=c(0, 1.5, 0), tck=-0.04)
  
  abline(lm.fit.p2.delay.conservative, col="red")
  
  abline(h=0, lty=3, lwd=cex)
  
  mtext(text=expression(Viral~replication~rate~(hour^-1)),
        side=1, line=4, cex=cex_labels)
  
  text(panellabelratio * 5 - 4, panellabelratio * 7 - 1, labels="B", cex=cex, font=2)
  
  # Panel C: Pre-symptomatic transmission potential vs. symptom onset
  
  plot(transmission_potential ~ jitter(symptom_onset), data=transmission.latent.delays.data,
          col=colors[factor(dose)], xlim=c(0, 4),
        pch=ifelse(delay < 0, 16, 1),
          ylim=c(-1, 15), cex.axis=cex, cex=cex,
          ylab = "",
          xlab = "",
       xaxt="n",
       yaxt="n", bty='l')
  
  axis(1, cex.axis=cex, mgp=c(0, 1.5, 0), tck=-0.04)
  axis(2, cex.axis=cex, mgp=c(0, 1.5, 0), tck=-0.04, at=seq(0, 15, by=5),
       labels=c(expression(10^0),
                expression(10^5),
                expression(10^10),
                expression(10^15)),
       las=2)

  mtext(text="Average viral shedding \n prior to symptom onset",
        side=2, line=5, cex=cex_labels)
  mtext(text="Symptom onset (day)", side=1, line=3, cex=cex_labels)

  text(panellabelratio * 4, panellabelratio * 16 - 1, labels="C", cex=cex, font=2)
  
  dev.off()
}
```

We again plot delay vs. inoculum dose and delay vs. viral growth rate using
the non-conservative measure of pre-symptomatic transmission.

```{r}

id.nonconservative <- delays.nonconservative$dose
id.nonconservative[id.nonconservative==4.8] <- 1
id.nonconservative[id.nonconservative==48] <- 2
id.nonconservative[id.nonconservative==4800] <- 3

cex <- 0.9
cex_labels <- 0.9
panellabelratio <- .975

{
  tiff(filename="Atmar Presymptomatic Transmission (non-conservative).tiff", 
             height=6, width=3.5, units="in", res=500)
  par(mfrow=c(2, 1), oma=c(1.75, 1, 0, 0), mai=c(0.5, 0.82, 0.5, 0.1))
  
  # Panel A: Delay between symptom onset and peak viral shedding vs. inoculum
  # dose
  
    plot(jitter(id.nonconservative), delays.nonconservative$delay,
          col=colors[factor(id.nonconservative)], xlim=c(.5, 3.5),
          ylim=c(-2, 4), cex.axis=cex, cex=cex,
          pch=ifelse(delays.nonconservative$delay < 0, 16, 1),
          xaxt="n",
          ylab = "",
          xlab = "",
       mgp=c(0, 0.8, 0), tck=-0.04, las=2, bty='l')

  axis(1, at=seq(1,3), labels=c("4.8", "48", "4800"), cex.axis=cex, 
       mgp=c(0, 0.8, 0), tck=-0.04)
  mtext(text="Day of peak viral shedding \n minus day of symptom onset",
        side=2, line=-1.5, cex=cex_labels, outer=TRUE)
  mtext(text="Inoculum dose", side=1, line=2, cex=cex_labels)
  
    #Average delays per inoculum dose
  segments(0.7, 
           mean(delays.nonconservative[delays.nonconservative$dose == 4.8,]$delay), 
           1.3, col=colors[1], lwd=cex)
  segments(1.7,
           mean(delays.nonconservative[delays.nonconservative$dose == 48,]$delay), 
           2.3, col=colors[2], lwd=cex)
  segments(2.7,
           mean(delays.nonconservative[delays.nonconservative$dose == 4800,]$delay), 
           3.3, col=colors[3], lwd=cex)
  
  #Pre-symptomatic vs post-symptomatic line
  abline(h=0, lty=3, lwd=cex)
  
  text(panellabelratio * 3 + .5, panellabelratio * 6 - 2, labels="A", cex=cex, font=2)
 
  # Panel B: Delay between symptom onset and peak viral shedding vs. viral
  # growth rate
  
  plot(params.localmin[order(params.localmin$participant),]$p2,
     delays.nonconservative[order(delays.nonconservative$participant),]$delay, 
     col=colors[factor(delays.nonconservative[order(delays.nonconservative$participant),]$dose)],
     pch=ifelse(delays.nonconservative[order(delays.nonconservative$participant),]$delay < 0, 16, 1),
     cex=cex, cex.axis=cex,
     ylab="", xlab="", xaxt="n",
     ylim=c(-2, 4),
     mgp=c(0, 0.8, 0), tck=-0.04, las=2, bty='l')
  
  axis(1, at=seq(-4, 1), labels=c(expression(10^-4), 
                                             expression(10^-3), 
                                             expression(10^-2),
                                             expression(10^-1),
                                             expression(10^0),
                                             expression(10^1)), cex.axis=cex, 
       mgp=c(0, 0.8, 0), tck=-0.04)
  
  abline(lm.fit.p2.delay.nonconservative, col="red")
  
  abline(h=0, lty=3, lwd=cex)
  
  mtext(text=expression(Viral~replication~rate~(hour^-1)),
        side=1, line=2.5, cex=cex_labels)

    text(panellabelratio * 5 - 4, panellabelratio * 6 - 2, labels="B", cex=cex, font=2)
  
    legend(-3.5, 2.9, legend=c("Pre-symptomatic", "Post-symptomatic", 
                          "Least squares regression line"), pch=c(19, 1, NA),
                          lty=c(NA, NA, 1), col=c("black", "black", "red"),
       cex=0.6, bty="n")
  
  dev.off()
}
```

Lastly, we break down the correlation between pre-symptomatic transmission
and viral growth rate by examining the relationships between time of
peak shedding and viral growth rate as well as between symptom onset
and viral growth rate.

```{r}

participant.peaktimes <- data.frame("participant"=participants.conservative, 
                    "peaktime"=rep(0, length(participants.conservative)), 
                    "dose"=rep(0, length(participants.conservative)))

for (p in participants.conservative) {
  
  vldata <- atmardata[atmardata$participant == p, ]
  
  t <- seq(min(vldata$x)*24-19, 
        vldata[maxt.firstmin[maxt.firstmin$participant==p,]$max.point,]$x*24-19, 
        by=1)
  
  params <- unlist(params.localmin[params.localmin$participant==p, 2:5])
  
  predictedvl <- generate.estimates(t, params)
  
  participant.peaktime <- t[which(predictedvl==max(predictedvl))]
  
  dose <- unique(vldata$dose)
  
  participant.peaktimes[participant.peaktimes$participant==p, ]$dose <- dose
  
  participant.peaktimes[participant.peaktimes$participant==p, ]$peaktime <- 
    participant.peaktime
}

params.peaktime.data <- merge(merge(participant.params, participant.peaktimes),
                              delays.conservative)
lm.fit.p2.peaktime <- lm(peaktime/24 ~ p2, data=params.peaktime.data)
p2.peaktime.pval <- summary(lm.fit.p2.peaktime)$coefficients[2,4]
p2.peaktime.pval

params.symp.data <- merge(merge(participant.params, atmarsymptoms),
                          delays.conservative)
lm.fit.p2.symponset <- lm(symp.onset ~ p2, data=params.symp.data)
p2.symponset.pval <- summary(lm.fit.p2.symponset)$coefficients[2,4]
p2.symponset.pval

cex <- 0.9
cex_labels <- 0.9

{
  tiff(filename="Broken up p2 impact.tiff", 
             height=3, width=6.5, units="in", res=500)
  par(mfrow=c(1, 2), mai=c(0.5, 0.65, 0.1, 0.3), oma=c(1.75, 1, 0, 0))
  
  plot(peaktime/24 ~ p2, data=params.peaktime.data,
            col=colors[factor(dose)], xlim=c(-4, 1),
            ylim=c(1, 6), cex.axis=cex, cex=cex,
            pch=ifelse(delay < 0, 16, 1),
            xaxt="n",
            ylab = "",
            xlab = "",
         mgp=c(0, 0.8, 0), tck=-0.04, las=2, bty='l')
  axis(1, at=seq(-4, 1), labels=c(expression(10^-4), 
                                  expression(10^-3),
                                  expression(10^-2),
                                  expression(10^-1),
                                  expression(10^0),
                                  expression(10^1)), cex.axis=cex, 
       mgp=c(0, 0.8, 0), tck=-0.04)
  mtext(text="Day of peak shedding",
        side=2, line=2, cex=cex_labels)
  mtext(text=expression(Viral~replication~rate~(hour^-1)),
        side=1, line=0, cex=cex_labels, outer = TRUE, adj=0.52)
  
  abline(lm.fit.p2.peaktime, col="red")
  
  text(panellabelratio * 5 - 4, panellabelratio * 5 + 1, labels="A", cex=cex, font=2)
    
  plot(symp.onset ~ p2, data=params.symp.data,
    col=colors[factor(dose)], xlim=c(-4, 1),
            ylim=c(1, 4), cex.axis=cex, cex=cex,
            pch=ifelse(delay < 0, 16, 1),
            yaxt="n",
            xaxt="n",
            ylab = "",
            xlab = "", bty='l')
  
  axis(1, at=seq(-4, 1), labels=c(expression(10^-4), 
                                  expression(10^-3),
                                  expression(10^-2),
                                  expression(10^-1),
                                  expression(10^0),
                                  expression(10^1)), cex.axis=cex, 
       mgp=c(0, 0.8, 0), tck=-0.04)
  mtext(text="Day of symptom onset",
        side=2, line=2, cex=cex)
  
  axis(2, at=seq(1, 4), cex.axis=cex, mgp=c(0, 0.8, 0), tck=-0.04, las=2)
  
  abline(lm.fit.p2.symponset, col="red")
  
  text(panellabelratio * 5 - 4, panellabelratio * 3 + 1, labels="B", cex=cex, font=2)
  
  dev.off()
}
```

The correlation between peak shedding and viral growth rate is verified
by running a linear regression between the p2 and p3 parameters from the
statistical model

```{r}

lm.fit.p2.p3 <- 
  lm(p3 ~ p2, data=participant.params)
p2.p3.pval <- 
  summary(lm.fit.p2.p3)$coefficients[2,4]
p2.p3.pval

```