---
title: "Analyzing human infection data (norovirus and SARS-CoV-2)"
author: "Kayla Zhang, Damie Pak, Megan Greischar"
date: "2024-04-27"
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

Packages we use
```{r}
library(here)
library(pzfx)
library(data.table)
library(spatialEco)
library(Polychrome)
library(pracma)
```

We use the shedding and symptom timing data that Ge *et al*., 2023 obtained 
from a norovirus study (Atmar *et al*., 2008). We exclude participants with 
less than 5 quantitative viral shedding points to avoid overfitting the 
statistical model. This data also contains qualitative test results for 
viral shedding values (positive or negative) falling below the qRT-PCR detection 
threshold. Any shedding values below the detection threshold for quantitative
results are set to NA.
```{r, echo = TRUE, eval = FALSE}
exclude.participants <- function(p, data) {
  pdata <- data[data$ID == p, ]
  
  if(nrow(pdata[!is.na(pdata$y_virus),]) < 5) {
    return(p)
  }
}

noroData <- readRDS(here("data", "norodata.rds")) 

write.table(noroData$ftsshedding, "noroFtsData.txt")
noro.FShedData <- read.table("noroFtsData.txt", stringsAsFactors = F)
noro.FShedData <- noro.FShedData[, c(1, 2, 5, 6, 8)]

write.table(noroData$incubation, "noroIncubationData.txt")
noro.IncubData <- read.table("noroIncubationData.txt", stringsAsFactors = F)
noro.IncubData <- noro.IncubData[, c(1, 2, 5)]

# Get rid of viral shedding numbers below the qRT-PCR detection threshold 
# (4e7 copies /ml)
noro.detection.threshold <- 15000
noro.detection.threshold.2 <- 4e7
noro.FShedData[noro.FShedData$y_virus <= noro.detection.threshold.2,]$y_virus <- NA

# Exclude participants with less than 5 points of data 
noro.excluded <- unlist(lapply(unique(noro.FShedData$ID), exclude.participants, 
                               data=noro.FShedData))

noro.excluded <- noro.excluded[!is.null(noro.excluded)]

noro.FShedData <- noro.FShedData[!(noro.FShedData$ID %in% noro.excluded),]
noro.IncubData <- noro.IncubData[!(noro.IncubData$ID %in% noro.excluded),]

# Average across viral shedding values at the same time point
aggregated_virus <- aggregate(y_virus ~ ID + days, 
                              noro.FShedData,
                              mean,
                              na.action=na.pass)
aggregated_virus <- aggregated_virus[order(as.numeric(substring(aggregated_virus$ID, 3, 4)),
                       aggregated_virus$days),]$y_virus

noro.FShedData <- noro.FShedData[!duplicated(noro.FShedData[, c(1, 3)]),]
noro.FShedData$y_virus <- aggregated_virus
```

We extract the shedding and symptom timing data from Zhou *et al*., 2023, 
a SARS-CoV-2 study as well.
```{r}
extract.incubation <- function(symp.data) {
  
  symptom.indexes <- symp.data[which(symp.data[, 2] > 0),]$ROWTITLE
  
  if (length(symptom.indexes) == 0) {
    symp.onset <- NA
  } else {
    symp.onset <- min(symptom.indexes) 
  }
  
  incub.data <- data.frame("ID" = colnames(symp.data)[2], "incubation" = symp.onset)
  
  return(incub.data)
}

sars.TShedData <- read_pzfx(here("data", "SARS-Supp-2.pzfx"), table=37)[, c(1, 20:37)]
sars.TShedData <- melt(as.data.table(sars.TShedData), id.vars=c("Day"), 
                       variable.name = "ID", value.name = "y_virus", 
                       variable.factor = F)
colnames(sars.TShedData)[1] <- "days"

incub.indexes <- 37:54
sars.IncubData <- data.frame(do.call(rbind, 
                                     lapply(incub.indexes, 
                                            function(x) 
                                              extract.incubation(
                                                read_pzfx(here("data", "SARS-Fig-1.pzfx"), 
                                                    table=x)))))

# Participants were matched up by visually comparing viral shedding values

ID_dict_sars <- c("680539" = "Throat_10",
"673353" = "Throat_18",
"667186" = "Throat_11",
"550630" = "Throat_7",
"635495" = "Throat_1",
"666482" = "Throat_9",
"651806" = "Throat_15",
"635331" = "Throat_5",
"634105" = "Throat_2",
"635779" = "Throat_16",
"635729" = "Throat_4",
"637340" = "Throat_12",
"627506" = "Throat_6",
"666660" = "Throat_8",
"676586" = "Throat_13",
"666427" = "Throat_14",
"647785" = "Throat_3",
"674700" = "Throat_17"
)

sars.IncubData$ID <- ID_dict_sars[sars.IncubData$ID]

# Get rid of viral shedding below the detection threshold (1200 copies/ml)
sars.detection.threshold <- 1200
sars.TShedData[sars.TShedData$y_virus < sars.detection.threshold,]$y_virus <- NA

# Exclude participants with less than 5 points of data 
sars.excluded <- unlist(lapply(unique(sars.TShedData$ID), exclude.participants, data=sars.TShedData))

sars.excluded <- sars.excluded[!is.null(sars.excluded)]

sars.TShedData <- sars.TShedData[!(sars.TShedData$ID %in% sars.excluded),]
sars.IncubData <- sars.IncubData[!(sars.IncubData$ID %in% sars.excluded),]
```

We calculate each study's average shedding and symptom onset.
```{r}
shed.onset <- function(vldata) {
  if("y_type" %in% colnames(vldata)) {
    return(min(vldata[grepl('POS', vldata$y_type),]$days))
  } else {
    return(min(vldata[!is.na(vldata$y_virus),]$days))
  }
}

noroParticipants <- unique(noro.FShedData$ID)

noro.avg.shed.onset <- mean(unlist(lapply(noroParticipants, function(x) 
  shed.onset(noro.FShedData[noro.FShedData$ID == x, ]))))
noro.avg.shed.onset

noro.avg.symp.onset <- mean(noro.IncubData$incubation)
noro.avg.symp.onset

sarsParticipants <- unique(sars.TShedData$ID)

sars.avg.shed.onset <- mean(unlist(lapply(sarsParticipants, function(x) 
  shed.onset(sars.TShedData[sars.TShedData$ID == x, ]))))
sars.avg.shed.onset

sars.avg.symp.onset <- mean(sars.IncubData$incubation, na.rm=T)
sars.avg.symp.onset
```

We fit a statistical model (Li & Handel, 2014, Holder & Beauchemin, 2011) to each
individual's shedding data that estimates viral load over time using the 
least squares method. There is a typo in Li & Handel, 2014 in the denominator, 
where the term exp(p4 * (t-p3)) is written as exp(-p4 * (t-p3)) (Holder & 
Beauchemin, 2011). We make the assumption that viral load is a 1-to-1 mapping to 
viral shedding. 

When model estimates incorrectly lie outside a given detection range,
we calculate the SSE of that point using the nearest detection threshold as 
the "real" data point.
```{r}
generate.estimates <- function(t, pars) {
  
  p1 <- exp(pars[1])
  p2 <- exp(pars[2])
  p3 <- exp(pars[3])
  p4 <- exp(pars[4])
  
  estimate <- (2 * p1) / (exp(-p2*(t-p3)) + exp(p4*(t-p3)))
  return(log10(estimate))
  
} 

sse.viralLoad <- function(pars, data, detect.threshold, detect.threshold.2 = NA) {
  vl <- log10(data$y_virus)
  t <- data$days * 24
  
  estimates <- generate.estimates(t, pars)
  
  # Punishing model estimates that incorrectly lie above the detection
  # threshold
  vl[which(is.na(vl) & estimates > detect.threshold)] <- detect.threshold
  
  # Punishing model estimates that incorrectly lie outside the detection range of 
  # qualitative results
  vl[which(grepl('~', data$y_type) & estimates < detect.threshold)] <- detect.threshold 
  vl[which(grepl('~', data$y_type) & estimates > detect.threshold.2)] <- detect.threshold.2
  
  sse <- sum((estimates - vl)^2, na.rm=TRUE)

  if (sse > 1e20) {
    return(1e20)
  } else {
    return(sse)
  }
  
}

generate.possible.params <- function(data, detect.threshold,
                                     detect.threshold.2 = NA) {
  
  # These are ln(param) values
  
  p1 <- runif(1, -1, log(1e12))
  p2 <- runif(1, -1, 1) 
  p3 <- runif(1, -1, log(4 * 24))
  p4 <- runif(1, -5, 1)
  
  params <- c(p1, p2, p3, p4)

  fit <- optim(params, sse.viralLoad, data=data, detect.threshold=detect.threshold,
               detect.threshold.2=detect.threshold.2)
  
  param.guess <- data.frame(unique(data$ID), 
                         fit$par[1],
                         fit$par[2],
                         fit$par[3],
                         fit$par[4],
                         fit$convergence, 
                         fit$value)
  
  return(param.guess)
  
}

generate.fits <- function(data, numguess, detect.threshold, detect.threshold.2 = NA) {

  fits <- do.call(rbind, 
                  lapply(1:numguess,
                         function(x) generate.possible.params(data, detect.threshold,
                                                              detect.threshold.2))) 
  
  colnames(fits) <- c('ID', 'p1','p2','p3','p4','convergence','sse')
  return(fits)
}

params.minsse <- function(fits) {
  
  minsse <- min(fits[fits$convergence==0, ]$sse)
  pars <- unlist((fits[fits$sse == minsse, ])[1:5])
  return(pars)
}
 
```

For endpoints, we looked at values on or after the fifth viral load value, to
ensure we are not overfitting the model to too few data points. We define two 
endpoints: 1) stopping at the first local minimum in viral load (two greater 
points at each side), and 2) stopping at the first viral load value that is 
equal to or less than the inoculum dose/initial viral load. If there are 
no data points that meet these criteria, we set the endpoint to the last 
viral shedding value.
```{r}
first.local.min <- function(data) {
       
  quantifiable.vl <- data[!is.na(data$y_virus),]$y_virus
  quantifiable.vl <- quantifiable.vl[3:length(quantifiable.vl)]
  
  minima <- local.min.max(quantifiable.vl)$minima

  for (mi in minima) {
    
    min.index <- min(which(quantifiable.vl == mi))
    data.context <- quantifiable.vl[(min.index - 2): (min.index + 2)]
    
    if(length(na.omit(data.context)) == 5 & (min(data.context) == mi)) {
      return(min(which(data$y_virus==quantifiable.vl[min.index])))
    }
  }

  return(max(which(data$y_virus == quantifiable.vl[length(quantifiable.vl)])))
  
}

back.to.initial <- function(data) {
  
  quantifiable.vl <- data[!is.na(data$y_virus),]$y_virus 
  less.than.initial <- which(quantifiable.vl[5:length(quantifiable.vl)] <= 
                               quantifiable.vl[1])

  if (length(less.than.initial) == 0) {
    return(which(data$y_virus == quantifiable.vl[length(quantifiable.vl)]))
  } else {
    return(which(data$y_virus == quantifiable.vl[min(less.than.initial)+4]))
  }
}

all.fits <- function(data, maxt, numguess, detect.threshold, detect.threshold.2 = NA) {
  
  participants <- unique(data$ID)

  participantfits <- do.call(rbind,
                             lapply(participants,
                                    function(x) 
                                      generate.fits(data[data$ID==x,][1:(maxt[maxt$ID==x,]$max.point),],
                                                    numguess, detect.threshold,
                                                    detect.threshold.2)))
  
  return(participantfits)
}

participant.params.minsse <- function(fits) {
  
  participants <- unique(fits$ID)
  
  participantparams <- data.frame(do.call(rbind, 
                               lapply(participants, 
                                      function(x) params.minsse(fits[fits$ID == x,]))))
  
  colnames(participantparams) <- c("ID", "p1", "p2",
                               "p3", "p4")
  
  participantparams[, 2:5] <- apply(participantparams[, 2:5], 2, as.numeric)
  
  return(participantparams)
}

# Setting the endpoint as the first local minimum
noro.maxt.firstmin <- do.call(rbind, 
                         lapply(noroParticipants, 
                                function(x) data.frame(x, 
                                                       first.local.min(noro.FShedData[noro.FShedData$ID == x,]))))
colnames(noro.maxt.firstmin) <- c('ID', 'max.point')

# SARS-CoV-2 data
sars.maxt.firstmin <- do.call(rbind, 
                         lapply(sarsParticipants, 
                                function(x) data.frame(x, 
                                                       first.local.min(sars.TShedData[sars.TShedData$ID == x,]))))
colnames(sars.maxt.firstmin) <- c('ID', 'max.point')

# Setting the endpoint as the first time that viral load decreases to the 
# initial viral load or lower

noro.maxt.backtoinitial <- do.call(rbind, 
                         lapply(noroParticipants, 
                                function(x) data.frame(x, 
                                                       back.to.initial(noro.FShedData[noro.FShedData$ID == x,]))))
colnames(noro.maxt.backtoinitial) <- c('ID', 'max.point')

# SARS-CoV-2 data
sars.maxt.backtoinitial <- do.call(rbind, 
                         lapply(sarsParticipants, 
                                function(x) data.frame(x, 
                                                       back.to.initial(sars.TShedData[sars.TShedData$ID == x,]))))
colnames(sars.maxt.backtoinitial) <- c('ID', 'max.point')

```

Note that because this model is statistical, not deterministic, the parameters
will fluctuate with each run. We account for this uncertainty by 1) showing
the range of values for each parameter in the fits with the lowest 5% SSE, and
by 2) producing confidence bounds in our analyses via bootstrapping/case 
resampling.
```{r, eval = FALSE}
noro.fits.localmin <- all.fits(noro.FShedData, noro.maxt.firstmin, 100,
                               detect.threshold = log10(noro.detection.threshold), 
                               detect.threshold.2 = log10(noro.detection.threshold.2))
noro.params.localmin <- participant.params.minsse(noro.fits.localmin)

sars.fits.localmin <- all.fits(sars.TShedData, sars.maxt.firstmin, 100,
                               detect.threshold = log10(sars.detection.threshold))
sars.params.localmin <- participant.params.minsse(sars.fits.localmin)

noro.fits.backtoinitial <- all.fits(noro.FShedData, noro.maxt.backtoinitial, 100,
                                    detect.threshold = log10(noro.detection.threshold), 
                                    detect.threshold.2 = log10(noro.detection.threshold.2))
noro.params.backtoinitial <- participant.params.minsse(noro.fits.backtoinitial)

sars.fits.backtoinitial <- all.fits(sars.TShedData, sars.maxt.backtoinitial, 100,
                                    detect.threshold = log10(sars.detection.threshold))
sars.params.backtoinitial <- participant.params.minsse(sars.fits.backtoinitial)
```

As a result, we saved the results of our first iteration in a csv file.
```{r}
write.csv(noro.fits.localmin, "Atmar Fits local minimum.csv", row.names=F)
write.csv(noro.params.localmin, "Atmar Parameters local minimum.csv", row.names=F)
write.csv(sars.fits.localmin, "Zhou Fits local minimum.csv", row.names=F)
write.csv(sars.params.localmin, "Zhou Parameters local minimum.csv", row.names=F)

write.csv(noro.fits.backtoinitial, "Atmar Fits back to initial.csv", row.names=F)
write.csv(noro.params.backtoinitial, "Atmar Parameters back to initial.csv", row.names=F)
write.csv(sars.fits.backtoinitial, "Zhou Fits back to initial.csv", row.names=F)
write.csv(sars.params.backtoinitial, "Zhou Parameters back to initial.csv", row.names=F)

noro.fits.localmin <- read.csv("Ge Fits local minimum.csv")
noro.params.localmin <- read.csv("Ge Parameters local minimum.csv")

sars.fits.localmin <- read.csv("Zhou Fits local minimum.csv")
sars.params.localmin <- read.csv("Zhou Parameters local minimum.csv")

noro.fits.backtoinitial <- read.csv("Ge Fits back to initial.csv")
noro.params.backtoinitial <- read.csv("Ge Parameters back to initial.csv")

sars.fits.backtoinitial <- read.csv("Zhou Fits back to initial.csv")
sars.params.backtoinitial <- read.csv("Zhou Parameters back to initial.csv")
```

We plot the range (max and min) of each parameter for the fitting runs with the
lowest 5% SSE from each infection to assess parameter uncertainty.
```{r}
param.ranges <- function(fits) {
  
  bottom5sse <- quantile(fits$sse, probs=c(0.05))
    
  bottom5params <- fits[fits$convergence==0 &
                            fits$sse <= bottom5sse,][1:5]
  
  return(bottom5params)
}

plot.paramranges <- function(ranges, cols) {
  
  participants <- unique(ranges$ID)
  globalmin <- min(ranges[,2])
  globalmax <- max(ranges[,2])
  
  param.name <- colnames(ranges)[2]
  
  if (param.name == 'p1') {
    xlab <- "Peak Viral Load"
    label <- "A"
  } else if (param.name == 'p2') {
    xlab <- expression(Viral~replication~rate~(hour^-1))
    label <- "B"
  } else if (param.name == 'p3') {
    xlab <- "Time of Peak Viral Load (hour)"
    label <- "C"
  } else {
    xlab <- expression(Viral~decay~rate~(hour^-1))
    label <- "D"
  }
  
  plot(1, type="n", xlab="", ylab="", yaxt="n", 
     xlim=c(globalmin, globalmax), ylim=c(0, 20), bty="l",
     mgp=c(2, 0.75, 0), tck=-0.04)
  
  text(.975 * (globalmax - globalmin) + globalmin, .975*20, label)
  
  axis(2, lwd.ticks=0, labels = NA)

  for (i in 1:length(participants)) {
    participantrange <- ranges[ranges$ID == participants[i],]
    p.range <- c(min(participantrange[,2]), max(participantrange[,2])) 
    
    points(p.range, rep(i, 2), col=cols[i])
    lines(p.range, rep(i, 2), col=cols[i])

  }
  
  mtext(xlab, side = 1, line= 2.3)
  
}

# Norovirus data
topnoro.params.localmin <- do.call(rbind, 
                              lapply(noroParticipants,
                                     function(x) param.ranges(noro.fits.localmin[noro.fits.localmin$ID == x,]))) 

cols <- glasbey.colors(length(noroParticipants)+1)[2:(length(noroParticipants)+1)]

tiff(filename="Parameter Ranges Norovirus.tiff", height=6.5, width=6.5, units = "in",
    res=500)
par(mfrow=c(2, 2), mai=c(0.72, 0.42, 0.42, 0.02))

{plot.paramranges(topnoro.params.localmin[, c(1, 2)], cols)
mtext("Parameter ranges with the lowest 5% of SSE (norovirus)", side = 3, adj=-0.75, line=1.5)
plot.paramranges(topnoro.params.localmin[, c(1, 3)], cols)
legend(10, 10, legend=noroParticipants, ncol=3,
         lty=1, col=cols, cex=0.6)
plot.paramranges(topnoro.params.localmin[, c(1, 4)], cols)
mtext("Participants", side = 2, line=1.2)
plot.paramranges(topnoro.params.localmin[, c(1, 5)], cols)}

dev.off()

# SARS-CoV-2 Data
topsars.params.localmin <- do.call(rbind, 
                              lapply(sarsParticipants,
                                     function(x) param.ranges(sars.fits.localmin[sars.fits.localmin$ID == x,]))) 

cols <- glasbey.colors(length(sarsParticipants)+1)[2:(length(sarsParticipants)+1)]

tiff(filename="Parameter Ranges SARS-CoV-2.tiff", height=6.5, width=6.5, units = "in",
    res=500)
par(mfrow=c(2, 2), mai=c(0.72, 0.42, 0.42, 0.02))

{plot.paramranges(topsars.params.localmin[, c(1, 2)], cols)
mtext("Parameter ranges with the lowest 5% of SSE (SARS-CoV-2)", side = 3, 
      adj=-0.6, line=1.5)
plot.paramranges(topsars.params.localmin[, c(1, 3)], cols)
plot.paramranges(topsars.params.localmin[, c(1, 4)], cols)
mtext("Participants", side = 2, line=1.2)
plot.paramranges(topsars.params.localmin[, c(1, 5)], cols)
legend(-20, 10, legend=sarsParticipants, ncol=3,
         lty=1, col=cols, cex=0.6)}
dev.off()
```

We proceed to plot trajectories for every individual.
```{r, fig.height = 7.5, fig.width = 3.5}
plot.viralload <- function(data, pars, sympdata, max.modeltime, cex=1,
                           xaxisLabs = T, yaxisLabs = T, col, pch, ylim.lower=3,
                           cex_labels, cex.axis, xlabel, ylabel, panellabel,
                           example=FALSE) {

  # Plotting real shedding values
  plot(1, col=col, xlim=c(0, 25), 
       ylim=c(ylim.lower, 15), xlab="", ylab="", xaxt="n", yaxt="n", cex.lab=cex, 
       cex.axis=cex, font.lab=2, bty='n')
  
  points(data$days, log10(data$y_virus), col=col, cex=cex, pch=pch)
  
  # Grey rectangles where viral shedding is a range (qualitative results)
  range.indexes <- which(is.na(data$y_virus) & grepl('POS', data$y_type))
  rectangle.x_vals <- lapply(range.indexes, function(x) c(data[x-1,]$days + 0.25, 
                                                          data[x+1,]$days - 0.25))
  lapply(rectangle.x_vals, function(x) 
    rect(x[1], log10(15000), x[2], log10(4e7), border=NA, col="lightgray"))
  
  
  # Plotting simulated VL
  linet <- seq(min(data[!is.na(data$y_virus),]$days) * 24, 
                 data[max.modeltime,]$days * 24, by=1)

  predictedvl <- generate.estimates(linet, pars)

  lines(linet / 24, predictedvl, col=col)
  
  # Symptom onset
  symp.onset <- sympdata$incubation
  
  abline(v = symp.onset, col=col, lty=2)
  
  # Peak shedding
  t.peak <- linet[which(predictedvl == max(predictedvl))]/24

  if (is.na(symp.onset)) {
    redpointstyle <- NA
  } else if (t.peak < symp.onset) {
    redpointstyle <- 19
  } else {
    redpointstyle <- 1
  }
  
  if (!example) {
    text(x=18, y=12, labels=unique(data$ID), font=2)
  }
  
  ylabel.line <- 4.5
  xlabel.line <- 3.5

  points(t.peak, max(predictedvl), pch=redpointstyle,
         col="red", lwd=cex, cex=cex)

  xLabs = NA
  if(xaxisLabs==T){
        xLabs = c(expression(0),
                  expression(5),
                  expression(10),
                  expression(15),
                  expression(20),
                  expression(25))}
  axis(1, cex.axis=cex.axis, at=seq(0, 25, by=5), labels = xLabs,
       line=1.25,
       mgp=c(2, 0.75, 0), tck=-0.04)
  
  yLabs = NA
  if(yaxisLabs==T){
      yLabs = c(expression(10^3),
                expression(10^6),
                expression(10^9),
                expression(10^12),
                expression(10^15))}
  axis(2, cex.axis=cex.axis, at=seq(3, 15, by=3), 
       labels=yLabs,
       line=1.25,
       las=2, 
       mgp=c(2, 1, 0), tck= -0.04)
  
  if (example) {
    mtext(text=xlabel, side=1, line=xlabel.line, cex=cex_labels, adj=3.25)
    mtext(text=ylabel, side=2, line=ylabel.line, cex=cex_labels, adj=5)
    text(.93 * 25, .93 * (15-3) + 3, panellabel, cex=cex_labels,
         font=2)
  }
  else {
    mtext(text=xlabel, side=1, line=xlabel.line, cex=cex_labels)
    mtext(text=ylabel, side=2, line=ylabel.line, cex=cex_labels)
  }
  
}
```

First we show examples of pre- and post-symptomatic transmission.
```{r}
sars.color <- "#56b3e9"
noro.color <- "#871a6e"
sars.pch <- 5
noro.pch <- 2

tiff(filename="Example Trajectories.tiff", height=5, 
     width=6.5, units = "in", res=500)

participantexamples <- c('Throat_16', 'Throat_5', 'ID12')

par(mfrow=c(2, 2), mai=c(0.42, 0.52, 0.52, 0.1), oma=c(2, 3, 1, 2))

for(i in 1: length(participantexamples)) {
  
  p <- participantexamples[i]
  
  if(grepl('Throat', p)) {
    participantdata <- sars.TShedData[sars.TShedData$ID==p, ]
    sympdata <- sars.IncubData[sars.IncubData$ID==p, ]
    params <- unlist(sars.params.localmin[sars.params.localmin$ID==p,][2:5])
    maxt <- sars.maxt.firstmin[sars.maxt.firstmin$ID==p,]$max.point
    col <- sars.color
    pch <- sars.pch
  } else {
    participantdata <- noro.FShedData[noro.FShedData$ID==p, ]
    sympdata <- noro.IncubData[noro.IncubData$ID == p, ]
    params <- unlist(noro.params.localmin[noro.params.localmin$ID==p,][2:5])
    maxt <- noro.maxt.firstmin[noro.maxt.firstmin$ID==p,]$max.point
    col <- noro.color
    pch <- noro.pch
  }

  if (i == 3) {
    ylabel <- "Viral shedding"
    xlabel <- "Hours after inoculation"
  } else {
    ylabel <- ""
    xlabel <- ""
  }
  
  plot.viralload(data=participantdata, 
                 pars=params, 
                 sympdata=sympdata, 
                 max.modeltime=maxt,
                 cex=1.33, 
                 col=col,
                 pch=pch,
                 cex_labels=1,
                 cex.axis=1,
                 xlabel=xlabel,
                 ylabel=ylabel,
                 panellabel=LETTERS[i],
                 example=TRUE)
  
  if (i == 1) {
    mtext("Post-symptomatic", side=3, line=2, font=2)
  } else if (p==participantexamples[2]) {
    mtext("Pre-symptomatic", side=3, line=2, font=2)
  }
}

plot.new()

legend("center", legend=c("SARS-CoV-2", "Norovirus"), col=c(sars.color, noro.color),
       pch=c(5, 2))

dev.off()
```

We then plot the trajectories of each participant from both studies using both 
sets of endpoints. Since they produce qualitatively similar fits, we arbitrarily 
use the first local minimum in pathogen shedding as the endpoint going forward.
```{r, fig.height = 7.5, fig.width = 6.5}
tiff("Participant trajectories norovirus.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(noroParticipants)) {
  
  p <- noroParticipants[i]
  participantdata <- noro.FShedData[noro.FShedData$ID==p, ]
  sympdata <- noro.IncubData[noro.IncubData$ID==p, ]
  params <- unlist(noro.params.localmin[noro.params.localmin$ID==p,][2:5])
  maxt <- noro.maxt.firstmin[noro.maxt.firstmin$ID==p,]$max.point
  col <- noro.color
  pch <- noro.pch
  
  if (i==16) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding")
  } else if (i %in% seq(1, 16, by=3)) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      xaxisLabs=F,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding")
  } else if (i %in% c(15, 17)) {
     plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      yaxisLabs=F,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="")
  } else {
    plot.viralload(
    data=participantdata,
    pars=params,
    sympdata=sympdata,
    max.modeltime=maxt,
    xaxisLabs=F,
    yaxisLabs=F,
    col=col,
    pch=pch,
    ylim.lower = 7,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="")
  }
}
dev.off()

#Repeating with a different endpoint

tiff("Participant trajectories norovirus alternative endpoint.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(noroParticipants)) {
  
  p <- noroParticipants[i]
  participantdata <- noro.FShedData[noro.FShedData$ID==p, ]
  sympdata <- noro.IncubData[noro.IncubData$ID==p, ]
  params <- unlist(noro.params.backtoinitial[noro.params.backtoinitial$ID==p,][2:5])
  maxt <- noro.maxt.backtoinitial[noro.maxt.backtoinitial$ID==p,]$max.point
  col <- noro.color
  pch <- noro.pch
  
  if (i==16) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      col=noro.color,
      pch=noro.pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding")
  } else if (i %in% seq(1, 16, by=3)) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      xaxisLabs=F,
      col=noro.color,
      pch=noro.pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding")
  } else if (i %in% c(15, 17)) {
     plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      max.modeltime=maxt,
      yaxisLabs=F,
      col=noro.color,
      pch=noro.pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="")
  } else {
    plot.viralload(
    data=participantdata,
    pars=params,
    sympdata=sympdata,
    max.modeltime=maxt,
    xaxisLabs=F,
    yaxisLabs=F,
    col=noro.color,
    pch=noro.pch,
    ylim.lower = 7,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="")
  }
}

dev.off()

tiff("Participant Trajectories SARS-CoV-2.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(7, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(sarsParticipants)) {
  
  p <- sarsParticipants[i]
  participantdata <- sars.TShedData[sars.TShedData$ID==p, ]
  sympdata <- sars.IncubData[sars.IncubData$ID==p, ]
  params <- unlist(sars.params.localmin[sars.params.localmin$ID==p,][2:5])
  maxt <- sars.maxt.firstmin[sars.maxt.firstmin$ID==p,]$max.point
  
  if (i %in% seq(1, 13, by=3)) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      xaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding")
  } else if (i %in% c(17, 18)) {
     plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      yaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="")
  } else if (i==16) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding")
  } else {
    plot.viralload(
    data=participantdata,
    sympdata=sympdata,
    pars=params,
    max.modeltime=maxt,
    xaxisLabs=F,
    yaxisLabs=F,
    col=sars.color,
    pch=sars.pch,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="")
  }
}

dev.off()

# Repeating with a different endpoint

tiff("Participant Trajectories SARS-CoV-2 alternative endpoint.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(7, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(sarsParticipants)) {
  
  p <- sarsParticipants[i]
  participantdata <- sars.TShedData[sars.TShedData$ID==p, ]
  sympdata <- sars.IncubData[sars.IncubData$ID==p, ]
  params <- unlist(sars.params.backtoinitial[sars.params.backtoinitial$ID==p,][2:5])
  maxt <- sars.maxt.backtoinitial[sars.maxt.backtoinitial$ID==p,]$max.point
  
  if (i %in% seq(1, 13, by=3)) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      xaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding")
  } else if (i %in% c(17, 18)) {
     plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      yaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="")
  } else if (i==16) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      max.modeltime=maxt,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding")
  } else {
    plot.viralload(
    data=participantdata,
    sympdata=sympdata,
    pars=params,
    max.modeltime=maxt,
    xaxisLabs=F,
    yaxisLabs=F,
    col=sars.color,
    pch=sars.pch,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="")
  }
}

dev.off()
```

We exclude some participants due to unclear peaks.
```{r}
noroParticipants.all <- unique(noro.FShedData$ID)
noroParticipants <- unique(noro.FShedData[!(noro.FShedData$ID %in% c('ID5', 'ID20')),]$ID)

sarsParticipants.all <- unique(sars.TShedData$ID)
sarsParticipants <- unique(sars.TShedData[!(sars.TShedData$ID %in% c('Throat_1', 'Throat_12', 'Throat_13', 'Throat_14')),]$ID)
```

Now, we calculate the delay between symptom onset and peak pathogen load for
each participant in each study. The delay is in days.
```{r}
calculate.peaktime <- function(vldata, maxt, params) {
  
  t <- seq(0, vldata[maxt$max.point,]$days*24, 
        by=1)
  
  predictedvl <- generate.estimates(t, params)
  
  peak.vl.time <- t[which(predictedvl==max(predictedvl))]
  
  return(peak.vl.time)
}

calculate.delay.conservative <- function(vldata, sympdata, params, maxt) {
  
  peak.vl.time <- calculate.peaktime(vldata, maxt, params)

  delay <- peak.vl.time - (sympdata$incubation * 24)
  
  return(delay/24)
}


delays.conservative.noro <- do.call(rbind,  
                               lapply(noroParticipants, 
                                      function(x) 
                                        data.frame(x, 
                                                   calculate.delay.conservative(noro.FShedData[noro.FShedData$ID == x, ],
                                                        noro.IncubData[noro.IncubData$ID == x,],
                                                        unlist(noro.params.localmin[noro.params.localmin$ID == x,][2:5]), 
                                                        noro.maxt.firstmin[noro.maxt.firstmin$ID == x,]))))

colnames(delays.conservative.noro) <- c('ID', 'delay')

delays.conservative.sars <- do.call(rbind,  
                               lapply(sarsParticipants, 
                                      function(x) 
                                        data.frame(x, 
                                                   calculate.delay.conservative(sars.TShedData[sars.TShedData$ID == x, ],
                                                        sars.IncubData[sars.IncubData$ID == x,],
                                                        unlist(sars.params.localmin[sars.params.localmin$ID == x,][2:5]), 
                                                        sars.maxt.firstmin[sars.maxt.firstmin$ID == x,]))))

colnames(delays.conservative.sars) <- c('ID', 'delay')
```

We run a linear regression between the delay between symptom onset and peak
pathogen load and each of the 4 parameters from the statistical model as well
as between viral growth and decay rate. In the Ge *et al.*, 2023 study, 
p1, p2, and p3, i.e. peak viral load, viral replication rate, and time of peak 
viral load, respectively, were found to have significant correlations with delay. 
```{r}
params.and.delays.conservative.noro <- merge(noro.params.localmin, delays.conservative.noro)

# Peak viral load vs. delay
lm.fit.p1.delay.conservative.noro <- 
  lm(delay ~ p1, data=params.and.delays.conservative.noro)
p1.delay.conservative.pval.noro <- 
  summary(lm.fit.p1.delay.conservative.noro)$coefficients[2,4]
p1.delay.conservative.pval.noro

# Viral replication rate vs. delay
lm.fit.p2.delay.conservative.noro <- 
  lm(delay ~ p2, data=params.and.delays.conservative.noro)
p2.delay.conservative.pval.noro <- 
  summary(lm.fit.p2.delay.conservative.noro)$coefficients[2,4]
p2.delay.conservative.pval.noro
p2.delay.conservative.rsquared.noro <- summary(lm.fit.p2.delay.conservative.noro)$r.squared
p2.delay.conservative.rsquared.noro

# Time of peak viral load vs. delay
lm.fit.p3.delay.conservative.noro <- 
  lm(delay ~ p3, data=params.and.delays.conservative.noro)
p3.delay.conservative.pval.noro <- 
  summary(lm.fit.p3.delay.conservative.noro)$coefficients[2,4]
p3.delay.conservative.pval.noro

# Viral decay rate vs. delay
lm.fit.p4.delay.conservative.noro <- 
  lm(delay ~ p4, data=params.and.delays.conservative.noro)
p4.delay.conservative.pval.noro <- 
  summary(lm.fit.p4.delay.conservative.noro)$coefficients[2,4]
p4.delay.conservative.pval.noro

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative.noro <- 
  lm(p4 ~ p2, data=params.and.delays.conservative.noro)
p4.p2.conservative.pval.noro <- 
  summary(lm.fit.p4.p2.conservative.noro)$coefficients[2,4]
p4.p2.conservative.pval.noro
```

There were no significant correlations found within the Zhou *et al.*, 2023 study,
looking at the same parameters.
```{r}
params.and.delays.conservative.sars <- merge(sars.params.localmin, delays.conservative.sars)

# Peak viral load vs. delay
lm.fit.p1.delay.conservative.sars <- 
  lm(delay ~ p1, data=params.and.delays.conservative.sars)
p1.delay.conservative.pval.sars <- 
  summary(lm.fit.p1.delay.conservative.sars)$coefficients[2,4]
p1.delay.conservative.pval.sars

# Viral replication rate vs. delay
lm.fit.p2.delay.conservative.sars <- 
  lm(delay ~ p2, data=params.and.delays.conservative.sars)
p2.delay.conservative.pval.sars <- 
  summary(lm.fit.p2.delay.conservative.sars)$coefficients[2,4]
p2.delay.conservative.pval.sars

# Time of peak viral load vs. delay
lm.fit.p3.delay.conservative.sars <- 
  lm(delay ~ p3, data=params.and.delays.conservative.sars)
p3.delay.conservative.pval.sars <- 
  summary(lm.fit.p3.delay.conservative.sars)$coefficients[2,4]
p3.delay.conservative.pval.sars

# Viral decay rate vs. delay
lm.fit.p4.delay.conservative.sars <- 
  lm(delay ~ p4, data=params.and.delays.conservative.sars)
p4.delay.conservative.pval.sars <- 
  summary(lm.fit.p4.delay.conservative.sars)$coefficients[2,4]
p4.delay.conservative.pval.sars

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative.sars <- 
  lm(p4 ~ p2, data=params.and.delays.conservative.sars)
p4.p2.conservative.pval.sars <- 
  summary(lm.fit.p4.p2.conservative.sars)$coefficients[2,4]
p4.p2.conservative.pval.sars
```

We now look at a non-conservative measure of pre-symptomatic transmission, 
where we calculate the delay between shedding onset and symptom onset. A negative 
delay still indicates pre-symptomatic transmission.
```{r}
calculate.delay.nonconservative <- function(vldata, sympdata) {

  delay <- shed.onset(vldata) - sympdata$incubation 
  
  return(delay)
}

delays.nonconservative.noro <- do.call(rbind,  
                               lapply(noroParticipants.all, 
                                      function(x) 
                                        data.frame(x, 
                                                   calculate.delay.nonconservative(noro.FShedData[noro.FShedData$ID == x, ],
                                                        noro.IncubData[noro.IncubData$ID == x,]))))

colnames(delays.nonconservative.noro) <- c('ID', 'delay')

delays.nonconservative.sars <- do.call(rbind,  
                               lapply(sarsParticipants.all, 
                                      function(x) 
                                        data.frame(x, 
                                                   calculate.delay.nonconservative(sars.TShedData[sars.TShedData$ID == x, ],
                                            sars.IncubData[sars.IncubData$ID == x,]))))

colnames(delays.nonconservative.sars) <- c('ID', 'delay')
```

As before, we test for any significant correlations between the statistical 
model parameters and the duration of the latency period. With this 
non-conservative estimate of pre-symptomatic transmission, we no longer find any 
statistically significant correlations in either dataset.
```{r}
params.and.delays.nonconservative.noro <- merge(noro.params.localmin, delays.nonconservative.noro)

lm.fit.p1.delay.nonconservative.noro <- 
  lm(delay ~ p1, data=params.and.delays.nonconservative.noro)
p1.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p1.delay.nonconservative.noro)$coefficients[2,4]
p1.delay.nonconservative.pval.noro

lm.fit.p2.delay.nonconservative.noro <- 
  lm(delay ~ p2, data=params.and.delays.nonconservative.noro)
p2.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p2.delay.nonconservative.noro)$coefficients[2,4]
p2.delay.nonconservative.pval.noro

lm.fit.p3.delay.nonconservative.noro <- 
  lm(delay ~ p3, data=params.and.delays.nonconservative.noro)
p3.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p3.delay.nonconservative.noro)$coefficients[2,4]
p3.delay.nonconservative.pval.noro

lm.fit.p4.delay.nonconservative.noro <- 
  lm(delay ~ p4, data=params.and.delays.nonconservative.noro)
p4.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p4.delay.nonconservative.noro)$coefficients[2,4]
p4.delay.nonconservative.pval.noro

params.and.delays.nonconservative.sars <- merge(sars.params.localmin, delays.nonconservative.sars)

lm.fit.p1.delay.nonconservative.sars <- 
  lm(delay ~ p1, data=params.and.delays.nonconservative.sars)
p1.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p1.delay.nonconservative.sars)$coefficients[2,4]
p1.delay.nonconservative.pval.sars

lm.fit.p2.delay.nonconservative.sars <- 
  lm(delay ~ p2, data=params.and.delays.nonconservative.sars)
p2.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p2.delay.nonconservative.sars)$coefficients[2,4]
p2.delay.nonconservative.pval.sars

lm.fit.p3.delay.nonconservative.sars <- 
  lm(delay ~ p3, data=params.and.delays.nonconservative.sars)
p3.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p3.delay.nonconservative.sars)$coefficients[2,4]
p3.delay.nonconservative.pval.sars

lm.fit.p4.delay.nonconservative.sars <- 
  lm(delay ~ p4, data=params.and.delays.nonconservative.sars)
p4.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p4.delay.nonconservative.sars)$coefficients[2,4]
p4.delay.nonconservative.pval.sars
```

We also run a linear regression between pre-symptomatic transmission potential,
measured as average shedding prior to symptoms, and time of symptom onset. 
There is a correlation in the norovirus data, but there are only 5 data points
included in that analysis. There is no correlation in the SARS-CoV-2 data.
```{r}
presymp.transmission.potential <- function(vldata, sympdata) {
  
  vl.latent <- vldata[vldata$days < sympdata$incubation,]
  
  if(length(vl.latent[grepl('~', vl.latent$y_type),]$y_virus) > 0) {
    vl.latent[grepl('~', vl.latent$y_type),]$y_virus <- mean(c(15000, 4e7))
  }

  avg.shedding.latent <- log10(mean(vl.latent$y_virus, na.rm=T))
  
  return(avg.shedding.latent)
  
}

transmission.latent.noro <- do.call(rbind, 
                               lapply(noroParticipants.all, 
                                      function(x) 
                                        data.frame(x,
                                                   presymp.transmission.potential(noro.FShedData[noro.FShedData$ID == x,],
                                                                                  noro.IncubData[noro.IncubData$ID == x,]),
                                                   noro.IncubData[noro.IncubData$ID == x,]$incubation)))
  
colnames(transmission.latent.noro) <- c('ID', 'transmission_potential', 'symptom_onset')
transmission.latent.noro <- transmission.latent.noro[!is.na(transmission.latent.noro$transmission_potential),]

transmission.latent.fit.noro <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent.noro)
transmission.latent.pval.noro <- 
  summary(transmission.latent.fit.noro)$coefficients[2,4]

transmission.latent.pval.noro

transmission.latent.rsquared.noro <- summary(transmission.latent.fit.noro)$r.squared

transmission.latent.rsquared.noro

# SARS-CoV-2 data 

transmission.latent.sars <- do.call(rbind, 
                               lapply(sarsParticipants.all, 
                                      function(x) 
                                        data.frame(x,
                                                   presymp.transmission.potential(sars.TShedData[sars.TShedData$ID == x,],
                                                                                  sars.IncubData[sars.IncubData$ID == x,]),
                                                   sars.IncubData[sars.IncubData$ID == x,]$incubation)))
  
colnames(transmission.latent.sars) <- c('ID', 'transmission_potential', 'symptom_onset')
transmission.latent.sars <- transmission.latent.sars[!is.na(transmission.latent.sars$transmission_potential),]

transmission.latent.fit.sars <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent.sars)
transmission.latent.pval.sars <- 
  summary(transmission.latent.fit.sars)$coefficients[2,4]

transmission.latent.pval.sars
```

We ran the same analysis using cumulative pre-symptomatic shedding to measure
pre-symptomatic transmission potential. The results were exactly the same: 
statistically significant correlation found for norovirus (still only 5 data 
points), and no statistically significant correlation found for SARs-CoV-2.
```{r}
presymp.transmission.potential.cum <- function(vldata, sympdata) {
  
  vl.latent <- vldata[vldata$days < sympdata$incubation,]
  
  if(length(vl.latent[grepl('~', vl.latent$y_type),]$y_virus) > 0) {
    vl.latent[grepl('~', vl.latent$y_type),]$y_virus <- mean(c(15000, 4e7))
  }
  
  vl.latent[is.na(vl.latent$y_virus),]$y_virus <- 0

  if (length(vl.latent$y_virus) == 1) {
    auc.vl.latent <- log10(vl.latent$y_virus)
  } else {
    auc.vl.latent <- log10(trapz(vl.latent$days, vl.latent$y_virus))
  }
  
  if (is.infinite(auc.vl.latent)) {return(NaN)} else {return(auc.vl.latent)}
}

transmission.latent.cum.noro <- do.call(rbind, 
                               lapply(noroParticipants.all, 
                                      function(x) 
                                        data.frame(x,
                                                   presymp.transmission.potential.cum(noro.FShedData[noro.FShedData$ID == x,],
                                                                                      noro.IncubData[noro.IncubData$ID == x,]),
                                                   noro.IncubData[noro.IncubData$ID == x,]$incubation)))
  
colnames(transmission.latent.cum.noro) <- c('ID', 'transmission_potential', 'symptom_onset')

transmission.latent.cum.fit.noro <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent.cum.noro)
transmission.latent.cum.pval.noro <- 
  summary(transmission.latent.cum.fit.noro)$coefficients[2,4]

transmission.latent.cum.pval.noro

# SARS-CoV-2 data

transmission.latent.cum.sars <- do.call(rbind, 
                               lapply(sarsParticipants.all, 
                                      function(x) 
                                        data.frame(x,
                                                   presymp.transmission.potential.cum(sars.TShedData[sars.TShedData$ID == x,],
                                                                                      sars.IncubData[sars.IncubData$ID == x,]),
                                                   sars.IncubData[sars.IncubData$ID == x,]$incubation)))
  
colnames(transmission.latent.cum.sars) <- c('ID', 'transmission_potential', 'symptom_onset')

transmission.latent.cum.fit.sars <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent.cum.sars)
transmission.latent.cum.pval.sars <- 
  summary(transmission.latent.cum.fit.sars)$coefficients[2,4]

transmission.latent.cum.pval.sars
```

Lastly, we break down the correlation between pre-symptomatic transmission
and viral growth rate by examining the relationships between time of
peak shedding and viral growth rate as well as between symptom onset
and viral growth rate. For both, there is a significant correlation between
time of peak shedding and viral growth rate, but not between symptom onset 
and viral growth rate.
```{r, fig.height = 3, fig.width = 6.5}
participant.peaktimes.noro <- do.call(rbind, 
                               lapply(noroParticipants, 
                                      function(x) 
                                        data.frame(x,
                                                   calculate.peaktime(noro.FShedData[noro.FShedData$ID == x,],
                                                                      noro.maxt.firstmin[noro.maxt.firstmin$ID == x,],
                                                                      unlist(noro.params.localmin[noro.params.localmin$ID==x, 2:5])))))

colnames(participant.peaktimes.noro) <- c("ID", "peaktime")

params.peaktime.data.noro <- merge(merge(noro.params.localmin, participant.peaktimes.noro),
                              delays.conservative.noro)

lm.fit.p2.peaktime.noro <- lm(peaktime/24 ~ p2, data=params.peaktime.data.noro)
p2.peaktime.pval.noro <- summary(lm.fit.p2.peaktime.noro)$coefficients[2,4]
p2.peaktime.pval.noro
p2.peaktime.rsquared.noro <- summary(lm.fit.p2.peaktime.noro)$r.squared
p2.peaktime.rsquared.noro

params.symp.data.noro <- merge(merge(noro.params.localmin, noro.IncubData[,c(1, 3)]),
                          delays.conservative.noro)
lm.fit.p2.symponset.noro <- lm(incubation ~ p2, data=params.symp.data.noro)
p2.symponset.pval.noro <- summary(lm.fit.p2.symponset.noro)$coefficients[2,4]
p2.symponset.pval.noro

# SARS-CoV-2 data
participant.peaktimes.sars <- do.call(rbind, 
                               lapply(sarsParticipants, 
                                      function(x) 
                                        data.frame(x,
                                                   calculate.peaktime(sars.TShedData[sars.TShedData$ID == x,],
                                                                      sars.maxt.firstmin[sars.maxt.firstmin$ID == x,],
                                                                      unlist(sars.params.localmin[sars.params.localmin$ID==x, 2:5])))))

colnames(participant.peaktimes.sars) <- c("ID", "peaktime")

params.peaktime.data.sars <- merge(merge(sars.params.localmin, participant.peaktimes.sars),
                              delays.conservative.sars)

lm.fit.p2.peaktime.sars <- lm(peaktime/24 ~ p2, data=params.peaktime.data.sars)
p2.peaktime.pval.sars <- summary(lm.fit.p2.peaktime.sars)$coefficients[2,4]
p2.peaktime.pval.sars
p2.peaktime.rsquared.sars <- summary(lm.fit.p2.peaktime.sars)$r.squared
p2.peaktime.rsquared.sars

params.symp.data.sars <- merge(merge(sars.params.localmin, sars.IncubData),
                          delays.conservative.sars)
lm.fit.p2.symponset.sars <- lm(incubation ~ p2, data=params.symp.data.sars)
p2.symponset.pval.sars <- summary(lm.fit.p2.symponset.sars)$coefficients[2,4]
p2.symponset.pval.sars
```

We use bootstrapping to create 95% confidence bounds for each correlation
we show in Figure 3.
```{r}
iters <- 1000

delay.sample.slopes.noro <- c()
delay.sample.intercepts.noro <- c()

delay.nonconservative.sample.slopes.noro <- c()
delay.nonconservative.sample.intercepts.noro <- c()

peakshedding.sample.slopes.noro <- c()
peakshedding.sample.intercepts.noro <- c()

symponset.sample.slopes.noro <- c()
symponset.sample.intercepts.noro <- c()

presymp.shed.sample.slopes.noro <- c()
presymp.shed.sample.intercepts.noro <- c()

delay.sample.slopes.sars <- c()
delay.sample.intercepts.sars <- c()

delay.nonconservative.sample.slopes.sars <- c()
delay.nonconservative.sample.intercepts.sars <- c()

peakshedding.sample.slopes.sars <- c()
peakshedding.sample.intercepts.sars <- c()

symponset.sample.slopes.sars <- c()
symponset.sample.intercepts.sars <- c()

presymp.shed.sample.slopes.sars <- c()
presymp.shed.sample.intercepts.sars <- c()

for (i in 1:iters) {
  
  # Panel B: delay between symptom onset and peak viral shedding vs. viral 
  # replication rate
  delay.resamp.noro <- params.and.delays.conservative.noro[
    sample(1:nrow(params.and.delays.conservative.noro), 
                         size=nrow(params.and.delays.conservative.noro),
                         replace=T), c(3, 6)]
  
  delay.slope.noro <- coef(lm(delay ~ p2, data=delay.resamp.noro))[2]
  delay.intercept.noro <- coef(lm(delay ~ p2, data=delay.resamp.noro))[1]
  
  delay.sample.slopes.noro[i] <- delay.slope.noro
  delay.sample.intercepts.noro[i] <- delay.intercept.noro
  
  delay.resamp.sars <- params.and.delays.conservative.sars[
    sample(1:nrow(params.and.delays.conservative.sars), 
                         size=nrow(params.and.delays.conservative.sars),
                         replace=T), c(3, 6)]
  
  delay.slope.sars <- coef(lm(delay ~ p2, data=delay.resamp.sars))[2]
  delay.intercept.sars <- coef(lm(delay ~ p2, data=delay.resamp.sars))[1]
  
  delay.sample.slopes.sars[i] <- delay.slope.sars
  delay.sample.intercepts.sars[i] <- delay.intercept.sars
  
  # Panel B for Figure S7: delay between symptom onset and shedding onset vs. viral 
  # replication rate
  delay.nonconservative.resamp.noro <- params.and.delays.nonconservative.noro[
    sample(1:nrow(params.and.delays.nonconservative.noro), 
                         size=nrow(params.and.delays.nonconservative.noro),
                         replace=T), c(3, 6)]
  
  delay.nonconservative.slope.noro <- coef(lm(delay ~ p2, 
                                              data=delay.nonconservative.resamp.noro))[2]
  delay.nonconservative.intercept.noro <- coef(lm(delay ~ p2, 
                                                  data=delay.nonconservative.resamp.noro))[1]
  
  delay.nonconservative.sample.slopes.noro[i] <- delay.nonconservative.slope.noro
  delay.nonconservative.sample.intercepts.noro[i] <- delay.nonconservative.intercept.noro
  
  delay.nonconservative.resamp.sars <- params.and.delays.nonconservative.sars[
    sample(1:nrow(params.and.delays.nonconservative.sars), 
                         size=nrow(params.and.delays.nonconservative.sars),
                         replace=T), c(3, 6)]
  
  delay.nonconservative.slope.sars <- coef(lm(delay ~ p2, 
                                              data=delay.nonconservative.resamp.sars))[2]
  delay.nonconservative.intercept.sars <- coef(lm(delay ~ p2, 
                                                  data=delay.nonconservative.resamp.sars))[1]
  
  delay.nonconservative.sample.slopes.sars[i] <- delay.nonconservative.slope.sars
  delay.nonconservative.sample.intercepts.sars[i] <- delay.nonconservative.intercept.sars
  
  # Panel C: day of peak shedding vs. viral replication rate 
  peakshedding.resamp.noro <- params.peaktime.data.noro[
    sample(1:nrow(params.peaktime.data.noro), 
                         size=nrow(params.peaktime.data.noro),
                         replace=T), c(3, 6)]
  
  peakshedding.slope.noro <- coef(lm(peaktime ~ p2, data=peakshedding.resamp.noro))[2]
  peakshedding.intercept.noro <- coef(lm(peaktime ~ p2, data=peakshedding.resamp.noro))[1]
  
  peakshedding.sample.slopes.noro[i] <- peakshedding.slope.noro
  peakshedding.sample.intercepts.noro[i] <- peakshedding.intercept.noro
  
  peakshedding.resamp.sars <- params.peaktime.data.sars[
    sample(1:nrow(params.peaktime.data.sars), 
                         size=nrow(params.peaktime.data.sars),
                         replace=T), c(3, 6)]
  
  peakshedding.slope.sars <- coef(lm(peaktime ~ p2, data=peakshedding.resamp.sars))[2]
  peakshedding.intercept.sars <- coef(lm(peaktime ~ p2, data=peakshedding.resamp.sars))[1]
  
  peakshedding.sample.slopes.sars[i] <- peakshedding.slope.sars
  peakshedding.sample.intercepts.sars[i] <- peakshedding.intercept.sars
  
  # Panel D: day of symptom onset vs. viral replication rate 
  symponset.resamp.noro <- params.symp.data.noro[sample(1:nrow(params.symp.data.noro),
                                                        size=nrow(params.symp.data.noro),
                                                        replace=T), c(3, 6)]
  
  symponset.slope.noro <- coef(lm(incubation ~ p2, data=symponset.resamp.noro))[2]
  symponset.intercept.noro <- coef(lm(incubation ~ p2, data=symponset.resamp.noro))[1]
  
  symponset.sample.slopes.noro[i] <- symponset.slope.noro
  symponset.sample.intercepts.noro[i] <- symponset.intercept.noro
  
  symponset.resamp.sars <- params.symp.data.sars[sample(1:nrow(params.symp.data.sars),
                                                        size=nrow(params.symp.data.sars),
                                                        replace=T), c(3, 6)]
  
  symponset.slope.sars <- coef(lm(incubation ~ p2, data=symponset.resamp.sars))[2]
  symponset.intercept.sars <- coef(lm(incubation ~ p2, data=symponset.resamp.sars))[1]
  
  symponset.sample.slopes.sars[i] <- symponset.slope.sars
  symponset.sample.intercepts.sars[i] <- symponset.intercept.sars
  
  # Panel E: average pre-symptomatic shedding vs. symptom onset 
  presymp.shed.resamp.noro <- transmission.latent.noro[sample(1:nrow(transmission.latent.noro),
                                                              size=nrow(transmission.latent.noro),
                                                              replace=T), c(2, 3)] 
  
  presymp.shed.slope.noro <- coef(lm(transmission_potential ~ symptom_onset, 
                                     data=presymp.shed.resamp.noro))[2]
  presymp.shed.intercept.noro <- coef(lm(transmission_potential ~ symptom_onset,
                                         data=presymp.shed.resamp.noro))[1]
  
  presymp.shed.sample.slopes.noro[i] <- presymp.shed.slope.noro 
  presymp.shed.sample.intercepts.noro[i] <- presymp.shed.intercept.noro 
 
  presymp.shed.resamp.sars <- transmission.latent.sars[sample(1:nrow(transmission.latent.sars),
                                                              size=nrow(transmission.latent.sars),
                                                              replace=T), c(2, 3)] 
  
  presymp.shed.slope.sars <- coef(lm(transmission_potential ~ symptom_onset, 
                                     data=presymp.shed.resamp.sars))[2]
  presymp.shed.intercept.sars <- coef(lm(transmission_potential ~ symptom_onset,
                                         data=presymp.shed.resamp.sars))[1]
  
  presymp.shed.sample.slopes.sars[i] <- presymp.shed.slope.sars
  presymp.shed.sample.intercepts.sars[i] <- presymp.shed.intercept.sars  
}

# Panel B
lower.bound.delay.mod.noro <- c(quantile(delay.sample.slopes.noro, 0.025),
                                quantile(delay.sample.intercepts.noro, 0.025))

upper.bound.delay.mod.noro <- c(quantile(delay.sample.slopes.noro, 0.975), 
                                quantile(delay.sample.intercepts.noro, 0.975))

lower.bound.delay.mod.sars <- c(quantile(delay.sample.slopes.sars, 0.025),
                                quantile(delay.sample.intercepts.sars, 0.025))
                                
upper.bound.delay.mod.sars <- c(quantile(delay.sample.slopes.sars, 0.975),
                                quantile(delay.sample.intercepts.sars, 0.975))

# Panel B for Figure S7
lower.bound.delay.nonconservative.mod.noro <- c(quantile(delay.nonconservative.sample.slopes.noro, 0.025),
                                quantile(delay.nonconservative.sample.intercepts.noro, 0.025))

upper.bound.delay.nonconservative.mod.noro <- c(quantile(delay.nonconservative.sample.slopes.noro, 0.975), 
                                quantile(delay.nonconservative.sample.intercepts.noro, 0.975))

lower.bound.delay.nonconservative.mod.sars <- c(quantile(delay.nonconservative.sample.slopes.sars, 0.025),
                                quantile(delay.nonconservative.sample.intercepts.sars, 0.025))
                                
upper.bound.delay.nonconservative.mod.sars <- c(quantile(delay.nonconservative.sample.slopes.sars, 0.975),
                                quantile(delay.nonconservative.sample.intercepts.sars, 0.975))

# Panel C
lower.bound.peakshedding.mod.noro <- c(quantile(peakshedding.sample.slopes.noro, 0.025),
                                       quantile(peakshedding.sample.intercepts.noro, 0.025))

upper.bound.peakshedding.mod.noro <- c(quantile(peakshedding.sample.slopes.noro, 0.975),
                                         quantile(peakshedding.sample.intercepts.noro, 0.975))

lower.bound.peakshedding.mod.sars <- c(quantile(peakshedding.sample.slopes.sars, 0.025),
                                       quantile(peakshedding.sample.intercepts.sars, 0.025))
upper.bound.peakshedding.mod.sars <- c(quantile(peakshedding.sample.slopes.sars, 0.975),
                                       quantile(peakshedding.sample.intercepts.sars, 0.975))

# Panel D
lower.bound.symponset.mod.noro <- c(quantile(symponset.sample.slopes.noro, 0.025),
                                      quantile(symponset.sample.intercepts.noro, 0.025))

upper.bound.symponset.mod.noro <- c(quantile(symponset.sample.slopes.noro, 0.975),
                                    quantile(symponset.sample.intercepts.noro, 0.975))

lower.bound.symponset.mod.sars <- c(quantile(symponset.sample.slopes.sars, 0.025),
                                    quantile(symponset.sample.intercepts.sars, 0.025))

upper.bound.symponset.mod.sars <- c(quantile(symponset.sample.slopes.sars, 0.975),
                                      quantile(symponset.sample.intercepts.sars, 0.975))

# Panel E

# The norovirus data only has 5 people demonstrating pre-symptomatic shedding,
# so sometimes resampling will return the same person 5 times. We remove those
# samples (na.rm = T).
lower.bound.presymp.shed.mod.noro <- c(quantile(presymp.shed.sample.slopes.noro, 0.025,
                                                na.rm=T),
                                       quantile(presymp.shed.sample.intercepts.noro, 0.025,
                                                na.rm=T))

upper.bound.presymp.shed.mod.noro <- c(quantile(presymp.shed.sample.slopes.noro, 0.975,
                                                na.rm=T),
                                       quantile(presymp.shed.sample.intercepts.noro, 0.975,
                                                na.rm=T))

lower.bound.presymp.shed.mod.sars <- c(quantile(presymp.shed.sample.slopes.sars, 
                                                0.025, na.rm=T),
                                       quantile(presymp.shed.sample.intercepts.sars, 
                                                0.025, na.rm=T))

upper.bound.presymp.shed.mod.sars <- c(quantile(presymp.shed.sample.slopes.sars, 
                                                0.975, na.rm=T),
                                       quantile(presymp.shed.sample.intercepts.sars, 
                                                0.975, na.rm=T))

```

We plot the delay between symptom onset and peak viral shedding for each virus,
delay vs. viral replication rate, day of peak shedding vs. viral replication rate,
day of symptom onset vs. viral replication rate, and average viral shedding 
prior to symptoms vs. symptom onset in a 5-panel plot.
```{r, fig.height = 7.5, fig.width = 3.5}

plot.fig3 <- function(params.delays.data, 
                      b.mod.noro, 
                      b.mod.lower.noro,
                      b.mod.upper.noro,
                      b.mod.sars,
                      b.mod.lower.sars,
                      b.mod.upper.sars,
                      c.mod.noro=NA,
                      c.mod.lower.noro=NA,
                      c.mod.upper.noro=NA,
                      c.mod.sars=NA,
                      c.mod.lower.sars=NA,
                      c.mod.upper.sars=NA,
                      params.peaktime.data=NA,
                      d.mod.noro=NA, d.mod.sars=NA,
                      d.mod.lower.noro=NA,
                      d.mod.upper.noro=NA,
                      d.mod.lower.sars=NA,
                      d.mod.upper.sars=NA,
                      params.symp.data=NA,
                      transmission.latent.incubation.data=NA, 
                      e.mod.noro=NA, e.mod.sars=NA,
                      e.mod.lower.noro=NA,
                      e.mod.upper.noro=NA,
                      e.mod.lower.sars=NA,
                      e.mod.upper.sars=NA,
                      cex, cex_labels, 
                      panellabelratio, nonconservative = FALSE) {

  par(mfrow=c(5, 1), oma=c(0.5, 2, 0.5, 0), mai=c(0.3, 0.5, 0.1, 0.1))
  
  # Panel A: Delay between symptom onset and peak viral shedding 
  ylim.ab <- c(floor(min(params.delays.data$delay, na.rm=T))-1, 
               ceiling(max(params.delays.data$delay, na.rm=T))+1)
  
  plot(delay ~ jitter(ifelse(pathogen == 'Norovirus', 1, 2)), 
       data=params.delays.data,
       col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       xlim=c(.5, 2.5),
       ylim=ylim.ab, cex.axis=cex, cex=cex,
       pch=ifelse(delay < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(delay > 0 & pathogen == 'Norovirus', 2,
                         ifelse(delay < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(delay > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
       bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       yaxt="n",
       xaxt="n",
       ylab = "",
       xlab = "",
       mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  axis(1, at=seq(1,2), labels=unique(params.delays.data$pathogen), cex.axis=cex, 
       mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
       las=2)
  
  #Average delays per pathogen
  for (i in 1:length(unique(params.delays.data$pathogen))) {
    segments(-0.3 + i, 
             mean(params.delays.data[params.delays.data$pathogen == 
                           unique(params.delays.data$pathogen)[i],]$delay,
                  na.rm=T),
             -0.3 + i + 0.6, col=ifelse(unique(
               params.delays.data$pathogen)[i] == 'Norovirus', 
                                        noro.color, sars.color), 
             lwd=cex)
  }

  #Pre-symptomatic vs post-symptomatic line
  abline(h=0, lty=3, lwd=cex)
  
  text(panellabelratio * 2 + .5, panellabelratio * (ylim.ab[2] - ylim.ab[1]) + 
         ylim.ab[1], labels="A", cex=cex, font=2)
  
  # Panel B: Delay between symptom onset and peak viral shedding vs. viral
  # growth rate
  
  xlim.bcd <- c(floor(min(params.delays.data$p2)) - 1,
                ceiling(max(params.delays.data$p2)) + 1)
    
  plot(delay ~ p2, data=params.delays.data, 
     col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
     pch=ifelse(delay < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(delay > 0 & pathogen == 'Norovirus', 2,
                         ifelse(delay < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(delay > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
     bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     cex=cex, cex.axis=cex,
     ylab="", xlab="", xaxt="n", yaxt="n",
     ylim=ylim.ab,
     xlim=xlim.bcd,
     mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  
  bcd.xlabels <- lapply(seq(xlim.bcd[1], xlim.bcd[2], by=3), function(x) 
         bquote(10^.(x)))
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], by=3),
       labels=do.call(expression, bcd.xlabels),
       cex.axis=cex, 
       mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
       las=2)
  
  noro.r.range <- seq(min(
    params.delays.data[params.delays.data$pathogen == 'Norovirus',]$p2),
    max(params.delays.data[params.delays.data$pathogen == 'Norovirus',]$p2),
    length=10)
  
  sars.r.range <- seq(min(
    params.delays.data[params.delays.data$pathogen == 'SARS-CoV-2',]$p2),
    max(params.delays.data[params.delays.data$pathogen == 'SARS-CoV-2',]$p2),
    length=10)
  
  # Lines of best fit
  noro.bpred <- data.frame(p2 = noro.r.range)
  noro.bpred <- transform(noro.bpred, y_predicted = predict(b.mod.noro, newdata=noro.bpred))
  lines(y_predicted ~ p2, data=noro.bpred, col=noro.color)
  
  sars.bpred <- data.frame(p2 = sars.r.range)
  sars.bpred <- transform(sars.bpred, y_predicted = predict(b.mod.sars, newdata=sars.bpred))
  lines(y_predicted ~ p2, data=sars.bpred, col=sars.color)
  
  # Confidence bounds 
  noro.x <- noro.r.range
  noro.y.lower <- b.mod.lower.noro[1] * noro.x + b.mod.lower.noro[2]
  noro.y.higher <- b.mod.upper.noro[1] * noro.x + b.mod.upper.noro[2]
  
  polygon(c(rev(noro.x), noro.x), c(rev(noro.y.higher), noro.y.lower), 
          col = rgb(135, 26, 110, alpha=100,
                  maxColorValue=255), border = NA)
  
  sars.x <- sars.r.range
  sars.y.lower <- b.mod.lower.sars[1] * sars.x + b.mod.lower.sars[2]
  sars.y.higher <- b.mod.upper.sars[1] * sars.x + b.mod.upper.sars[2]
  
  
  polygon(c(rev(sars.x), sars.x), c(rev(sars.y.higher), sars.y.lower), 
          col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
  
  abline(h=0, lty=3, lwd=cex)
  
  if (nonconservative) {
    mtext(text=expression(Viral~replication~rate~(hour^-1)),
          side=1, line=2.5, cex=cex_labels)
    mtext(text="Day of shedding onset \n minus day of symptom onset",
          side=2, line=2.1, cex=cex_labels)
  } else {
    mtext(text="Day of peak viral shedding \n minus day of symptom onset",
          side=2, line=2.1, cex=cex_labels, adj=-0.5)
  }
  
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1])  + xlim.bcd[1], 
       panellabelratio * (ylim.ab[2] - ylim.ab[1]) + ylim.ab[1], labels="B", 
       cex=cex, font=2)
  
  if (!nonconservative) {
    ylim.c <- c(floor(min(params.peaktime.data$peaktime/24)) - 1, 
                ceiling(max(params.peaktime.data$peaktime/24)) + 1)
    # Panel C: Time of peak shedding vs. viral replication rate
    plot(peaktime/24 ~ p2, data=params.peaktime.data,
         col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
         xlim=xlim.bcd,
         ylim=ylim.c, cex.axis=cex, cex=cex,
         pch=ifelse(delay < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(delay > 0 & pathogen == 'Norovirus', 2,
                         ifelse(delay < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(delay > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
         bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
         xaxt="n", yaxt="n", ylab = "", xlab = "",
         mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
    
    axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], by=2), 
       labels=do.call(expression, bcd.xlabels),
       cex.axis=cex, 
       mgp=c(0, 0.5, 0), tck=-0.04)
    axis(2, at=seq(ylim.c[1],ylim.c[2], by=2), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
    
    # Lines of best fit
    noro.cpred <- data.frame(p2 = noro.r.range)
    noro.cpred <- transform(noro.cpred, y_predicted = predict(c.mod.noro, newdata=noro.cpred))
    lines(y_predicted ~ p2, data=noro.cpred, col=noro.color)
    
    sars.cpred <- data.frame(p2 = sars.r.range)
    sars.cpred <- transform(sars.cpred, y_predicted = predict(c.mod.sars, newdata=sars.cpred))
    lines(y_predicted ~ p2, data=sars.cpred, col=sars.color)
    
    # Confidence bounds 
    noro.y.lower <- (c.mod.lower.noro[1]/24) * noro.x +
      (c.mod.lower.noro[2]/24)
    noro.y.higher <- (c.mod.upper.noro[1]/24) * noro.x +
      (c.mod.upper.noro[2]/24)
    
    polygon(c(rev(noro.x), noro.x), c(rev(noro.y.higher), noro.y.lower), 
          col=rgb(135, 26, 110, alpha=100,
                  maxColorValue=255), border = NA)

    
    sars.y.lower <- (c.mod.lower.sars[1]/24) * sars.x +
      (c.mod.lower.sars[2]/24)
    sars.y.higher <- (c.mod.upper.sars[1]/24) * sars.x +
      (c.mod.upper.sars[2]/24)
    
    polygon(c(rev(sars.x), sars.x), c(rev(sars.y.higher), sars.y.lower), 
            col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
    
    mtext(text="Day of \npeak shedding", side=2, line=2, cex=cex_labels)
    
    text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
         panellabelratio * (ylim.c[2] - ylim.c[1]) + ylim.c[1], labels="C", 
         cex=cex, font=2)
      
    # Panel D: Symptom onset vs. viral replication rate
    ylim.d <- c(floor(min(params.symp.data$incubation, na.rm=T)) - 1, 
                ceiling(max(params.symp.data$incubation, na.rm=T)) + 1)
    
    plot(incubation ~ p2, data=params.symp.data,
         col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
         xlim=xlim.bcd, ylim=ylim.d, cex.axis=cex, cex=cex,
         pch=ifelse(delay < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(delay > 0 & pathogen == 'Norovirus', 2,
                         ifelse(delay < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(delay > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
         bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
         yaxt="n", xaxt="n", ylab = "", xlab = "", bty='l')
    
    axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], by=2), 
       labels=do.call(expression, bcd.xlabels), 
       cex.axis=cex, 
       mgp=c(0, 0.5, 0), tck=-0.04)
    
    axis(2, at=seq(ylim.d[1], ylim.d[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
    
      # Lines of best fit
    noro.dpred <- data.frame(p2 = noro.r.range)
    noro.dpred <- transform(noro.dpred, y_predicted = predict(d.mod.noro, newdata=noro.dpred))
    lines(y_predicted ~ p2, data=noro.dpred, col=noro.color)
    
    sars.dpred <- data.frame(p2 = sars.r.range)
    sars.dpred <- transform(sars.dpred, y_predicted = predict(d.mod.sars, newdata=sars.dpred))
    lines(y_predicted ~ p2, data=sars.dpred, col=sars.color)
    
    # Confidence bounds 
    noro.y.lower <- d.mod.lower.noro[1] * noro.x + d.mod.lower.noro[2]
    noro.y.higher <- d.mod.upper.noro[1] * noro.x + d.mod.upper.noro[2]
    
    polygon(c(rev(noro.x), noro.x), c(rev(noro.y.higher), noro.y.lower), 
            col = rgb(135, 26, 110, alpha=100,
                    maxColorValue=255), border = NA)
    
    sars.y.lower <- d.mod.lower.sars[1] * sars.x + d.mod.lower.sars[2]
    sars.y.higher <- d.mod.upper.sars[1] * sars.x + d.mod.upper.sars[2]
    
    polygon(c(rev(sars.x), sars.x), c(rev(sars.y.higher), sars.y.lower), 
            col=rgb(86, 179, 233, alpha=100,
                    maxColorValue=255), border = NA)
    
    mtext(text="Day of \nsymptom onset", side=2, line=2, cex=cex_labels)
    mtext(text=expression(Viral~replication~rate~(hour^-1)),
          side=1, line=2, cex=cex_labels)
    text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
         panellabelratio * (ylim.d[2] - ylim.d[1]) + ylim.d[1], 
         labels="D", cex=cex, font=2)
  
    # Panel E: Pre-symptomatic transmission potential vs. symptom onset
    xlim.e <- c(floor(min(transmission.latent.incubation.data$symptom_onset,
                    na.rm=T)) - 1,
                ceiling(max(transmission.latent.incubation.data$symptom_onset,
                    na.rm=T)) + 1)
    
    ylim.e <- c(floor(min(transmission.latent.incubation.data$transmission_potential,
                    na.rm=T)) - 1,
                ceiling(max(transmission.latent.incubation.data$transmission_potential,
                    na.rm=T)) + 1)
    
    plot(transmission_potential ~ jitter(symptom_onset), 
         data=merge(params.delays.data, transmission.latent.incubation.data, by='ID',
                    all.y=T),
         col=ifelse(pathogen.y == 'Norovirus', noro.color, sars.color), 
         pch=ifelse(is.na(ifelse(delay < 0 & pathogen.y == 'Norovirus', 17, 
                ifelse(delay > 0 & pathogen.y == 'Norovirus', 2,
                       ifelse(delay < 0 & pathogen.y == 'SARS-CoV-2',
                              23, 5)))) | delay == 0,
                9, 
                ifelse(delay < 0 & pathogen.y == 'Norovirus', 17, 
                ifelse(delay > 0 & pathogen.y == 'Norovirus', 2,
                       ifelse(delay < 0 & pathogen.y == 'SARS-CoV-2',
                              23, 5)))),
         bg=ifelse(pathogen.y == 'Norovirus', noro.color, sars.color), 
         ylim=ylim.e,
         xlim=xlim.e,
         cex.axis=cex, cex=cex,
         ylab = "",
         xlab = "",
         xaxt="n",
         yaxt="n", bty='l')
    
    e.ylabels <- lapply(seq(ylim.e[1], ylim.e[2], by=2.5), function(x) 
         bquote(10^.(x)))
    axis(1, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04)
    axis(2, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, at=seq(ylim.e[1], ylim.e[2], by=2.5),
         labels=do.call(expression, e.ylabels),
         las=2)
    
    legend(.375 * (xlim.e[2] - xlim.e[1]) + xlim.e[1], 
           .9 * (ylim.e[2] - ylim.e[1]) + ylim.e[1], 
           legend=c("Pre-symptomatic", "Post-symptomatic", 
                    "Unclear peak (this panel only)"), 
       pch=c(19, 1, 9), col=c("black", "black", "black"), cex=cex_labels, bty="n")
  
    mtext(text="Average viral shedding \n prior to symptom onset",
          side=2, line=3, cex=cex_labels)
    mtext(text="Symptom onset (day)", side=1, line=1.5, cex=cex_labels)
    
      # Lines of best fit
    noro.symponset.range <- seq(min(transmission.latent.incubation.data[
                                 transmission.latent.incubation.data$pathogen == 'Norovirus',]$symptom_onset, na.rm=T), 
                                  max(transmission.latent.incubation.data[
                                    transmission.latent.incubation.data$pathogen == 'Norovirus',]$symptom_onset, na.rm=T), length=10)
    noro.epred <- data.frame(symptom_onset= noro.symponset.range)
    noro.epred <- transform(noro.epred, y_predicted = predict(e.mod.noro, newdata=noro.epred))
    lines(y_predicted ~ symptom_onset, data=noro.epred, col=noro.color)
    
    sars.symponset.range <- seq(min(transmission.latent.incubation.data[
                                 transmission.latent.incubation.data$pathogen == 'SARS-CoV-2',]$symptom_onset, na.rm=T), 
                                  max(transmission.latent.incubation.data[
                                    transmission.latent.incubation.data$pathogen == 'SARS-CoV-2',]$symptom_onset, na.rm=T), length=10)
    sars.epred <- data.frame(symptom_onset = sars.symponset.range)
    sars.epred <- transform(sars.epred, y_predicted = predict(e.mod.sars, newdata=sars.epred))
    lines(y_predicted ~ symptom_onset, data=sars.epred, col=sars.color)
    
    # Confidence bounds 
    noro.x <- noro.symponset.range
    noro.y.lower <- e.mod.lower.noro[1] * noro.x + 
      e.mod.lower.noro[2]
    noro.y.higher <- e.mod.upper.noro[1] * noro.x + 
      e.mod.upper.noro[2]
    
    polygon(c(rev(noro.x), noro.x), c(rev(noro.y.higher), noro.y.lower), 
            col = rgb(135, 26, 110, alpha=100,
                    maxColorValue=255), border = NA)
    
    sars.x <- sars.symponset.range
    sars.y.lower <- e.mod.lower.sars[1] * sars.x + 
      e.mod.lower.sars[2]
    sars.y.higher <- e.mod.upper.sars[1] * sars.x + 
      e.mod.upper.sars[2]
    
    polygon(c(rev(sars.x), sars.x), c(rev(sars.y.higher), sars.y.lower), 
            col=rgb(86, 179, 233, alpha=100,
                    maxColorValue=255), border = NA)
    
    text(panellabelratio * (xlim.e[2] - xlim.e[1]) + xlim.e[1], 
         panellabelratio * (ylim.e[2] - ylim.e[1]) + ylim.e[1], 
         labels="E", cex=cex, font=2)
  }

}

params.and.delays.conservative.combined <- rbind(params.and.delays.conservative.noro, 
                                    params.and.delays.conservative.sars)
params.and.delays.conservative.combined$pathogen <- ifelse(grepl('ID', 
                                                      params.and.delays.conservative.combined$ID), 
                                                'Norovirus', 'SARS-CoV-2')

transmission.latent.combined <- rbind(transmission.latent.noro, transmission.latent.sars)
transmission.latent.combined$pathogen <- ifelse(grepl('ID', 
                                                      transmission.latent.combined$ID), 
                                                'Norovirus', 'SARS-CoV-2')

combined.params.peaktime.data <- rbind(params.peaktime.data.noro, params.peaktime.data.sars)
combined.params.peaktime.data$pathogen <- ifelse(grepl('ID', 
                                                      combined.params.peaktime.data$ID), 
                                                'Norovirus', 'SARS-CoV-2')

combined.params.sympdata <- rbind(params.symp.data.noro, params.symp.data.sars)
combined.params.sympdata$pathogen <- ifelse(grepl('ID', 
                                                      combined.params.sympdata$ID), 
                                                'Norovirus', 'SARS-CoV-2')

tiff("Pre-symptomatic transmission analysis.tiff", res=500, height=6.5, width=3.5, unit="in")

plot.fig3(
  params.delays.data = params.and.delays.conservative.combined,
  b.mod.noro = lm.fit.p2.delay.conservative.noro,
  b.mod.lower.noro = lower.bound.delay.mod.noro,
  b.mod.upper.noro = upper.bound.delay.mod.noro,
  b.mod.sars = lm.fit.p2.delay.conservative.sars,
  b.mod.lower.sars = lower.bound.delay.mod.sars,
  b.mod.upper.sars = upper.bound.delay.mod.sars,
  c.mod.noro = lm.fit.p2.peaktime.noro,
  c.mod.lower.noro = lower.bound.peakshedding.mod.noro,
  c.mod.upper.noro = upper.bound.peakshedding.mod.noro,
  c.mod.sars = lm.fit.p2.peaktime.sars,
  c.mod.lower.sars = lower.bound.peakshedding.mod.sars,
  c.mod.upper.sars = upper.bound.peakshedding.mod.sars,
  params.peaktime.data = combined.params.peaktime.data,
  d.mod.noro = lm.fit.p2.symponset.noro,
  d.mod.sars = lm.fit.p2.symponset.sars,
  d.mod.lower.noro = lower.bound.symponset.mod.noro,
  d.mod.upper.noro = upper.bound.symponset.mod.noro,
  d.mod.lower.sars = lower.bound.symponset.mod.sars,
  d.mod.upper.sars = upper.bound.symponset.mod.sars,
  params.symp.data = combined.params.sympdata,
  transmission.latent.incubation.data = transmission.latent.combined,
  e.mod.noro = transmission.latent.fit.noro,
  e.mod.sars = transmission.latent.fit.sars,
  e.mod.lower.noro = lower.bound.presymp.shed.mod.noro,
  e.mod.upper.noro = upper.bound.presymp.shed.mod.noro,
  e.mod.lower.sars = lower.bound.presymp.shed.mod.sars,
  e.mod.upper.sars = upper.bound.presymp.shed.mod.sars,
  cex=1.1, cex_labels=0.75, panellabelratio=0.975)
dev.off()
```

We again plot delay vs. virus and delay vs. viral replication rate using
the non-conservative measure of pre-symptomatic transmission (shedding onset
preceding symptom onset).
```{r, fig.height = 6, fig.width = 3.5}

params.and.delays.nonconservative.combined <- rbind(params.and.delays.nonconservative.noro, 
                                    params.and.delays.nonconservative.sars)
params.and.delays.nonconservative.combined$pathogen <- ifelse(grepl('ID', 
                                                      params.and.delays.nonconservative.combined$ID), 
                                                'Norovirus', 'SARS-CoV-2')

tiff("Non-conservative pre-symptomatic transmission analysis.tiff", res=500, height=6.5, width=3.5, unit="in")

plot.fig3(
  params.delays.data = params.and.delays.nonconservative.combined,
  b.mod.noro = lm.fit.p2.delay.nonconservative.noro, 
  b.mod.sars = lm.fit.p2.delay.nonconservative.sars,
  b.mod.lower.noro = lower.bound.delay.nonconservative.mod.noro,
  b.mod.upper.noro = upper.bound.delay.nonconservative.mod.noro,
  b.mod.lower.sars = lower.bound.delay.nonconservative.mod.sars,
  b.mod.upper.sars = upper.bound.delay.nonconservative.mod.sars,
  cex=1.1, cex_labels=0.75, panellabelratio=0.975,
  nonconservative = TRUE)

dev.off()

```

We plot the differences in incubation periods reported in Ge *et al.*, 2023,
and Atmar *et al.*, 2008.
```{r}
atmarIDs <- read.csv("AtmarIDs.csv")

# Using peak viral load, inoculum dose, and type of symptoms to match up participant IDs

geSymptomData <- noroData$mvs
geSymptomData$type <- ""
geSymptomData$type[geSymptomData$mvsc3>0] <- "v"
# mvsc3 indicates duration of vomiting
# mvsc1 indicates duration of diarrhea
geSymptomData$type[geSymptomData$mvsc3>0 & geSymptomData$mvsc1>0] <- "dv"

# Participant ID4 had reported viral shedding in vomit; we and the authors
# of the study do not know why mvsc3 = 0
geSymptomData[geSymptomData$ID == 'ID4',]$type <- "dv"

geIDs <- do.call(rbind, 
                  lapply(noroParticipants,
                         function(x) data.frame(
                           geID=x,
                           dose=unique(noro.FShedData[noro.FShedData$ID == x,]$dose),
                           peak=round(log10(max(noro.FShedData[noro.FShedData$ID == x,]$y_virus, na.rm=T)), 1),
                           symptom.type=geSymptomData[geSymptomData$ID==x,]$type)))

ID_dict_noro <- merge(atmarIDs, geIDs, by=c("dose", "peak", "symptom.type"))

combinedNoroData <- merge(merge(ID_dict_noro, noro.IncubData, by.x = 'geID', 
                              by.y='ID')[, c(1, 5, 8, 6)], participant.peaktimes.noro,
                          by.x="geID", by.y="ID")
colnames(combinedNoroData) <- c('geID', 'atmarID', 'geIncubation', 'atmarIncubation',
                                'peaktime')

# Manually adding in ID10/720 because their trajectories are identical, but
# their symptoms do not match up between the two datasets
combinedNoroData <- rbind(combinedNoroData, 
                          data.frame(geID = 'ID10', atmarID='720', 
                                     geIncubation = noro.IncubData[
                                       noro.IncubData$ID == 'ID10',]$incubation,
                                     atmarIncubation = atmarIDs[
                                       (atmarIDs$atmarID == '720') & 
                                         (!is.na(atmarIDs$atmarID)),]$incubation,
                                     peaktime = participant.peaktimes.noro[
                                       participant.peaktimes.noro$ID == 'ID10',]$peaktime))

# Manual ordering of IDs
combinedNoroData <- combinedNoroData[c(1, 7, 8, 9, 10, 11, 12, 13, 2, 3, 4, 5, 6),]



atmarColor <- "#f1a340"
geColor <- "#871a6e"
tiff(filename="Symptom Onset Discrepancy.tiff", height=3.5, width=5, units = "in",
    res=500)
{par(mar=c(4.1, 3.6, 0.5, 0.1), oma=c(2, 0, 0, 0))
plot(1:nrow(combinedNoroData), combinedNoroData$atmarIncubation, xaxt="n",
     ylim=c(0, 5), col=atmarColor, xlab="", ylab="",
     bty="n")
axis(1, at=1:nrow(combinedNoroData), 
     labels=combinedNoroData$geID,las=2, col.axis=geColor)
axis(1, at=1:(nrow(combinedNoroData)), 
     labels=paste(combinedNoroData$atmarID, ' /        ', sep=''), las=2, 
     col.axis=atmarColor)
mtext("Participant ID", side=1, line=5)
mtext("Day", side=2, line=2.5)
legend(4, 5, pch=c(1, 1, 16), col=c(atmarColor, geColor, 'black'), 
       legend=c(bquote("Atmar" ~ italic("et al.") ~ 2008 ~ "\u2014 symptom onset"), 
                bquote("Ge" ~ italic("et al.") ~ 2023 ~ "\u2014 symptom onset"),
                "Peak viral shedding"))
points(1:nrow(combinedNoroData), combinedNoroData$geIncubation, col=geColor)
points(1:nrow(combinedNoroData), combinedNoroData$peaktime/24, pch=16)
dev.off()}
```
