---
title: "Analyzing human infection data (norovirus and SARS-CoV-2)"
author: "Kayla Zhang, Damie Pak, Megan Greischar"
date: "2025-05-24"
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

Packages we use
```{r}
library(here)
library(pzfx)
library(data.table)
library(spatialEco)
library(pracma)
library(dplyr)
library(Polychrome)
```

### Importing/cleaning data

We use the shedding and symptom timing data that Ge *et al*., 2023 obtained 
from a norovirus study (Atmar *et al*., 2008). We exclude participants with 
less than 5 quantitative viral shedding points to avoid overfitting the 
statistical model. This data also contains qualitative test results for 
viral shedding values (positive or negative) falling below the qRT-PCR detection 
threshold. Any shedding values below the detection threshold for quantitative
results are set to NA.
```{r, echo = TRUE, eval = FALSE}
exclude.participants <- function(p, data) {
  
  pdata <- data[data$ID == p, ]
  
  if ('y_type' %in% colnames(pdata)) {
    negative_readings <- which(grepl('NEG', pdata$y_type))
  } else {
    negative_readings <- which(is.na(pdata$y_virus))
  }
  
  if (max(negative_readings) < 7) {
    clearance <- nrow(pdata)
  } else {
    clearance <- min(negative_readings[negative_readings > 7]) - 1
  }
  
  pdata <- pdata[1:clearance,]
    
  if(nrow(pdata[!is.na(pdata$y_virus),]) < 5) {
    return(p)
  }

}

noroData <- readRDS(here("data", "norodata.rds")) 

write.table(noroData$ftsshedding, "noroFtsData.txt")
noro.FShedData.unaggregated <- read.table("noroFtsData.txt", stringsAsFactors = F)
noro.FShedData.unaggregated <- noro.FShedData.unaggregated[, c(1, 2, 5, 6, 8)]

write.table(noroData$incubation, "noroIncubationData.txt")
noro.IncubData <- read.table("noroIncubationData.txt", stringsAsFactors = F)
noro.IncubData <- noro.IncubData[, c(1, 2, 5)]

# Get rid of viral shedding numbers below the qRT-PCR detection threshold 
# (4e7 copies /ml)
noro.detection.threshold <- 15000
noro.detection.threshold.2 <- 4e7
noro.FShedData.unaggregated[noro.FShedData.unaggregated$y_virus <= 
                              noro.detection.threshold.2,]$y_virus <- NA

# Exclude participants with less than 5 points of data 
noro.excluded <- unlist(lapply(unique(noro.FShedData.unaggregated$ID), exclude.participants, 
                               data=noro.FShedData.unaggregated))

noro.excluded <- noro.excluded[!is.null(noro.excluded)]

noro.FShedData.unaggregated <- noro.FShedData.unaggregated[
  !(noro.FShedData.unaggregated$ID %in% noro.excluded),]
noro.IncubData <- noro.IncubData[!(noro.IncubData$ID %in% noro.excluded),]

# Average across viral shedding values at the same time point
aggregated_virus <- aggregate(y_virus ~ ID + days, 
                              noro.FShedData.unaggregated,
                              mean,
                              na.action=na.pass)
aggregated_virus <- aggregated_virus[order(as.numeric(substring(aggregated_virus$ID, 3, 4)),
                       aggregated_virus$days),]$y_virus

noro.FShedData <- noro.FShedData.unaggregated[
  !duplicated(noro.FShedData.unaggregated[, c(1, 3)]),]
noro.FShedData$y_virus <- aggregated_virus
```

We extract the shedding and symptom timing data from Zhou *et al*., 2023, 
a SARS-CoV-2 study, as well.
```{r}
extract.incubation <- function(symp.data) {
  
  symptom.indexes <- symp.data[which(symp.data[, 2] > 0),]$ROWTITLE
  
  if (length(symptom.indexes) == 0) {
    symp.onset <- NA
  } else {
    symp.onset <- min(symptom.indexes) 
  }
  
  incub.data <- data.frame("ID" = colnames(symp.data)[2], "incubation" = symp.onset)
  
  return(incub.data)
}

sars.TShedData <- read_pzfx(here("data", "SARS-Supp-2.pzfx"), table=37)[, c(1, 20:37)]
sars.TShedData <- melt(as.data.table(sars.TShedData), id.vars=c("Day"), 
                       variable.name = "ID", value.name = "y_virus", 
                       variable.factor = F)
colnames(sars.TShedData)[1] <- "days"

incub.indexes <- 37:54
sars.IncubData <- data.frame(do.call(rbind, 
                                     lapply(incub.indexes, 
                                            function(x) 
                                              extract.incubation(
                                                read_pzfx(here("data", "SARS-Fig-1.pzfx"), 
                                                    table=x)))))

# Participants were matched up by visually comparing viral shedding values

ID_dict_sars <- c("680539" = "Throat_10",
"673353" = "Throat_18",
"667186" = "Throat_11",
"550630" = "Throat_7",
"635495" = "Throat_1",
"666482" = "Throat_9",
"651806" = "Throat_15",
"635331" = "Throat_5",
"634105" = "Throat_2",
"635779" = "Throat_16",
"635729" = "Throat_4",
"637340" = "Throat_12",
"627506" = "Throat_6",
"666660" = "Throat_8",
"676586" = "Throat_13",
"666427" = "Throat_14",
"647785" = "Throat_3",
"674700" = "Throat_17"
)

sars.IncubData$ID <- ID_dict_sars[sars.IncubData$ID]

# Get rid of viral shedding below the detection threshold (1200 copies/ml)
sars.detection.threshold <- 1200
sars.TShedData[sars.TShedData$y_virus < sars.detection.threshold,]$y_virus <- NA

# Exclude participants with less than 5 points of data 
sars.excluded <- unlist(lapply(unique(sars.TShedData$ID), exclude.participants, data=sars.TShedData))

sars.excluded <- sars.excluded[!is.null(sars.excluded)]

sars.TShedData <- sars.TShedData[!(sars.TShedData$ID %in% sars.excluded),]
sars.IncubData <- sars.IncubData[!(sars.IncubData$ID %in% sars.excluded),]
```

We calculate each study's average shedding and symptom onset.
```{r}
shed.onset <- function(vldata) {
  if("y_type" %in% colnames(vldata)) {
    return(min(vldata[grepl('POS', vldata$y_type),]$days))
  } else {
    return(min(vldata[!is.na(vldata$y_virus),]$days))
  }
}

noroParticipants <- unique(noro.FShedData$ID)

noro.avg.shed.onset <- mean(unlist(lapply(noroParticipants, function(x) 
  shed.onset(noro.FShedData[noro.FShedData$ID == x, ]))))
noro.avg.shed.onset

noro.avg.symp.onset <- mean(noro.IncubData$incubation)
noro.avg.symp.onset

sarsParticipants <- unique(sars.TShedData$ID)

sars.avg.shed.onset <- mean(unlist(lapply(sarsParticipants, function(x) 
  shed.onset(sars.TShedData[sars.TShedData$ID == x, ]))))
sars.avg.shed.onset

sars.avg.symp.onset <- mean(sars.IncubData$incubation, na.rm=T)
sars.avg.symp.onset
```

### Fitting statistical model to observed shedding values

We fit a statistical model (Li & Handel, 2014, Holder & Beauchemin, 2011) to each
individual's shedding data that estimates viral load over time using the 
least squares method. There is a typo in Li & Handel, 2014 in the denominator, 
where the term exp(p4 * (t-p3)) is written as exp(-p4 * (t-p3)) (Holder & 
Beauchemin, 2011). We make the assumption that viral load is a 1-to-1 mapping to 
viral shedding. 

When model estimates incorrectly lie outside a given detection range,
we calculate the SSE of that point using the nearest detection threshold as 
the "observed" data point.
```{r}
generate.estimates <- function(t, p1, p2, p3, p4) {

  estimate <- (2 * p1) / (exp(-p2*(t-p3)) + exp(p4*(t-p3)))

  # set zeros to very small number to avoid log10(0)
  estimate[estimate==0] <- 1e-16

  res <- log10(estimate)
  
  return(res)

}

sse.viralLoad <- function(pars, data, detect.threshold, 
                          detect.threshold.2 = NA, optimization = T) {
  vl <- log10(data$y_virus)
  t <- data$days * 24
  
  finite.range <- 24*range(data$days[is.finite(data$y_virus)],na.rm=T)
  
  p1 <- exp(pars[1])
  p2 <- exp(pars[2])
  p3 <- exp(-exp(pars[3]))*diff(finite.range) + finite.range[1]
  p4 <- exp(pars[4])
  
  if(is.na(optimization) | optimization==F){
    # return hi resolution predicted time series
      hiResT <- seq(min(data$days[data$days > 0]) * 24, 
                   max(data$days) * 24, by=1)
      hiResEst <- generate.estimates(hiResT, p1, p2, p3, p4)
      if(is.na(optimization)){
        # return peak height, replication rate, time of peak, decay rate
        res <- c(max(hiResEst),p2,hiResT[which.max(hiResEst)],p4)
        }
      if(!is.na(optimization) & optimization==F){
        res <- data.frame(t = hiResT, predData = hiResEst)}
        }

  if(!is.na(optimization) & optimization == T){
    estimates <- generate.estimates(t, p1, p2, p3, p4)
    
    # Punishing model estimates that incorrectly lie above the detection
    # threshold or extremely low (< 1)
    vl[which(is.na(vl) & estimates > detect.threshold)] <- detect.threshold
    vl[which(is.na(vl) & estimates < 1)] <- 1
    
    # Punishing model estimates that incorrectly lie outside the detection range of 
    # qualitative results
    vl[which(grepl('~', data$y_type) & estimates < detect.threshold)] <- detect.threshold 
    vl[which(grepl('~', data$y_type) & estimates > detect.threshold.2)] <- detect.threshold.2
    
    # Keep NA if estimates lie between two thresholds
    vl[which(grepl('~', data$y_type) & estimates < detect.threshold.2 & estimates > detect.threshold)] <- NA
    
    sse <- sum((estimates - vl)^2, na.rm=TRUE)
    
    res <- sse
    } # end conditional: optimize?
  
  return(res)
  } # end sse.viralLoad function

generate.possible.params <- function(data, detect.threshold,
                                     detect.threshold.2 = NA) {
  
  convergenceValue <- 1
  
  # These are ln(param) values
  while(convergenceValue>0){
    p1 <- runif(1, log(1e9), 
                   log(1e13))
    p2 <- runif(1, log(0.1), log(1)) 
    p3 <- runif(1, -10, 10)
    p4 <- runif(1, log(10^(-2)), log(0.1))
    
    params <- c(p1, p2, p3, p4)
    
    fit <- optim(params, sse.viralLoad, data=data, detect.threshold=detect.threshold,
                 detect.threshold.2=detect.threshold.2)
    convergenceValue <- fit$convergence
    
    } # end while loop
  
  transformed.params <- sse.viralLoad(fit$par, data = data, optimization = NA)
  param.guess <- data.frame(unique(data$ID), 
                         fit$par[1],
                         fit$par[2],
                         fit$par[3],
                         fit$par[4],
                         fit$convergence, 
                         fit$value,
                         transformed.params[1],
                         transformed.params[2],
                         transformed.params[3],
                         transformed.params[4])
  
  return(param.guess)
  
}

generate.fits <- function(data, numguess, detect.threshold, detect.threshold.2 = NA) {

  fits <- do.call(rbind, 
                  lapply(1:numguess,
                         function(x) generate.possible.params(data, detect.threshold,
                                                              detect.threshold.2))) 
  
  colnames(fits) <- c('ID', 'p1','p2','p3','p4','convergence','sse','transp1','transp2','transp3','transp4')
  return(fits)
}

minsse <- function(fits) {
  
  minsse <- min(fits$sse)
  pars <- unlist((fits[fits$sse == minsse, ])[1,])
  return(pars)
}

all.fits <- function(data, endpoints, numguess, detect.threshold, detect.threshold.2 = NA) {
  
  participants <- unique(data$ID)

  participantfits <- do.call(rbind,
                             lapply(participants,
                                    function(x) 
                                      generate.fits(data[data$ID==x,]
                                                    [1:(endpoints[endpoints$ID==x,]$endPt),],
                                                    numguess, detect.threshold,
                                                    detect.threshold.2)))
  
  return(participantfits)
}

participant.fits.minsse <- function(fits) {
    participants <- unique(fits$ID)
  
    participantparams <- data.frame(do.call(rbind, 
                               lapply(participants, 
                                      function(x) minsse(fits[fits$ID == x,]))))
  
    participantparams[, 2:length(participantparams[1,])] <- apply(participantparams[, 2:length(participantparams[1,])], 2, as.numeric)
  
    return(participantparams)
}
```

For endpoints, we looked for the first local minimum in the data. If such 
a value does not exist, we look for any value lesser than all subsequent shedding
values. If such a value does not exist, we look for the first value below
the detection threshold after the maximum detected viral shedding value. Lastly,
if this value does not exist, we define the endpoint as the length of the 
shedding time series.
```{r, results='hide'}
whichLower <- function(index, vec){
  # function indicates whether index entry of vec
    # is less than all subsequent values in vec
  return(all(vec[(index+1):length(vec)] > vec[index]))
  }

end.point <- function(data) {
  # endPt is defined as the length of the full viral abundance time series
    # which is replaced by the following in order (if defined):
    # 1. the first value below the detection threshold after the max. detected abundance
       # if such a value exists
    # 2. any value in prior to the endPt defined by (1) for which all subsequent values were greater
    # 3. as the local minimum (defined as a timepoint with a negative derivative two points prior,
       # and a positive derivative two points following) if such a value occurs earlier than the
       # endPt defined in (2)
  endPt <- length(data$y_virus)
  
  # 1. locate first value below detection threshold, if it exists
      # consider only values for which data stay below detection
      # threshold for at least one day
  allMissingValues <- which(is.na(data$y_virus)) # below detection limit
  # discard missing values if the gap is less than one day
  gaps <- c(diff(data$days),NA)
  missingValues <- allMissingValues[which(gaps[allMissingValues]>=1)]

  # find first missing value after max. viral shedding
  maxViralAbundance <- which.max(data$y_virus)

  if(length(missingValues)>0 & any(missingValues > maxViralAbundance)){
    lastDetectable <- min(missingValues[which(missingValues > maxViralAbundance)])
    endPt <- lastDetectable
    }

  # 2. check for values less than all subsequent values in detectable run
  remainingNonmissing <- which(is.finite(data$y_virus[1:endPt]))
  minBeforeEnd <- sapply(1:length(remainingNonmissing), whichLower,
                        vec = data$y_virus[remainingNonmissing])
  if(any(minBeforeEnd)){
    if(any(remainingNonmissing[which(minBeforeEnd)]>maxViralAbundance)){
      endPt <- min(remainingNonmissing[minBeforeEnd])
      }
    }

  # 3. check for earlier end of infection wave
    # defined as a local minimum with a derivative going from negative
    # to positive twice on either side
  if(endPt>=5){
    earlyLocalMinTest <- sapply(3:(endPt-2), function(x) all(sign(diff(data$y_virus[max(c(1, x-2)):min(c(endPt,(x+2)))]))==c(-1,-1,1,1)))
  if(any(earlyLocalMinTest,na.rm=T)){
    endPt <- which(earlyLocalMinTest)+2
    }
  }

  return(endPt)
}

noro.endpoints <- do.call(rbind, 
                         lapply(noroParticipants, 
                                function(x) data.frame(x, 
                                                       end.point(noro.FShedData[noro.FShedData$ID == x,]))))
colnames(noro.endpoints) <- c('ID', 'endPt')

sars.endpoints <- do.call(rbind, 
                         lapply(sarsParticipants, 
                                function(x) data.frame(x, 
                                                       end.point(sars.TShedData[sars.TShedData$ID == x,]))))
colnames(sars.endpoints) <- c('ID', 'endPt')

```

Note that because this model is statistical, not deterministic, the parameters
will fluctuate with each run. We account for this uncertainty later on by 
producing confidence bounds in our analyses via bootstrapping/case resampling.
```{r, eval = FALSE}
numFits <- 1000

noro.fits <- all.fits(noro.FShedData, noro.endpoints, numFits,
                                    detect.threshold = log10(noro.detection.threshold),
                                    detect.threshold.2 = log10(noro.detection.threshold.2))

sars.fits <- all.fits(sars.TShedData, sars.endpoints, numFits,
                                    detect.threshold = log10(sars.detection.threshold))
```
 
### Calculating our measures of interest

Now, we calculate the delay between symptom onset and peak pathogen load for
each participant in each study, which serves as a conservative measure of 
pre-symptomatic transmission. We also look at a 
non-conservative measure of pre-symptomatic transmission, i.e. the 
delay between shedding onset and symptom onset. A negative delay indicates 
pre-symptomatic transmission, while a positive one indicates post-symptomatic
transmission. The delay is in days.

```{r}
calc.measures <- function(fits, participants, shedding.data,
                           incub.data) {
  if (!is.null(incub.data$dose[1])) {
    fits$dose <- rep(incub.data$dose, each = numFits)
  } else {
    fits$dose <- rep(10, each = numFits)
  }
  
  fits$incubation <- rep(incub.data$incubation, each = numFits)
  fits$sheddingOnset <- rep(unlist(lapply(participants, 
                                                       function(x) 
    shed.onset(shedding.data[shedding.data$ID == x, ]))), each = numFits)
  fits$repRate <- log10(fits$transp2)
  fits$peakDay <- fits$transp3/24
  fits$symptomsToPeakDay <- fits$peakDay - fits$incubation
  fits$symptomsToStartShedding <- fits$sheddingOnset - fits$incubation
  
  return(fits)
}

noro.fits <- calc.measures(noro.fits, noroParticipants,
                                  noro.FShedData, noro.IncubData)

sars.fits <- calc.measures(sars.fits, sarsParticipants,
                                  sars.TShedData, sars.IncubData)


noro.params <- participant.fits.minsse(noro.fits)
sars.params <- participant.fits.minsse(sars.fits)
```

We exclude participants from some of the upcoming analyses due to unclear peaks
or not enough data points.
```{r}
positivePts = function(data){
  quantifiedAbundance = is.finite(data$y_virus)
  positiveDetection = quantifiedAbundance
  if("y_type" %in% colnames(data)){positiveDetection = grepl("POS", data$y_type)}
  return(length(which(quantifiedAbundance | positiveDetection)))
  }

# check how many points were fitted
noro.PtsFitted <- do.call(rbind, 
                         lapply(noroParticipants, 
                                function(x) data.frame(x, 
                                                       positivePts((noro.FShedData[noro.FShedData$ID == x,])
                                                                   [1:noro.endpoints$endPt[noro.endpoints$ID==x],]))))
colnames(noro.PtsFitted) <- c('ID', 'points.fit')

# check how many points were fitted
sars.PtsFitted <- do.call(rbind, 
                         lapply(sarsParticipants, 
                                function(x) data.frame(x, 
                                                       positivePts((sars.TShedData[sars.TShedData$ID == x,])
                                                                   [1:sars.endpoints$endPt[sars.endpoints$ID==x],]))))
colnames(sars.PtsFitted) <- c('ID', 'points.fit')

noro.unclearpeaks <- c(noro.PtsFitted[noro.PtsFitted$points.fit < 5,]$ID)

sars.unclearpeaks <- c('Throat_2', sars.PtsFitted[sars.PtsFitted$points.fit < 5,]$ID)

noro.fits.clearpeaks <- noro.fits[which(!noro.fits$ID%in%noro.unclearpeaks),]

noro.params.clearpeaks <- noro.params[which(!noro.params$ID%in%noro.unclearpeaks),]

sars.fits.clearpeaks <- sars.fits[which(!sars.fits$ID%in%sars.unclearpeaks),]

sars.params.clearpeaks <- sars.params[which(!sars.params$ID%in%sars.unclearpeaks),]
```

We then use bootstrapping to create 95% confidence intervals (n = 1000) around 
the linear regressions we show in Figures 3 and S6 as well as the mean values
per dose we show in Figures 3 and S4.
```{r} 
# Produce projected values as per a linear regression with predictor 
# variable = viral replication rate
lmMetrics <- function(dataSet, columnName, xValues = NA){
  lmFit <- lm(formula = reformulate("repRate", response = columnName), data = dataSet)
  res <- predict.lm(lmFit, newdata = data.frame(repRate = xValues))
  return(res)
} 

rep.summary.table <- function(fits, params, iters, n_points) {
  
  r.range <- seq(min(params$repRate),
                    max(params$repRate),
                    length=n_points)

  summary <- data.frame(t = r.range,
                         peakDay = lmMetrics(params, 'peakDay', r.range),
                         symp = lmMetrics(params, 'incubation', r.range), 
                         sympToPeakDay = lmMetrics(params, 'symptomsToPeakDay', r.range),
                         shedding = lmMetrics(params, 'sheddingOnset', r.range),
                         sympToShed = lmMetrics(params, 'symptomsToStartShedding', r.range)
                        )
  
  peakDayResamples <- matrix(NA, nrow = iters, ncol = n_points)
  sympResamples <- matrix(NA, nrow = iters, ncol = n_points)
  sympToPeakDayResamples <- matrix(NA, nrow = iters, ncol = n_points)
  sheddingResamples <- matrix(NA, nrow = iters, ncol = n_points)
  sympToShedResamples <- matrix(NA, nrow = iters, ncol = n_points)
  
  for(i in 1:iters){
    params.resampled <- participant.fits.resampled(fits)
  
    peakDayResamples[i, ] <- lmMetrics(params.resampled, 'peakDay', r.range)
    sympResamples[i, ] <- lmMetrics(params.resampled, 'incubation', r.range)
    sympToPeakDayResamples[i, ] <- lmMetrics(params.resampled, 'symptomsToPeakDay', r.range)
    sheddingResamples[i, ] <- lmMetrics(params.resampled, 'sheddingOnset', r.range)
    sympToShedResamples[i, ] <- lmMetrics(params.resampled, 'symptomsToStartShedding', r.range)
      
  }
  
  summary$peakDayLCI <- 
    apply(peakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$peakDayUCI <- 
    apply(peakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  summary$sympLCI <- 
    apply(sympResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$sympUCI <- 
    apply(sympResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  summary$sympToPeakDayLCI <- 
    apply(sympToPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$sympToPeakDayUCI <- 
    apply(sympToPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  summary$sheddingLCI <- 
    apply(sheddingResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$sheddingUCI <- 
    apply(sheddingResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  summary$sympToShedLCI <- 
    apply(sympToShedResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$sympToShedUCI <- 
    apply(sympToShedResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  return(summary)
}

dose.summary.table <- function(fits, params, iters, pathogen) {
  
  doses <- rev(unique(params$dose))
  
  n_points <- length(doses)

  if(pathogen == 'Norovirus') {
    summary <- rev(as.data.frame(params %>% group_by(dose) %>% summarise(across(c(symptomsToPeakDay,
                                                     incubation,
                                                     peakDay,
                                                     repRate), 
                                            ~ mean(.x, na.rm=TRUE)))))
    
    avgPeakDayResamples <- matrix(NA, nrow = iters, ncol = n_points)
    avgRepRateResamples <- matrix(NA, nrow = iters, ncol = n_points)
  } else {
    summary <- as.data.frame(params %>% group_by(dose) %>% summarise(across(symptomsToPeakDay, 
                                            ~ mean(.x, na.rm=TRUE))))
  }
  
  avgSympToPeakDayResamples <- matrix(NA, nrow = iters, ncol = n_points)
  
  if(pathogen == 'Norovirus') {
    for (i in 1:iters) {
      params.resampled <- participant.fits.resampled(fits)
      
      avgSympToPeakDayResamples[i, ] <- unlist(lapply(doses, function(d) {
        mean(params.resampled[params.resampled$dose == d,]$symptomsToPeakDay)
      }))
      
      avgPeakDayResamples[i, ] <- unlist(lapply(doses, function(d) {
        mean(params.resampled[params.resampled$dose == d,]$peakDay)
      }))
      
      avgRepRateResamples[i, ] <- unlist(lapply(doses, function(d) {
        mean(params.resampled[params.resampled$dose == d,]$repRate)
      }))
    }
    
    summary$avgPeakDayLCI <- 
      apply(avgPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
    summary$avgPeakDayUCI <- 
      apply(avgPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
    
    summary$avgRepRateLCI <- 
      apply(avgRepRateResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
    summary$avgRepRateUCI <- 
      apply(avgRepRateResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  } else {
    for (i in 1:iters) {
      params.resampled <- participant.fits.resampled(fits)
      
      avgSympToPeakDayResamples[i, ] <- unlist(lapply(doses, function(d) {
        mean(params.resampled[params.resampled$dose == d,]$symptomsToPeakDay, na.rm=T)
      }))
      
    }
  }
  
  summary$avgSympToPeakDayLCI <- 
    apply(avgSympToPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.025))
  summary$avgSympToPeakDayUCI <- 
    apply(avgSympToPeakDayResamples, MARGIN = 2, FUN = function(c) quantile(c, 0.975))
  
  return(summary)
}

noroSummary.rep <- rep.summary.table(noro.fits.clearpeaks, 
                                 noro.params.clearpeaks,
                                 1000, 10)

sarsSummary.rep <- rep.summary.table(sars.fits.clearpeaks,
                                 sars.params.clearpeaks,
                                 1000, 10)

noroSummary.dose <- dose.summary.table(noro.fits.clearpeaks,
                                       noro.params.clearpeaks,
                                       1000,
                                       'Norovirus')

sarsSummary.dose <- dose.summary.table(sars.fits.clearpeaks,
                                       sars.params.clearpeaks,
                                       1000,
                                       'SARS-CoV-2')
```

We also calculate pre-symptomatic transmission potential,
measured as average shedding prior to symptoms, and the duration of
pre-symptomatic shedding (symptom onset minus shedding onset) for each person.
```{r}
presymp.transmission.potential <- function(vldata, sympdata, measure='concentration') {
  
  vl.presymp <- vldata[vldata$days < sympdata$incubation,]
  
  if(length(vl.presymp[grepl('~', vl.presymp$y_type),]$y_virus) > 0) {
    vl.presymp[grepl('~', vl.presymp$y_type),]$y_virus <- mean(c(15000, 4e7))
  }
  
  if(measure == 'total shedding') {
    avg.presymp.shedding <- log10(mean(vl.presymp$total_shedding, na.rm=T))
  } else {
    avg.presymp.shedding <- log10(mean(vl.presymp$y_virus, na.rm=T))
  }

  return(avg.presymp.shedding)
  
}

add.presymp.transmission.potential <- function(params, shedding.data,
                                               incub.data) {
  
  presymptrans.avg <- apply(params, 1, function(x) 
    presymp.transmission.potential(shedding.data[shedding.data$ID == x[1],],
                                   incub.data[incub.data$ID == x[1],]))
  
  presymp.dur <- apply(params, 1, 
                      function(x) -1 * params[params$ID == x[1],]$symptomsToStartShedding)
  
  params$avgSheddingBeforeSymptoms <- presymptrans.avg
  params$sheddingTimeBeforeSymptoms <- presymp.dur
  
  return(params)
}

noro.params <- add.presymp.transmission.potential(noro.params,
                                                  noro.FShedData,
                                                  noro.IncubData)

sars.params <- add.presymp.transmission.potential(sars.params,
                                                  sars.TShedData,
                                                  sars.IncubData)


```

We also used cumulative pre-symptomatic shedding as an alternative measure of
pre-symptomatic transmission potential. For norovirus, we look at volume
of each fecal sample * concentration of virus in each sample.
```{r}
noro.concentrationRaw <- read.csv(here("data", "Ge stool concentrations.csv"))
noro.concentrationRaw <- noro.concentrationRaw[noro.concentrationRaw$studyday > 0,]
noro.concentrationRaw <- noro.concentrationRaw[(noro.concentrationRaw$studyday - 1) <=
                                                 max(noro.IncubData$incubation),]
noro.concentrationRaw <- noro.concentrationRaw[
  noro.concentrationRaw$qual.pcr.result == 'POS',]

noro.weights <- read.csv(here("data", "Ge stool sample weights.csv"))
atmarIDs <- read.csv(here("data", "atmarIDs.csv"))

# Using peak viral load, inoculum dose, and type of symptoms to match up participant IDs

geSymptomData <- noroData$mvs
geSymptomData$type <- ""
geSymptomData$type[geSymptomData$mvsc3>0] <- "v"
# mvsc3 indicates duration of vomiting
# mvsc1 indicates duration of diarrhea
geSymptomData$type[geSymptomData$mvsc3>0 & geSymptomData$mvsc1>0] <- "dv"

# Participant ID4 had reported viral shedding in vomit; we and the authors
# of the study do not know why mvsc3 = 0
geSymptomData[geSymptomData$ID == 'ID4',]$type <- "dv"

geIDs <- do.call(rbind, 
                  lapply(noroParticipants,
                         function(x) data.frame(
                           geID=x,
                           dose=unique(noro.FShedData[noro.FShedData$ID == x,]$dose),
                           peak=round(log10(max(noro.FShedData[noro.FShedData$ID == x,]$y_virus, na.rm=T)), 1),
                           symptom.type=geSymptomData[geSymptomData$ID==x,]$type)))

ID_dict_noro <- merge(atmarIDs, geIDs, by=c("dose", "peak", "symptom.type"))[, c("atmarID", "geID")]

# Manually adding in ID10/720 because their trajectories are identical, but
# their symptoms do not match up between the two datasets
ID_dict_noro <- rbind(ID_dict_noro, 
                          data.frame(atmarID='720', geID = 'ID10'))

noro.concentrationRaw <- merge(noro.concentrationRaw, ID_dict_noro,
                               by.x = 'volnum', by.y = 'atmarID')

# Replace dates with time since inoculation (by matching up viral concentration
# of the samples, then for qualitative results, by manually filling in days 
# as order does not matter here)
noro.concentrationRaw <-  merge(noro.concentrationRaw, 
                                noro.FShedData.unaggregated[, c("days", "y_virus")],
                               by.x = 'quantity', by.y = 'y_virus', all.x = TRUE)

noro.concentrationRaw <- read.csv(here("data", "Ge stool sample concentrations mapped days.csv"))

# Combine with sample weight data to calculate shedding data
noro.FShedData.weightInc <- merge(noro.concentrationRaw, 
                                  noro.weights[, c("stoollognum", "weight")],
                                  by.x = 'lognum', by.y='stoollognum')

noro.FShedData.weightInc$total_shedding <- ifelse(grepl('<', noro.FShedData.weightInc$quantity),
                                                  mean(c(15000, 4e7)) * noro.FShedData.weightInc$weight,
                                                  as.numeric(noro.FShedData.weightInc$quantity) * noro.FShedData.weightInc$weight)

# Average total shedding values at the same time point
avg_shedding <- aggregate(total_shedding ~ geID + days, 
                              noro.FShedData.weightInc,
                              mean)

avg_shedding <- avg_shedding[order(as.numeric(substring(avg_shedding$geID, 3, 4)),
                       avg_shedding$days),]$total_shedding

noro.FShedData.weightInc <- noro.FShedData.weightInc[
  !duplicated(noro.FShedData.weightInc[, c('geID', 'days')]),]

noro.FShedData.weightInc <- noro.FShedData.weightInc[
  order(as.numeric(substring(noro.FShedData.weightInc$geID, 3, 4)),
                       noro.FShedData.weightInc$days),]

noro.FShedData.weightInc$total_shedding <- avg_shedding

noro.params$avgTotalSheddingBeforeSymptoms <- apply(noro.params, 1,  
                               function(x) presymp.transmission.potential(
                                 noro.FShedData.weightInc[noro.FShedData.weightInc$geID == x[1],],
                                 noro.IncubData[noro.IncubData$ID == x[1],],
                                 'total shedding'))
```

### Linear regressions
We run linear regressions between the delay between symptom onset and peak
pathogen load and each of the 4 parameters from the statistical model as well as 
inoculum dose. We also run linear regressions between viral replication and 
decay rate.
```{r}
# Peak viral load vs. delay
lm.fit.p1.delay.conservative.noro <- 
  lm(symptomsToPeakDay ~ p1, data=noro.params.clearpeaks)
p1.delay.conservative.pval.noro <- 
  summary(lm.fit.p1.delay.conservative.noro)$coefficients[2,4]
p1.delay.conservative.pval.noro

# Viral replication rate vs. delay
lm.fit.p2.delay.conservative.noro <- 
  lm(symptomsToPeakDay ~ p2, data=noro.params.clearpeaks)
p2.delay.conservative.pval.noro <- 
  summary(lm.fit.p2.delay.conservative.noro)$coefficients[2,4]
p2.delay.conservative.pval.noro
p2.delay.conservative.rsquared.noro <- summary(lm.fit.p2.delay.conservative.noro)$r.squared
p2.delay.conservative.rsquared.noro

# Time of peak viral load vs. delay
lm.fit.p3.delay.conservative.noro <- 
  lm(symptomsToPeakDay ~ p3, data=noro.params.clearpeaks)
p3.delay.conservative.pval.noro <- 
  summary(lm.fit.p3.delay.conservative.noro)$coefficients[2,4]
p3.delay.conservative.pval.noro

# Viral decay rate vs. delay
lm.fit.p4.delay.conservative.noro <- 
  lm(symptomsToPeakDay ~ p4, data=noro.params.clearpeaks)
p4.delay.conservative.pval.noro <- 
  summary(lm.fit.p4.delay.conservative.noro)$coefficients[2,4]
p4.delay.conservative.pval.noro

# Delay vs. dose 
lm.fit.dose.delay.conservative.noro <- 
  lm(symptomsToPeakDay ~ dose, data=noro.params.clearpeaks)
dose.delay.conservative.pval.noro <- 
  summary(lm.fit.dose.delay.conservative.noro)$coefficients[2,4]
dose.delay.conservative.pval.noro

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative.noro <- 
  lm(p4 ~ p2, data=noro.params.clearpeaks)
p4.p2.conservative.pval.noro <- 
  summary(lm.fit.p4.p2.conservative.noro)$coefficients[2,4]
p4.p2.conservative.pval.noro
```

We do the same for SARS-CoV-2, barring delay vs. dose since only one
inoculum dose was used.
```{r}
# Peak viral load vs. delay
lm.fit.p1.delay.conservative.sars <- 
  lm(symptomsToPeakDay ~ p1, data=sars.params.clearpeaks)
p1.delay.conservative.pval.sars <- 
  summary(lm.fit.p1.delay.conservative.sars)$coefficients[2,4]
p1.delay.conservative.pval.sars

# Viral replication rate vs. delay
lm.fit.p2.delay.conservative.sars <- 
  lm(symptomsToPeakDay ~ p2, data=sars.params.clearpeaks)
p2.delay.conservative.pval.sars <- 
  summary(lm.fit.p2.delay.conservative.sars)$coefficients[2,4]
p2.delay.conservative.pval.sars

# Time of peak viral load vs. delay
lm.fit.p3.delay.conservative.sars <- 
  lm(symptomsToPeakDay ~ p3, data=sars.params.clearpeaks)
p3.delay.conservative.pval.sars <- 
  summary(lm.fit.p3.delay.conservative.sars)$coefficients[2,4]
p3.delay.conservative.pval.sars

# Viral decay rate vs. delay
lm.fit.p4.delay.conservative.sars <- 
  lm(symptomsToPeakDay ~ p4, data=sars.params.clearpeaks)
p4.delay.conservative.pval.sars <- 
  summary(lm.fit.p4.delay.conservative.sars)$coefficients[2,4]
p4.delay.conservative.pval.sars

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative.sars <- 
  lm(p4 ~ p2, data=sars.params.clearpeaks)
p4.p2.conservative.pval.sars <- 
  summary(lm.fit.p4.p2.conservative.sars)$coefficients[2,4]
p4.p2.conservative.pval.sars
```

We also test for significant correlations between the statistical 
model parameters plus dose and the duration of shedding prior to
symptoms.
```{r}
# Peak viral load vs. delay (non-conservative)
lm.fit.p1.delay.nonconservative.noro <- 
  lm(symptomsToStartShedding ~ p1, data=noro.params.clearpeaks)
p1.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p1.delay.nonconservative.noro)$coefficients[2,4]
p1.delay.nonconservative.pval.noro

# Viral replication rate vs. delay (non-conservative)
lm.fit.p2.delay.nonconservative.noro <- lm(symptomsToStartShedding ~ p2, 
                                           data=noro.params.clearpeaks)
p2.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p2.delay.nonconservative.noro)$coefficients[2,4]
p2.delay.nonconservative.pval.noro

# Time of peak viral load vs. delay (non-conservative)
lm.fit.p3.delay.nonconservative.noro <- 
  lm(symptomsToStartShedding ~ p3, data=noro.params.clearpeaks)
p3.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p3.delay.nonconservative.noro)$coefficients[2,4]
p3.delay.nonconservative.pval.noro

# Viral decay rate vs. delay (non-conservative)
lm.fit.p4.delay.nonconservative.noro <- 
  lm(symptomsToStartShedding ~ p4, data=noro.params.clearpeaks)
p4.delay.nonconservative.pval.noro <- 
  summary(lm.fit.p4.delay.nonconservative.noro)$coefficients[2,4]
p4.delay.nonconservative.pval.noro

# Delay (non-conservative) vs. dose 
lm.fit.dose.delay.nonconservative.noro <- 
  lm(symptomsToStartShedding ~ dose, data=noro.params)
dose.delay.nonconservative.pval.noro <- 
  summary(lm.fit.dose.delay.nonconservative.noro)$coefficients[2,4]
dose.delay.nonconservative.pval.noro

# Same with SARS-CoV-2 data barring dose
lm.fit.p1.delay.nonconservative.sars <- 
  lm(symptomsToStartShedding ~ p1, data=sars.params.clearpeaks)
p1.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p1.delay.nonconservative.sars)$coefficients[2,4]
p1.delay.nonconservative.pval.sars

lm.fit.p2.delay.nonconservative.sars <- 
  lm(symptomsToStartShedding ~ p2, data=sars.params.clearpeaks)
p2.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p2.delay.nonconservative.sars)$coefficients[2,4]
p2.delay.nonconservative.pval.sars

lm.fit.p3.delay.nonconservative.sars <- 
  lm(symptomsToStartShedding ~ p3, data=sars.params.clearpeaks)
p3.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p3.delay.nonconservative.sars)$coefficients[2,4]
p3.delay.nonconservative.pval.sars

lm.fit.p4.delay.nonconservative.sars <- 
  lm(symptomsToStartShedding ~ p4, data=sars.params.clearpeaks)
p4.delay.nonconservative.pval.sars <- 
  summary(lm.fit.p4.delay.nonconservative.sars)$coefficients[2,4]
p4.delay.nonconservative.pval.sars
```

There is no correlation between average shedding prior to symptoms and the 
duration of shedding in the pre-symptomatic period (symptom onset minus shedding 
onset) in either the norovirus or SARS-CoV-2 data.
```{r}
transmission.presymp.avg.fit.noro <- lm(avgSheddingBeforeSymptoms ~ sheddingTimeBeforeSymptoms, 
                              data = noro.params)
transmission.presymp.avg.pval.noro <- 
  summary(transmission.presymp.avg.fit.noro)$coefficients[2,4]

transmission.presymp.avg.pval.noro

transmission.presymp.avg.totalshedding.fit.noro <- lm(avgTotalSheddingBeforeSymptoms ~ 
                                                        sheddingTimeBeforeSymptoms, data=noro.params)
transmission.presymp.avg.totalshedding.pval.noro <- summary(transmission.presymp.avg.totalshedding.fit.noro)$
  coefficients[2, 4]

transmission.presymp.avg.totalshedding.pval.noro

transmission.presymp.avg.fit.sars <- lm(avgSheddingBeforeSymptoms ~ sheddingTimeBeforeSymptoms, 
                              data = sars.params)
transmission.presymp.avg.pval.sars <- 
  summary(transmission.presymp.avg.fit.sars)$coefficients[2,4]

transmission.presymp.avg.pval.sars
```

Lastly, we break down the correlation between pre-symptomatic transmission
and viral replication rate by examining the relationships between time of
peak shedding and replication rate as well as between symptom onset
and replication rate.

For both norovirus and SARS-CoV-2, there is a significant correlation between
time of peak shedding and viral replication rate, but not between symptom onset 
and replication rate.
```{r}
lm.fit.p2.peaktime.noro <- lm(peakDay ~ p2, 
                              data=noro.params.clearpeaks)
p2.peaktime.pval.noro <- summary(lm.fit.p2.peaktime.noro)$coefficients[2,4]
p2.peaktime.pval.noro
p2.peaktime.rsquared.noro <- summary(lm.fit.p2.peaktime.noro)$r.squared
p2.peaktime.rsquared.noro

lm.fit.p2.shedonset.noro <- lm(sheddingOnset ~ p2, 
                                           data=noro.params.clearpeaks)
p2.shedonset.pval.noro <- 
  summary(lm.fit.p2.shedonset.noro)$coefficients[2,4]
p2.shedonset.pval.noro
p2.shedonset.rsquared.noro <- 
  summary(lm.fit.p2.shedonset.noro)$r.squared
p2.shedonset.rsquared.noro

lm.fit.p2.symponset.noro <- lm(incubation ~ p2, data=noro.params.clearpeaks)
p2.symponset.pval.noro <- summary(lm.fit.p2.symponset.noro)$coefficients[2,4]
p2.symponset.pval.noro

lm.fit.dose.symponset.noro <- lm(incubation ~ dose,
                                 data=noro.params)
dose.symponset.pval.noro <- summary(lm.fit.dose.symponset.noro)$coefficients[2,4]
dose.symponset.pval.noro

lm.fit.p2.peaktime.sars <- lm(peakDay ~ p2, data=sars.params.clearpeaks)
p2.peaktime.pval.sars <- summary(lm.fit.p2.peaktime.sars)$coefficients[2,4]
p2.peaktime.pval.sars
p2.peaktime.rsquared.sars <- summary(lm.fit.p2.peaktime.sars)$r.squared
p2.peaktime.rsquared.sars

lm.fit.p2.shedonset.sars <- lm(sheddingOnset ~ p2, 
                                           data=sars.params.clearpeaks)
p2.shedonset.pval.sars <- 
  summary(lm.fit.p2.shedonset.sars)$coefficients[2,4]
p2.shedonset.pval.sars

lm.fit.p2.symponset.sars <- lm(incubation ~ p2, data=sars.params.clearpeaks)
p2.symponset.pval.sars <- summary(lm.fit.p2.symponset.sars)$coefficients[2,4]
p2.symponset.pval.sars
```

### Plotting code

We plot trajectories for every individual.
```{r, fig.height = 7.5, fig.width = 3.5}
plot.viralload <- function(data, pars, sympdata, t.range, cex=1,
                           xaxisLabs = T, yaxisLabs = T, col, pch, ylim.lower=3,
                           cex_labels, cex.axis, xlabel, ylabel, panellabel,
                           example=FALSE) {

  # Plotting real shedding values
  plot(1, col=col, xlim=c(0, 25), 
       ylim=c(ylim.lower, 15), xlab="", ylab="", xaxt="n", yaxt="n", cex.lab=cex, 
       cex.axis=cex, font.lab=2, bty='n')
  
  points(data$days, log10(data$y_virus), col=col, cex=cex, pch=pch)
  
  # Grey rectangles where viral shedding is a range (qualitative results)
  range.indexes <- which(is.na(data$y_virus) & grepl('POS', data$y_type))
  rectangle.x_vals <- lapply(range.indexes, function(x) c(data[x-1,]$days + 0.25, 
                                                          data[x+1,]$days - 0.25))
  lapply(rectangle.x_vals, function(x) 
    rect(x[1], log10(15000), x[2], log10(4e7), border=NA, col="lightgray"))
  
  # Black lines where viral shedding is undetectable
  if ('y_type' %in% colnames(data)) {
    undetectable.indexes <- which(grepl('NEG', data$y_type))
  } else {
    undetectable.indexes <- which(is.na(data$y_virus) & data$days > 0)
  }
  lines.x_vals <- lapply(undetectable.indexes, function(x) c(data[x,]$days - 0.25, 
                                                          data[x,]$days + 0.25))
  
  lapply(lines.x_vals, lines, y=rep(ylim.lower, 2), lwd=2)
  
  # Plotting VL trajectories
  predValues <- sse.viralLoad(pars, data[1:t.range$endPt,], optimization = F)
  linet <- predValues$t
  predictedvl <- predValues$predData
  
  predictedvl[!is.finite(predictedvl)] <- 0

  lines(linet / 24, predictedvl, col=col)
  
  # Symptom onset
  symp.onset <- sympdata$incubation
  
  abline(v = symp.onset, col=col, lty=2)
  
  # Peak shedding
  t.peak <- sse.viralLoad(pars, data[1:t.range$endPt,], optimization = NA)[3]/24
  peakHeight <- sse.viralLoad(pars, data[1:t.range$endPt,], optimization = NA)[1]

  if (is.na(symp.onset)) {
    redpointstyle <- NA
  } else if (t.peak < symp.onset) {
    redpointstyle <- 19
  } else {
    redpointstyle <- 1
  }
  
  if (!example) {
    text(x=0.8 * 25, y=0.8 * (15 - ylim.lower) + ylim.lower, 
         labels=panellabel, font=2)
  }
  
  ylabel.line <- 4.5
  xlabel.line <- 3.5

  points(t.peak, peakHeight, pch=redpointstyle,
         col="red", lwd=cex, cex=cex)

  xLabs = NA
  if(xaxisLabs==T){
        xLabs = c(expression(0),
                  expression(5),
                  expression(10),
                  expression(15),
                  expression(20),
                  expression(25))}
  axis(1, cex.axis=cex.axis, at=seq(0, 25, by=5), labels = xLabs,
       line=1.25,
       mgp=c(2, 0.75, 0), tck=-0.04)
  
  yLabs = NA
  if(yaxisLabs==T){
      yLabs = c(expression(10^3),
                expression(10^6),
                expression(10^9),
                expression(10^12),
                expression(10^15))}
  axis(2, cex.axis=cex.axis, at=seq(3, 15, by=3), 
       labels=yLabs,
       line=1.25,
       las=2, 
       mgp=c(2, 1, 0), tck= -0.04)
  
  if (example) {
    mtext(text=xlabel, side=1, line=xlabel.line, cex=cex_labels, adj=3.25)
    mtext(text=ylabel, side=2, line=ylabel.line, cex=cex_labels, adj=5)
    text(.93 * 25, .93 * (15-3) + 3, panellabel, cex=cex_labels,
         font=2)
  }
  else {
    mtext(text=xlabel, side=1, line=xlabel.line, cex=cex_labels)
    mtext(text=ylabel, side=2, line=ylabel.line, cex=cex_labels)
  }
  
}
```

First, we show examples of pre- and post-symptomatic transmission.
```{r}
sars.color <- "#56b3e9"
noro.color <- "#871a6e"
sars.pch <- 5
noro.pch <- 2

tiff(filename="Example Trajectories.tiff", height=5, 
     width=6.5, units = "in", res=500)

participantexamples <- c('Throat_16', 'Throat_5', 'ID13')

par(mfrow=c(2, 2), mai=c(0.42, 0.52, 0.52, 0.1), oma=c(2, 3, 1, 2))

for(i in 1: length(participantexamples)) {
  
  p <- participantexamples[i]
  
  if(grepl('Throat', p)) {
    participantdata <- sars.TShedData[sars.TShedData$ID==p, ]
    sympdata <- sars.IncubData[sars.IncubData$ID==p, ]
    params <- unlist(sars.params.clearpeaks[sars.params.clearpeaks$ID==p,][2:5])
    t.range <- sars.endpoints[sars.endpoints$ID==p,]
    col <- sars.color
    pch <- sars.pch
  } else {
    participantdata <- noro.FShedData[noro.FShedData$ID==p, ]
    sympdata <- noro.IncubData[noro.IncubData$ID == p, ]
    params <- unlist(noro.params[noro.params$ID==p,][2:5])
    t.range <- noro.endpoints[noro.endpoints$ID==p,]
    col <- noro.color
    pch <- noro.pch
  }

  if (i == 3) {
    ylabel <- "Viral shedding"
    xlabel <- "Hours after inoculation"
  } else {
    ylabel <- ""
    xlabel <- ""
  }
  
  plot.viralload(data=participantdata, 
                 pars=params, 
                 sympdata=sympdata, 
                 t.range=t.range,
                 cex=1.33, 
                 col=col,
                 pch=pch,
                 cex_labels=1,
                 cex.axis=1,
                 xlabel=xlabel,
                 ylabel=ylabel,
                 panellabel=LETTERS[i],
                 example=TRUE)
  
  if (i == 1) {
    mtext("Post-symptomatic", side=3, line=2, font=2)
  } else if (p==participantexamples[2]) {
    mtext("Pre-symptomatic", side=3, line=2, font=2)
  }
}

plot.new()

legend("center", legend=c("SARS-CoV-2", "Norovirus"), col=c(sars.color, noro.color),
       pch=c(5, 2))

dev.off()
```

We then plot the trajectories of each participant from both studies. For norovirus,
we compare fits with those from Ge *et al.*, 2023.
```{r, fig.height = 7.5, fig.width = 6.5}
noroDataBayesian <- readRDS(here("data", "holder_curves_mu_id_con20.rds")) 
noro.FShedDataBayesian <- bind_rows(noroDataBayesian)

tiff("Participant trajectories comparison.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(noroParticipants)) {
  
  p <- noroParticipants[i]
  participantdata <- noro.FShedData[noro.FShedData$ID==p, ]
  sympdata <- noro.IncubData[noro.IncubData$ID==p, ]
  params <- unlist(noro.params[noro.params$ID==p,][2:5])
  participantdata.bayesian <- noro.FShedDataBayesian[noro.FShedDataBayesian$ID == p,]
  t.range <- noro.endpoints[noro.endpoints$ID==p,]
  col <- noro.color
  pch <- noro.pch
  
  if(p %in% noro.unclearpeaks) {
    plab <- paste(unique(participantdata$ID), ',\n excluded', sep='')
  } else{
    plab <- unique(participantdata$ID)
  }
  
  if (i==16) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      t.range=t.range,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding",
      panellabel=plab)
  } else if (i %in% seq(1, 15, by=3)) {
    plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      t.range=t.range,
      xaxisLabs=F,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding",
      panellabel=plab)
  } else if (i %in% c(15, 17)) {
     plot.viralload(
      data=participantdata,
      pars=params,
      sympdata=sympdata,
      t.range=t.range,
      yaxisLabs=F,
      col=col,
      pch=pch,
      ylim.lower = 7,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="",
      panellabel=plab)
  } else {
    plot.viralload(
    data=participantdata,
    pars=params,
    sympdata=sympdata,
    t.range=t.range,
    xaxisLabs=F,
    yaxisLabs=F,
    col=col,
    pch=pch,
    ylim.lower = 7,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="",
    panellabel=plab)
  }
  
  lines(participantdata.bayesian$days, 
        log10(exp(participantdata.bayesian$Estimate)), col="#22a841", lty=3)
  
}

plot.new()

legend("bottom", legend=c("Ge et al., 2023 fits", "Our fits"), lty=c(3, 1),
       col=c("#22a841", noro.color), bty='n')

dev.off()

tiff("Participant Trajectories SARS-CoV-2.tiff", res=500,
      height=7.5, width=6.5, unit="in")

par(mfrow=c(6, 3), mai=c(0.22, 0.42, 0.22, 0.22), oma=c(3.5, 4, 0, 0),
    mar=c(1, 2.1, 1, 1))

for(i in 1:length(sarsParticipants)) {
  
  p <- sarsParticipants[i]
  participantdata <- sars.TShedData[sars.TShedData$ID==p, ]
  sympdata <- sars.IncubData[sars.IncubData$ID==p, ]
  params <- unlist(sars.params[sars.params$ID==p,][2:5])
  t.range <- sars.endpoints[sars.endpoints$ID==p,]
  
  if(p %in% sars.unclearpeaks) {
    plab <- paste(unique(participantdata$ID), ',\n excluded', sep='')
  } else{
    plab <- unique(participantdata$ID)
  }
  
  if (i %in% seq(1, 14, by=3)) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      t.range=t.range,
      xaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="",
      ylabel="Viral shedding",
      panellabel=plab)
  } else if (i %in% c(15, 17)) {
     plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      t.range=t.range,
      yaxisLabs=F,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="",
      panellabel=plab)
  } else if (i==16) {
    plot.viralload(
      data=participantdata,
      sympdata=sympdata,
      pars=params,
      t.range=t.range,
      col=sars.color,
      pch=sars.pch,
      cex_labels=0.8,
      cex.axis=1.25,
      xlabel="Days after inoculation",
      ylabel="Viral shedding",
      panellabel=plab)
  } else {
    plot.viralload(
    data=participantdata,
    sympdata=sympdata,
    pars=params,
    t.range=t.range,
    xaxisLabs=F,
    yaxisLabs=F,
    col=sars.color,
    pch=sars.pch,
    cex_labels=0.8,
    cex.axis=1.25,
    xlabel="",
    ylabel="",
    panellabel=plab)
  }
}

dev.off()
```

We plot the delay between symptom onset and peak viral shedding for each virus,
delay vs. viral replication rate, day of peak shedding vs. viral replication rate,
day of symptom onset vs. viral replication rate, and average viral shedding 
prior to symptoms vs. duration of pre-symptomatic shedding in a 5-panel plot.
```{r, fig.height = 7.5, fig.width = 3.5}
measure.names <- c("ID", "dose", "incubation", "sheddingOnset", "repRate",
                   "peakDay", "symptomsToPeakDay", "symptomsToStartShedding",
                   "avgSheddingBeforeSymptoms", "sheddingTimeBeforeSymptoms")
measures.combined <- rbind(noro.params[colnames(noro.params) %in% measure.names], 
                           sars.params[colnames(sars.params) %in% measure.names])
measures.combined$pathogen <- ifelse(grepl('ID', measures.combined$ID), 
                                     'Norovirus', 'SARS-CoV-2')

noro.r.range <- seq(min(noro.params.clearpeaks$repRate),
                    max(noro.params.clearpeaks$repRate),
                    length=10)

sars.r.range <- seq(min(sars.params.clearpeaks$repRate),
                    max(sars.params.clearpeaks$repRate),
                    length=10)

cex <- 1.1
cex_labels <- 0.7
panellabelratio <- 0.975

tiff("Pre-symptomatic transmission analysis.tiff", res=500, height=6.5, width=3.5, unit="in")

par(mfrow=c(5, 1), oma=c(0.5, 2, 0.5, 0), mai=c(0.3, 0.5, 0.3, 0.1))

# Panel A: Delay between symptom onset and peak viral shedding 
{
  ylim.ab <- c(floor(min(measures.combined$symptomsToPeakDay, na.rm=T))-1, 
             ceiling(max(measures.combined$symptomsToPeakDay, na.rm=T))+1)
  
  plot(symptomsToPeakDay ~ jitter(ifelse(pathogen == 'Norovirus', 
                                        ifelse(dose == 4.8, .73,
                                               ifelse(dose == 48, 1.067, 
                                                      ifelse(dose == 4800, 1.4, 0))),
                                   2)), 
     data=measures.combined,
     col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     xlim=c(.4, 2.6),
     ylim=ylim.ab, cex.axis=cex, cex=cex,
     pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
                ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                       ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2', 23, 
                              ifelse(symptomsToPeakDay > 0 & pathogen == 'SARS-CoV-2', 5,
                                     9)))),
     bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     yaxt="n",
     xaxt="n",
     ylab = "",
     xlab = "",
     mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  axis(1, at=c(0.73, 1.067, 1.4, 2), 
       labels=c('4.8', '48', '4800', '10'), 
       cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  mtext(text='Norovirus dose\n(RT-PCR units)', side=1, line=2.7, cex=cex_labels,
        adj=-0.1)
  mtext(text = expression(paste('SARS-CoV-2 dose\n(', TCID[50], ')', sep="")), side=1, 
        line=2.7, cex=cex_labels, adj=0.85)
  
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
     las=2)
  
  #Average delays per pathogen
  segments(0.63, noroSummary.dose[noroSummary.dose$dose == 4.8,]$symptomsToPeakDay, 
           0.83, 
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$symptomsToPeakDay, 
           col=noro.color,
           lwd=cex)
  
  segments(0.967, noroSummary.dose[noroSummary.dose$dose == 48,]$symptomsToPeakDay, 
           1.167,
           noroSummary.dose[noroSummary.dose$dose == 48,]$symptomsToPeakDay, 
           col=noro.color,
           lwd=cex)
  
  segments(1.3, noroSummary.dose[noroSummary.dose$dose == 4800,]$symptomsToPeakDay, 
           1.5,
           noroSummary.dose[noroSummary.dose$dose == 4800,]$symptomsToPeakDay, 
           col=noro.color,
           lwd=cex)
  
  segments(1.9, sarsSummary.dose$symptomsToPeakDay, 
           2.1,
           sarsSummary.dose$symptomsToPeakDay, col=sars.color,
           lwd=cex)
  
   polygon(c(0.63, 0.83, 0.83, 0.63), c(noroSummary.dose[noroSummary.dose$dose == 4.8,]
                                       $avgSympToPeakDayLCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4.8,]
                                       $avgSympToPeakDayLCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4.8,]
                                       $avgSympToPeakDayUCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4.8,]
                                       $avgSympToPeakDayUCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(0.967, 1.167, 1.167, 0.967), c(noroSummary.dose[noroSummary.dose$dose == 48,]
                                           $avgSympToPeakDayLCI,
                                           noroSummary.dose[noroSummary.dose$dose == 48,]
                                           $avgSympToPeakDayLCI,
                                           noroSummary.dose[noroSummary.dose$dose == 48,]
                                           $avgSympToPeakDayUCI,
                                           noroSummary.dose[noroSummary.dose$dose == 48,]
                                           $avgSympToPeakDayUCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(1.3, 1.5, 1.5, 1.3), c(noroSummary.dose[noroSummary.dose$dose == 4800,]
                                       $avgSympToPeakDayLCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4800,]
                                       $avgSympToPeakDayLCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4800,]
                                       $avgSympToPeakDayUCI,
                                       noroSummary.dose[noroSummary.dose$dose == 4800,]
                                       $avgSympToPeakDayUCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(1.9, 2.1, 2.1, 1.9), c(sarsSummary.dose$avgSympToPeakDayLCI,
                                       sarsSummary.dose$avgSympToPeakDayLCI,
                                       sarsSummary.dose$avgSympToPeakDayUCI,
                                       sarsSummary.dose$avgSympToPeakDayUCI),
          col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
  
  #Pre-symptomatic vs post-symptomatic line
  abline(h=0, lty=3, lwd=cex)
  
  text(panellabelratio * 2.2 + .4, panellabelratio * (ylim.ab[2] - ylim.ab[1]) + 
       ylim.ab[1], labels="A", cex=cex, font=2)
}

# Panel B: Delay between symptom onset and peak viral shedding vs. viral
# growth rate

{
  xlim.bcd <- c(floor((min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]$repRate) * 2)) / 2,
              ceiling((max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]$repRate) * 2)) / 2)
  
  plot(symptomsToPeakDay ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),], 
   col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
   pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
                ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                       ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2', 23, 
                              ifelse(symptomsToPeakDay > 0 & pathogen == 'SARS-CoV-2', 5,
                                     9)))),
   bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
   cex=cex, cex.axis=cex,
   ylab="", xlab="", xaxt="n", yaxt="n",
   ylim=ylim.ab,
   xlim=xlim.bcd,
   mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  
  bcd.xlabels <- lapply(seq(xlim.bcd[1], xlim.bcd[2], length.out=5), function(x) 
       bquote(10^.(x)))
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5),
     labels=do.call(expression, bcd.xlabels),
     cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
     las=2)
  
  # Lines of best fit
  
  lines(sympToPeakDay ~ t, data=noroSummary.rep, col=noro.color)
  lines(sympToPeakDay ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$sympToPeakDayUCI,
                                                rev(noroSummary.rep$sympToPeakDayLCI)), 
        col = rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$sympToPeakDayUCI,
                                                rev(sarsSummary.rep$sympToPeakDayLCI)), 
        col=rgb(86, 179, 233, alpha=100,
                maxColorValue=255), border = NA)
  
  abline(h=0, lty=3, lwd=cex)

  mtext(text="Day of peak viral shedding \n minus day of symptom onset",
        side=2, line=2.1, cex=cex_labels, adj=-0.5)
  
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1])  + xlim.bcd[1], 
     panellabelratio * (ylim.ab[2] - ylim.ab[1]) + ylim.ab[1], labels="B", 
     cex=cex, font=2)
  
}

# Panel C: Time of peak shedding vs. viral replication rate

{
  ylim.c <- c(floor(min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $peakDay)) - 1, 
              ceiling(max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                            $peakDay)) + 1)
  
  plot(peakDay ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),],
     col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
     xlim=xlim.bcd,
     ylim=ylim.c, cex.axis=cex, cex=cex,
     pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
              ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                     ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2', 23, 
                            ifelse(symptomsToPeakDay > 0 & pathogen == 'SARS-CoV-2', 5,
                                   9)))),
     bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     xaxt="n", yaxt="n", ylab = "", xlab = "",
     mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5), 
     labels=do.call(expression, bcd.xlabels),
     cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.c[1],ylim.c[2], by=2), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
  
  # Lines of best fit
  lines(peakDay ~ t, data=noroSummary.rep, col=noro.color)

  lines(peakDay ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$peakDayUCI,
                                                rev(noroSummary.rep$peakDayLCI)), 
        col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$peakDayUCI,
                                                rev(sarsSummary.rep$peakDayLCI)), 
          col=rgb(86, 179, 233, alpha=100,
                maxColorValue=255), border = NA)
  
  mtext(text="Day of \npeak shedding", side=2, line=2, cex=cex_labels)
  
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
       panellabelratio * (ylim.c[2] - ylim.c[1]) + ylim.c[1], labels="C", 
       cex=cex, font=2)
}
  
# Panel D: Symptom onset vs. viral replication rate

{
  ylim.d <- c(floor(min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $incubation, na.rm=T)) - 1, 
            ceiling(max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $incubation, na.rm=T)) + 1)

  plot(incubation ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),],
       col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       xlim=xlim.bcd, ylim=ylim.d, cex.axis=cex, cex=cex,
       pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
                ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                       ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2', 23, 
                              ifelse(symptomsToPeakDay > 0 & pathogen == 'SARS-CoV-2', 5,
                                     9)))),
       bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       yaxt="n", xaxt="n", ylab = "", xlab = "", bty='l')
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5), 
     labels=do.call(expression, bcd.xlabels), 
     cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  
  axis(2, at=seq(ylim.d[1], ylim.d[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
  
  # Lines of best fit
  lines(symp ~ t, data=noroSummary.rep, col=noro.color)

  lines(symp ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$sympUCI,
                                                rev(noroSummary.rep$sympLCI)), 
          col = rgb(135, 26, 110, alpha=100,
                  maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$sympUCI,
                                                rev(sarsSummary.rep$sympLCI)), 
          col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
  
  mtext(text="Day of \nsymptom onset", side=2, line=2, cex=cex_labels)
  mtext(text=expression(Viral~replication~rate~(hour^-1)),
        side=1, line=2, cex=cex_labels)
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
       panellabelratio * (ylim.d[2] - ylim.d[1]) + ylim.d[1], 
       labels="D", cex=cex, font=2)
}

# Panel E: # Pre-symptomatic transmission potential vs. duration of pre-symptomatic
# shedding

{
  xrange <- c(floor(min(measures.combined[
    !is.na(measures.combined$avgSheddingBeforeSymptoms),]$sheddingTimeBeforeSymptoms,
                na.rm=T)) - 1,
            ceiling(max(measures.combined[
    !is.na(measures.combined$avgSheddingBeforeSymptoms),]$sheddingTimeBeforeSymptoms,
                na.rm=T)) + 1)

  yrange <- c(floor(min(measures.combined$avgSheddingBeforeSymptoms,
                  na.rm=T)) - 1,
              ceiling(max(measures.combined$avgSheddingBeforeSymptoms,
                  na.rm=T)) + 1)
  
  plot(avgSheddingBeforeSymptoms ~ jitter(sheddingTimeBeforeSymptoms), 
       data=measures.combined,
       col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
              ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                     ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2',
                            23, 5))),
       bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       ylim=yrange,
       xlim=xrange,
       cex.axis=cex, cex=cex,
       ylab = "",
       xlab = "",
       xaxt="n",
       yaxt="n", bty='l')
  
  ylabels <- lapply(seq(yrange[1], yrange[2], by=2.5), function(x) 
       bquote(10^.(x)))
  axis(1, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, at=seq(yrange[1], yrange[2], by=2.5),
       labels=do.call(expression, ylabels),
       las=2)
  
  legend(.5 * (xrange[2] - xrange[1]) + xrange[1], 
         1.1 * (yrange[2] - yrange[1]) + yrange[1], 
         legend=c("Pre-symptomatic", "Post-symptomatic"), 
     pch=c(19, 1), col=c("black", "black"), cex=0.9, bty="n")
  
  mtext(text="Average viral shedding \n prior to symptom onset",
        side=2, line=3, cex=cex_labels)
  mtext(text="Duration of shedding prior to symptom onset (days)", side=1, line=1.5, cex=cex_labels)
  
  # Lines of best fit
  noro.presymp.dur.range <- seq(min(noro.params
                              [!is.na(noro.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          max(noro.params
                              [!is.na(noro.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          length=10)
  
  noro.epred <- data.frame(sheddingTimeBeforeSymptoms= noro.presymp.dur.range)
  noro.epred <- transform(noro.epred, y_predicted = predict(transmission.presymp.avg.fit.noro, 
                                                          newdata=noro.epred))
  lines(y_predicted ~ sheddingTimeBeforeSymptoms, data=noro.epred, col=noro.color)
  
  sars.presymp.dur.range <- seq(min(sars.params
                              [!is.na(sars.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          max(sars.params
                              [!is.na(sars.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          length=10)

  sars.epred <- data.frame(sheddingTimeBeforeSymptoms = sars.presymp.dur.range)
  sars.epred <- transform(sars.epred, y_predicted = predict(transmission.presymp.avg.fit.sars, 
                                                           newdata=sars.epred))
  lines(y_predicted ~ sheddingTimeBeforeSymptoms, data=sars.epred, col=sars.color)
  
  text(panellabelratio * (xrange[2] - xrange[1]) + xrange[1], 
       panellabelratio * (yrange[2] - yrange[1]) + yrange[1], 
       labels="E", cex=cex, font=2)
}

dev.off()
```

We again plot delay vs. virus and delay vs. viral replication rate using
the non-conservative measure of pre-symptomatic transmission (shedding onset
minus symptom onset). We also plot shedding and symptom onset against 
replication rate.
```{r, fig.height = 6, fig.width = 3.5}

tiff("Non-conservative pre-symptomatic transmission analysis.tiff", res=500, height=6.5, width=3.5, unit="in")

par(mfrow=c(4, 1), oma=c(1, 2, 0.5, 0), mai=c(0.3, 0.5, 0.3, 0.1))

# Panel A: Delay between symptom onset and shedding onset
{  
  ylim.ab <- c(floor(min(measures.combined$symptomsToStartShedding, na.rm=T))-1, 
               ceiling(max(measures.combined$symptomsToStartShedding, na.rm=T))+1)
  
  plot(symptomsToStartShedding ~ jitter(ifelse(pathogen == 'Norovirus', 
                                        ifelse(dose == 4.8, .73,
                                               ifelse(dose == 48, 1.067, 1.4)),
                                   2), 1.5), 
       data=measures.combined,
       col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       xlim=c(.4, 2.6),
       ylim=ylim.ab, cex.axis=cex, cex=cex,
       pch=ifelse(symptomsToStartShedding < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(symptomsToStartShedding > 0 & pathogen == 'Norovirus', 2,
                         ifelse(symptomsToStartShedding < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(symptomsToStartShedding > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
       bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       yaxt="n",
       xaxt="n",
       ylab = "",
       xlab = "",
       mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
    axis(1, at=c(0.73, 1.067, 1.4, 2), 
       labels=c('4.8', '48', '4800', '10'), 
       cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  mtext(text='Norovirus\n(RT-PCR units)', side=1, line=3, cex=cex_labels,
        adj=-0.1)
  mtext(text = expression(paste('SARS-CoV-2\n(', TCID[50], ')', sep="")), side=1, 
        line=3, cex=cex_labels, adj=1.05)
  
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
       las=2)
  
  #Average delays per pathogen
  segments(0.63, 
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$symptomsToStartShedding, 
           0.83,
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$symptomsToStartShedding, 
           col=noro.color,
           lwd=cex)
  
  segments(0.967, 
           noroSummary.dose[noroSummary.dose$dose == 48,]$symptomsToStartShedding, 
           1.167,
           noroSummary.dose[noroSummary.dose$dose == 48,]$symptomsToStartShedding, 
           col=noro.color,
           lwd=cex)
  
  segments(1.3, 
           noroSummary.dose[noroSummary.dose$dose == 4800,]$symptomsToStartShedding, 
           1.5,
           noroSummary.dose[noroSummary.dose$dose == 4800,]$symptomsToStartShedding, 
           col=noro.color,
           lwd=cex)
  
  segments(1.9, sarsSummary.dose$symptomsToStartShedding, 2.1,
           sarsSummary.dose$symptomsToStartShedding, col=sars.color,
           lwd=cex)

  #Pre-symptomatic vs post-symptomatic line
  abline(h=0, lty=3, lwd=cex)
  
  text(panellabelratio * 2.2 + .4, panellabelratio * (ylim.ab[2] - ylim.ab[1]) + 
         ylim.ab[1], labels="A", cex=cex, font=2)
}
  
  # Panel B: Delay between symptom onset and peak viral shedding vs. viral
  # growth rate
  
{
  xlim.bcd <- c(floor(min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]$repRate) * 2) / 2,
              ceiling(max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]$repRate) * 2) / 2)
    
  plot(symptomsToStartShedding ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),], 
     col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
     pch=ifelse(symptomsToStartShedding < 0 & pathogen == 'Norovirus', 17, 
                  ifelse(symptomsToStartShedding > 0 & pathogen == 'Norovirus', 2,
                         ifelse(symptomsToStartShedding < 0 & pathogen == 'SARS-CoV-2', 23, 
                                ifelse(symptomsToStartShedding > 0 & pathogen == 'SARS-CoV-2', 5,
                                       9)))),
     bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     cex=cex, cex.axis=cex,
     ylab="", xlab="", xaxt="n", yaxt="n",
     ylim=ylim.ab,
     xlim=xlim.bcd,
     mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  
  bcd.xlabels <- lapply(seq(xlim.bcd[1], xlim.bcd[2], length.out=5), function(x) 
         bquote(10^.(x)))
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5),
       labels=do.call(expression, bcd.xlabels),
       cex.axis=cex, 
       mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.ab[1],ylim.ab[2], length.out=5), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04,
       las=2)
  
  # Lines of best fit
  lines(sympToShed ~ t, data=noroSummary.rep, col=noro.color)

  lines(sympToShed ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$sympToShedUCI, 
                                                rev(noroSummary.rep$sympToShedLCI)), 
          col = rgb(135, 26, 110, alpha=100,
                  maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$sympToShedUCI, 
                                                rev(sarsSummary.rep$sympToShedLCI)), 
          col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
  
  abline(h=0, lty=3, lwd=cex)

  mtext(text="Day of shedding onset \n minus day of symptom onset",
        side=2, line=2.1, cex=cex_labels)

  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1])  + xlim.bcd[1], 
       panellabelratio * (ylim.ab[2] - ylim.ab[1]) + ylim.ab[1], labels="B", 
       cex=cex, font=2)
}

# Panel C: Shedding onset vs. viral replication rate

{
  ylim.c <- c(floor(min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $sheddingOnset)) - 1, 
              ceiling(max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                            $sheddingOnset)) + 1)
  
  plot(sheddingOnset ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),],
     col=ifelse(pathogen == 'Norovirus', noro.color, sars.color),
     xlim=xlim.bcd,
     ylim=ylim.c, cex.axis=cex, cex=cex,
     pch=ifelse(symptomsToStartShedding < 0 & pathogen == 'Norovirus', 17, 
              ifelse(symptomsToStartShedding > 0 & pathogen == 'Norovirus', 2,
                     ifelse(symptomsToStartShedding < 0 & pathogen == 'SARS-CoV-2', 23, 
                            ifelse(symptomsToStartShedding > 0 & pathogen == 'SARS-CoV-2', 5,
                                   9)))),
     bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
     xaxt="n", yaxt="n", ylab = "", xlab = "",
     mgp=c(0, 0.5, 0), tck=-0.04, las=2, bty='l')
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5), 
     labels=do.call(expression, bcd.xlabels),
     cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, at=seq(ylim.c[1],ylim.c[2], by=2), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
  
  # Lines of best fit
  lines(shedding ~ t, data=noroSummary.rep, col=noro.color)

  lines(shedding ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$sheddingUCI,
                                                rev(noroSummary.rep$sheddingLCI)), 
        col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$sheddingUCI,
                                                rev(sarsSummary.rep$sheddingLCI)), 
          col=rgb(86, 179, 233, alpha=100,
                maxColorValue=255), border = NA)
  
  mtext(text="Day of \nshedding onset", side=2, line=2, cex=cex_labels)
  
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
       panellabelratio * (ylim.c[2] - ylim.c[1]) + ylim.c[1], labels="C", 
       cex=cex, font=2)
}
  
# Panel D: Symptom onset vs. viral replication rate

{
  ylim.d <- c(floor(min(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $incubation, na.rm=T)) - 1, 
            ceiling(max(measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),]
                        $incubation, na.rm=T)) + 1)

  plot(incubation ~ repRate, data=measures.combined[!measures.combined$ID %in% 
                                              c(noro.unclearpeaks, sars.unclearpeaks),],
       col=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       xlim=xlim.bcd, ylim=ylim.d, cex.axis=cex, cex=cex,
       pch=ifelse(symptomsToPeakDay < 0 & pathogen == 'Norovirus', 17, 
                ifelse(symptomsToPeakDay > 0 & pathogen == 'Norovirus', 2,
                       ifelse(symptomsToPeakDay < 0 & pathogen == 'SARS-CoV-2', 23, 
                              ifelse(symptomsToPeakDay > 0 & pathogen == 'SARS-CoV-2', 5,
                                     9)))),
       bg=ifelse(pathogen == 'Norovirus', noro.color, sars.color), 
       yaxt="n", xaxt="n", ylab = "", xlab = "", bty='l')
  
  axis(1, at=seq(xlim.bcd[1], xlim.bcd[2], length.out=5), 
     labels=do.call(expression, bcd.xlabels), 
     cex.axis=cex, 
     mgp=c(0, 0.5, 0), tck=-0.04)
  
  axis(2, at=seq(ylim.d[1], ylim.d[2], by=3), cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, las=2)
  
  # Lines of best fit
  lines(symp ~ t, data=noroSummary.rep, col=noro.color)

  lines(symp ~ t, data=sarsSummary.rep, col=sars.color)
  
  # Confidence bounds 
  polygon(c(noro.r.range, rev(noro.r.range)), c(noroSummary.rep$sympUCI,
                                                rev(noroSummary.rep$sympLCI)), 
          col = rgb(135, 26, 110, alpha=100,
                  maxColorValue=255), border = NA)
  
  polygon(c(sars.r.range, rev(sars.r.range)), c(sarsSummary.rep$sympUCI,
                                                rev(sarsSummary.rep$sympLCI)), 
          col=rgb(86, 179, 233, alpha=100,
                  maxColorValue=255), border = NA)
  
  mtext(text="Day of \nsymptom onset", side=2, line=2, cex=cex_labels)
  mtext(text=expression(Viral~replication~rate~(hour^-1)),
        side=1, line=2, cex=cex_labels)
  text(panellabelratio * (xlim.bcd[2] - xlim.bcd[1]) + xlim.bcd[1], 
       panellabelratio * (ylim.d[2] - ylim.d[1]) + ylim.d[1], 
       labels="D", cex=cex, font=2)
  
  legend(-0.75, 
         11, 
         legend=c("Pre-symptomatic", "Post-symptomatic", "Unclear"), 
     pch=c(19, 1, 9), col=c("black", "black", "black"), cex=0.9, bty="n")
}

dev.off()
```

Here, we plot average pre-symptomatic shedding factoring in sample weights vs. 
the duration of pre-symptomatic shedding for norovirus.
```{r}
tiff("Average pre-symptomatic shedding (including sample weights) vs. duration of pre-symptomatic period.tiff", res=500, 
     height=3.5, width=3.5, unit="in")

cex_labels <- 0.9
cex <- 0.9
par(oma=c(0.5, 1, 0, 0))
{
  xrange <- c(0,
            ceiling(max(noro.params[!is.na(noro.params$avgTotalSheddingBeforeSymptoms),]$
                          sheddingTimeBeforeSymptoms,
                na.rm=T) * 5) / 5)

  yrange <- c(floor(min(noro.params$avgTotalSheddingBeforeSymptoms,
                  na.rm=T)) - 1,
              ceiling(max(noro.params$avgTotalSheddingBeforeSymptoms,
                  na.rm=T)) + 1)

  plot(avgTotalSheddingBeforeSymptoms ~ jitter(sheddingTimeBeforeSymptoms),
       data=noro.params,
       col=noro.color,
       pch=ifelse(is.na(symptomsToPeakDay)| symptomsToPeakDay == 0, 9, 
                  ifelse(symptomsToPeakDay < 0, 17, 2)),
       ylim=yrange,
       xlim=xrange,
       cex.axis=cex, cex=cex,
       ylab = "",
       xlab = "",
       xaxt="n",
       yaxt="n", bty='l')

  ylabels <- lapply(seq(yrange[1], yrange[2], by=2.5), function(x)
       bquote(10^.(x)))
  axis(1, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04)
  axis(2, cex.axis=cex, mgp=c(0, 0.5, 0), tck=-0.04, at=seq(yrange[1], yrange[2], by=2.5),
       labels=do.call(expression, ylabels),
       las=2)

  mtext(text="Average total viral shedding \n prior to symptom onset",
        side=2, line=3, cex=cex_labels)
  mtext(text="Duration of pre-symptomatic\ntransmission (days)", side=1, line=2.5, cex=cex_labels)

  noro.presymp.dur.range <- seq(min(noro.params
                              [!is.na(noro.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          max(noro.params
                              [!is.na(noro.params$avgSheddingBeforeSymptoms),]
                              $sheddingTimeBeforeSymptoms),
                          length=10)
  # Lines of best fit
  noro.pred <- data.frame(sheddingTimeBeforeSymptoms= noro.presymp.dur.range)
  noro.pred <- transform(noro.pred, y_predicted = predict(transmission.presymp.avg.totalshedding.fit.noro,
                                                          newdata=noro.pred))
  lines(y_predicted ~ sheddingTimeBeforeSymptoms, data=noro.pred, col=noro.color)
}

dev.off()
```

We plot replication rate (p2), time of peak viral load, and time
of symptom onset against dose for the norovirus study participants.
```{r}
tiff("Dose impact pre-symptomatic transmission.tiff", res=500, 
     height=6.5, width=3.5, unit="in")

cex <- 1.3
cex_labels <- 0.9

par(mfrow=c(3, 1), oma=c(4, 3, 0.5, 0), mai=c(0.2, 0.3, 0.2, 0))

# Viral replication rate vs. dose
{
 
  ylim <- c(floor(min(noro.params.clearpeaks$repRate)) - 1,
              ceiling(max(noro.params.clearpeaks$repRate)) + 1)

  ylabels <- lapply(seq(ylim[1], ylim[2], length.out=5), function(x) 
         bquote(10^.(x))) 
  plot(repRate ~ jitter(ifelse(dose == 4.8, .73,
                   ifelse(dose == 48, 1.067, 
                          ifelse(dose == 4800, 1.4, 0)))), 
     data=noro.params.clearpeaks,
     bty='n', ylab="",
     xlab="", col=noro.color,
     xaxt='n', yaxt='n', ylim=ylim, cex=cex,
     xlim=c(0.6, 1.5))
  axis(1, at=c(.73, 1.067, 1.4), labels = c(4.8, 48, 4800), mgp=c(0, 0.75, 0),
       tck=-0.04, cex.axis=cex)
  axis(2, at=seq(ylim[1], ylim[2], length.out=5),
       labels=do.call(expression, ylabels),
       mgp=c(0, 0.75, 0), tck=-0.04, cex.axis=cex, las=2)
  
  segments(.65,
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$repRate, 
           .81, 
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$repRate,
           col=noro.color)
  segments(0.987,
           noroSummary.dose[noroSummary.dose$dose == 48,]$repRate, 
           1.147, 
           noroSummary.dose[noroSummary.dose$dose == 48,]$repRate,
           col=noro.color)
  segments(1.32,
           noroSummary.dose[noroSummary.dose$dose == 4800,]$repRate, 
           1.48, 
           noroSummary.dose[noroSummary.dose$dose == 4800,]$repRate,
           col=noro.color)
  mtext(text=expression(Viral~replication~rate~(hour^-1)), side=2, line=3.5,
        cex=cex_labels)
  
  # Confidence bounds
  polygon(c(.65, .81, .81, .65), c(noroSummary.dose[noroSummary.dose$dose==4.8,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgRepRateLCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgRepRateLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(.987, 1.147, 1.147, .987), c(noroSummary.dose[noroSummary.dose$dose==48,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgRepRateLCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgRepRateLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(1.32, 1.48, 1.48, 1.32), c(noroSummary.dose[noroSummary.dose$dose==4800,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgRepRateUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgRepRateLCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgRepRateLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
}

# Time of peak viral load vs. dose

{
 
  ylim <- c(floor(min(noro.params.clearpeaks$peakDay)) - 1,
              ceiling(max(noro.params.clearpeaks$peakDay)) + 1)

  plot(peakDay ~ jitter(ifelse(dose == 4.8, .73,
                        ifelse(dose == 48, 1.067, 
                          ifelse(dose == 4800, 1.4, 0)))), 
     data=noro.params.clearpeaks,
     bty='n', ylab="",
     xlab="", col=noro.color, yaxt='n',
     xaxt='n', ylim=ylim, cex.axis=cex, cex=cex,
     xlim=c(0.6, 1.5))
  axis(2, las=2, cex.axis=cex)
  axis(1, at=c(.73, 1.067, 1.4), labels = c(4.8, 48, 4800), mgp=c(0, 0.75, 0),
       tck=-0.04, cex.axis=cex)
   segments(.65,
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$peakDay,
           .81, 
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$peakDay,
           col=noro.color)
  segments(.987,
           noroSummary.dose[noroSummary.dose$dose == 48,]$peakDay,
           1.147, 
           noroSummary.dose[noroSummary.dose$dose == 48,]$peakDay,
           col=noro.color)
  segments(1.32,
           noroSummary.dose[noroSummary.dose$dose == 4800,]$peakDay, 
           1.48, 
           noroSummary.dose[noroSummary.dose$dose == 4800,]$peakDay,
           col=noro.color)
  mtext(text="Day of peak shedding", side=2, line=3.5, cex=cex_labels)
  
  # Confidence bounds
  polygon(c(.65, .81, .81, .65), c(noroSummary.dose[noroSummary.dose$dose==4.8,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgPeakDayLCI,
                                   noroSummary.dose[noroSummary.dose$dose==4.8,]$avgPeakDayLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(.987, 1.147, 1.147, .987), c(noroSummary.dose[noroSummary.dose$dose==48,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgPeakDayLCI,
                                   noroSummary.dose[noroSummary.dose$dose==48,]$avgPeakDayLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
  
  polygon(c(1.32, 1.48, 1.48, 1.32), c(noroSummary.dose[noroSummary.dose$dose==4800,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgPeakDayUCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgPeakDayLCI,
                                   noroSummary.dose[noroSummary.dose$dose==4800,]$avgPeakDayLCI),
          col=rgb(135, 26, 110, alpha=100,
                maxColorValue=255), border = NA)
}

# Time of symptom onset vs. dose

{
 
  ylim <- c(0,
              ceiling(max(noro.params.clearpeaks$incubation)) + 1)

  plot(incubation ~ jitter(ifelse(dose == 4.8, .73,
                        ifelse(dose == 48, 1.067, 
                          ifelse(dose == 4800, 1.4, 0)))), 
     data=noro.params,
     bty='n', ylab="",
     xlab="", col=noro.color,
     xaxt='n', ylim=ylim, yaxt='n',
     cex.axis=cex, cex=cex,
     xlim=c(0.6, 1.5))
  axis(2, las=2, cex.axis=cex)
  axis(1, at=c(.73, 1.067, 1.4), labels = c(4.8, 48, 4800), mgp=c(0, 0.75, 0),
       tck=-0.04, cex.axis=cex)
  
     segments(.65,
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$incubation, 
           .81, 
           noroSummary.dose[noroSummary.dose$dose == 4.8,]$incubation,
           col=noro.color)
     segments(.987,
           noroSummary.dose[noroSummary.dose$dose == 48,]$incubation, 
           1.147, 
           noroSummary.dose[noroSummary.dose$dose == 48,]$incubation,
           col=noro.color)
  segments(1.32,
           noroSummary.dose[noroSummary.dose$dose == 4800,]$incubation, 
           1.48, 
           noroSummary.dose[noroSummary.dose$dose == 4800,]$incubation,
           col=noro.color)
  
  mtext(text="Day of symptom onset", side=2, line=3.5, cex=cex_labels)
  mtext(text="Infectious dose\n(norovirus, RT-PCR units)", side=1, line=3.5,
        cex=cex_labels)
  
  text(1.45, 2.5, '*  No confidence intervals \n in this panel', adj=1)
  
}

dev.off()

```

Lastly, for clarification, we plot the differences in incubation periods reported in 
Ge *et al.*, 2023 and Atmar *et al.*, 2008.
```{r}
atmarIDs <- read.csv(here("data", "atmarIDs.csv"))

# Using peak viral load, inoculum dose, and type of symptoms to match up participant IDs

geSymptomData <- noroData$mvs
geSymptomData$type <- ""

# mvsc3 indicates duration of vomiting
# mvsc1 indicates duration of diarrhea
geSymptomData$type[geSymptomData$mvsc3>0] <- "v"
geSymptomData$type[geSymptomData$mvsc3>0 & geSymptomData$mvsc1>0] <- "dv"

# Participant ID4 had reported viral shedding in vomit; we and the authors
# of the study do not know why mvsc3 = 0
geSymptomData[geSymptomData$ID == 'ID4',]$type <- "dv"

geIDs <- do.call(rbind, 
                  lapply(noroParticipants,
                         function(x) data.frame(
                           geID=x,
                           dose=unique(noro.FShedData[noro.FShedData$ID == x,]$dose),
                           peak=round(log10(max(noro.FShedData[noro.FShedData$ID == x,]$y_virus, na.rm=T)), 1),
                           symptom.type=geSymptomData[geSymptomData$ID==x,]$type)))

ID_dict_noro <- merge(atmarIDs, geIDs, by=c("dose", "peak", "symptom.type"))

combinedNoroData <- merge(ID_dict_noro[, c("atmarID", "geID", "incubation")], 
                          noro.params[, c("ID", "incubation", "peakDay")], 
                          by.x = 'geID', by.y='ID')
colnames(combinedNoroData) <- c('geID', 'atmarID', 'atmarIncubation', 'geIncubation',
                                'peakDay')

# Manually adding in ID10/720 because their trajectories are identical, but
# their symptoms do not match up between the two datasets
combinedNoroData <- rbind(combinedNoroData, 
                          data.frame(geID = 'ID10', atmarID='720', 
                                     geIncubation = noro.params[
                                       noro.params$ID == 'ID10',]$incubation,
                                     atmarIncubation = atmarIDs[
                                       (atmarIDs$atmarID == '720') & 
                                         (!is.na(atmarIDs$atmarID)),]$incubation,
                                     peakDay = noro.params[
                                       noro.params$ID == 'ID10',]$peakDay))

# Manual ordering of IDs
combinedNoroData <- combinedNoroData[c(1, 7, 8, 9, 10, 11, 12, 13, 14, 2, 3, 4, 5, 6),]



atmarColor <- "#f1a340"
geColor <- "#871a6e"

tiff(filename="Symptom Onset Discrepancy.tiff", height=3.5, width=5, units = "in",
    res=500)

{par(mar=c(4.1, 3.6, 0.5, 0.1), oma=c(2, 0, 0, 0))
plot(1:nrow(combinedNoroData), combinedNoroData$atmarIncubation, xaxt="n",
     ylim=c(0, 5), col=atmarColor, xlab="", ylab="",
     bty="n")
axis(1, at=1:nrow(combinedNoroData), 
     labels=combinedNoroData$geID,las=2, col.axis=geColor)
axis(1, at=1:(nrow(combinedNoroData)), 
     labels=paste(combinedNoroData$atmarID, ' /        ', sep=''), las=2, 
     col.axis=atmarColor)
mtext("Participant ID", side=1, line=5)
mtext("Day", side=2, line=2.5)
legend(4, 5, pch=c(1, 1, 16), col=c(atmarColor, geColor, 'black'), 
       legend=c(bquote("Atmar" ~ italic("et al.") ~ 2008 ~ "\u2014 symptom onset"), 
                bquote("Ge" ~ italic("et al.") ~ 2023 ~ "\u2014 symptom onset"),
                "Peak viral shedding"))
points(1:nrow(combinedNoroData), combinedNoroData$geIncubation, col=geColor)
points(1:nrow(combinedNoroData), combinedNoroData$peakDay, pch=16)
}

dev.off()
```
