---
title: "Mechanistic Model Usage"
author: "Kayla Zhang, Damie Pak, Megan Greischar"
date: "2025-05-24"
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Our packages for ODE modeling and plotting:
```{r}
require(deSolve)
require(ggplot2)
require(gridExtra)
require(grid)
```

Here, we have our ODE model, based on the within-host model from King *et al*., 
2009, as well as a modified model implementing a pathogen load carrying capacity.
```{r}
within.host.model <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}

within.host.model.cap <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P*(1 - P/(5*10^5)) - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}
```

Below, we define some useful functions for setting parameters and creating 
dataframes containing the simulated data from our model. We focus on varying 
pathogen replication rate in the present study, so we maintain constant values 
for the remaining parameters based on values used in King *et al*., 2009 (except
for immune cell kill rate of the pathogen, which varies based on the within-host 
model to more easily show both pre- and post-symptomatic transmission later on). 
We truncate the dataframes when pathogen load reaches 0, assuming pathogen clearance. 

```{r}
get.params <- function(k, r) {
  return(c(y=10^-4, k=k, r=r, a=1, d=.5))
}

desolve_function <- function(tstep, k, r, model, id) {
  time <- seq(0,2, by=tstep)
  params <- get.params(k, r)

  df <- data.frame(lsoda(c(P=id, X=0), time, model, 
                       params))
  
  if(any(df$P <= 0)){
    clearance <- min(which(df$P <= 0))
    df[clearance:length(time),] <- NA
  }
  
  df <- data.frame(time=time, 
                  pathogen=df$P, 
                  immune=df$X, 
                  k=k, r=r, dose=id)
  
  return(df)
}

```

Here we provide functions for calculating time of peak pathogen load, 
symptom onset, and the end of symptoms, as well as the delay between symptom 
onset and peak pathogen load. If time of peak pathogen load comes before time of 
symptom onset, meaning the calculated delay is negative, then by our definition, 
there is pre-symptomatic transmission. 
```{r}
peak.pathogenloadtime <- function(df) {
  peak.pl <- max(df$pathogen, na.rm=TRUE)
  return(df[which(df$pathogen==peak.pl),]$time)
}

symptom.onset <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(min(symptomatic))
  } else {
    return(NA)
  }
}

symptom.end <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(max(symptomatic))
  } else {
    return(NA)
  }
}

delay.calculation <- function(df, symp.thresh) {
  return(peak.pathogenloadtime(df) - symptom.onset(df, 'immune', symp.thresh))
}

```

The first model we examine is one in which symptoms are dependent only on 
pathogen load. This is represented by a dotted horizontal red line depicting the 
threshold for symptoms; if pathogen load falls above that threshold, then symptoms 
are present. We have defined pre-symptomatic transmission as reaching peak pathogen 
load before symptom onset. This is a conservative method of estimating 
pre-symptomatic transmission. 

Given this definition, there are two potential outcomes: 1) symptom
threshold is below peak pathogen load, signifying post-symptomatic transmission, 
and 2) symptom threshold is above peak pathogen load, signifying 
asymptomatic infection. The outlier case (symptoms threshold is at peak pathogen 
load) is ignored. 

The second model we examine is one in which symptoms are
dependent only on immune response. When immune response rises above a certain 
threshold, the patient is symptomatic.

We produce simulations of these two models to demonstrate their dynamics.
```{r, fig.height = 7.5, fig.width = 6.5}
plot.model.simulations <- function(symp.thresh, data, 
                                    cex, label,
                                    xlimlow, xlimhigh, s.compartment) {
  
  col <- 'red'
  
  time <- data$time
  path.load <- data$pathogen
  immune <- data$immune
  peak.load.time <- peak.pathogenloadtime(data)
  
  if (s.compartment == 'pathogen') {
    plot(time, path.load, col=col, xlab="", ylab="", xlim=c(xlimlow, xlimhigh), log='y',
         ylim=10^c(-1, 7), cex.lab=cex, cex.axis=cex, mgp=c(0, 1, 0), 
         yaxt='n', tck=-0.04, type="l", lwd=3, bty='l')
    lines(time, immune, col=gray(0.25), type="l", lty=2)
    points(peak.load.time, path.load[which(time==peak.load.time,)],
         cex=1.5, pch=16, col="red")
    points(peak.load.time, immune[which(time==peak.load.time,)],
         col=gray(0.25), cex=1.5, pch = 21, bg = "white")
    axis(2, cex.axis=cex, at=10^(seq(-1, 7, by=2)), 
     labels=c(expression(10^-1), 
              expression(10^1),
              expression(10^3),
              expression(10^5),
              expression(10^7)), las=2, mgp=c(0, 1, 0))
  } else {
    plot(time, immune, col=col, xlab="", ylab="", xlim=c(xlimlow, xlimhigh), log='y',
         ylim=10^c(-1,7), cex.lab=cex, cex.axis=cex, mgp=c(0, 1, 0), 
         yaxt='n', lty = 2, tck = -0.04, type="l", lwd=3, bty='l')
    lines(time, path.load, col=gray(0.25), type="l")
    points(peak.load.time, immune[which(time==peak.load.time,)],
         cex=1.5, pch = 21, col = "red", bg = "white")
    points(peak.load.time, path.load[which(time==peak.load.time,)],
         col=gray(0.25), pch=16, cex=1.5)
    axis(2, cex.axis=cex, at=10^(seq(-1, 7, by=2)), 
     labels=c(expression(10^-1), 
              expression(10^1),
              expression(10^3),
              expression(10^5),
              expression(10^7)), las=2, mgp=c(0, 1, 0))
  }
  
  abline(h=symp.thresh, lty=3, col=col, lwd=2)
  
  polygon(x=c(symptom.onset(data, 
                            s.compartment,
                            symp.thresh), 
              symptom.onset(data, 
                            s.compartment,
                            symp.thresh),
              symptom.end(data,
                          s.compartment,
                          symp.thresh),
              symptom.end(data,
                          s.compartment,
                          symp.thresh)), 
          y=10^c(-5, 10, 10, -5), 
          col=rgb(200, 200, 200, alpha=100,
                  maxColorValue=255),
          border = NA)
  
  text(0.01 * (xlimhigh - xlimlow) + xlimlow, 10^(.85 * (8)) + 10^-1, 
       labels=label, cex=cex, adj=0)
}

data.lowreplication.nocap <- desolve_function(tstep=10^-4, 
                            k=3.5, r=30, model=within.host.model, id=10^4)
data.highreplication.nocap <- desolve_function(tstep=10^-4, 
                             k=3.5, r=200, model=within.host.model, id=10^4) 

data.lowreplication.cap <- desolve_function(tstep=10^-4, 
                            k=3.5, r=30, model=within.host.model.cap, id=10^4)
data.highreplication.cap <- desolve_function(tstep=10^-4, 
                             k=3.5, r=200, model=within.host.model.cap, id=10^4) 

path.symp.thresh <- 10^6
immune.sympthresh <- 10
immune.pathcap.sympthresh <- 2
cex <- 1.25
cex_labels <- 0.9

{
  par(mfrow=c(3, 2), oma=c(0, 2.5, 1.75, 0), mai=c(0.52, 0.32, 0.32, 0.42))
    
  plot.model.simulations(symp.thresh = path.symp.thresh,
                         data = data.lowreplication.nocap, 
                         cex = cex, 
                         label = "Asymptomatic", 
                         xlimlow = 0, xlimhigh = 0.3, 
                         s.compartment = 'pathogen')
  
  legend(-0.01, 10^4.2, legend=c('Peak pathogen load', 
                             'No symp/symp',
                             'Symptom threshold',
                             'Immune effector', 
                             'Pathogen'
                           ), 
         lty=c(NA, NA, 3, 2, 1), pch=c(16, NA, NA, NA, NA), bty="n",
         xpd=TRUE, cex=0.9)
  
  mtext(text="SLOW REPLICATION", font=2, side=3, line=2.5, cex=cex_labels)
  
  mtext(text="A. Pathogen abundance symptom threshold", side=3, line=1, 
        cex=cex_labels, adj=0.4)
  
  plot.model.simulations(symp.thresh = path.symp.thresh, 
                         data = data.highreplication.nocap, 
                         cex = cex, 
                         label = "Post-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.05, 
                         s.compartment = 'pathogen')
  
  mtext(text="FAST REPLICATION", font=2, side=3, line=2.5, cex=cex_labels)
  
  plot.model.simulations(symp.thresh = immune.sympthresh, 
                         data = data.lowreplication.nocap, 
                         cex = cex, 
                         label = "Pre-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.3, 
                         s.compartment = 'immune')
  
  mtext(text="B. Immune effector abundance symptom threshold, no pathogen carrying capacity", 
        side=3, line=1, cex=cex_labels, adj=0.1)
  
  plot.model.simulations(symp.thresh = immune.sympthresh, 
                         data = data.highreplication.nocap, 
                         cex = cex, 
                         label = "Post-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.05, 
                         s.compartment = 'immune')
  
  plot.model.simulations(symp.thresh = immune.pathcap.sympthresh, 
                         data = data.lowreplication.cap, 
                         cex = cex, 
                         label = "Post-symptomatic",
                         xlimlow = 0, xlimhigh = 0.3, 
                         s.compartment = 'immune')
  
  mtext(text="C. Immune effector abundance symptom threshold, pathogen carrying capacity", 
        side=3, line=1, cex=cex_labels, adj=0.105)
  
  plot.model.simulations(symp.thresh = immune.pathcap.sympthresh, 
                         data = data.highreplication.cap, 
                         cex = cex, 
                         label = "Pre-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.12, 
                         s.compartment = 'immune')
  
  mtext("Time", side=1, line=-1.75, outer=TRUE, adj=0.5)
  mtext("Abundance", side=2, line=1, outer=TRUE)
}
```

Now, we vary pathogen replication rate (r) and plot the delay between
symptom onset and peak pathogen load against various r values for both 
transmission models (with and without pathogen load cap) to
investigate the correlation between replication rate and
pre-symptomatic transmission. 
```{r, fig.height = 3.5, fig.width = 3.5}
get.delays.df <- function(data, param.vals, symp.thresh) {
  
  delays.df <- do.call(rbind, lapply(param.vals, function(x) 
    data.frame(r=x,
               delay.symponset.peakload=
                 delay.calculation(data[data$r==x,], symp.thresh),
               symp.thresh=symp.thresh))) 
  
  return(delays.df)
}

r_vals <- seq(10, 250, by=10)

varyingr.nocap <- data.frame(k= 3.5, r=r_vals, dose=10^5)
varyingr.nocap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  k=varyingr.nocap$k, 
                                  r=varyingr.nocap$r, 
                                  id=varyingr.nocap$dose,                                  
                                  MoreArgs=list(model=within.host.model),
                                  SIMPLIFY = FALSE))

varyingr.cap <- data.frame(k= 7, r=r_vals, dose=10^5)
varyingr.cap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  k=varyingr.cap$k, 
                                  r=varyingr.cap$r, 
                                  id=varyingr.cap$dose,                                  
                                  MoreArgs=list(model=within.host.model.cap),
                                  SIMPLIFY = FALSE))

delays.byr.lowsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=5)

delays.byr.medsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=10)

delays.byr.highsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=15)

delays.byr.lowsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=1)

delays.byr.medsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=2)

delays.byr.highsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=3)

plot.varyingr <- function(delaysbyr1, delaysbyr2, delaysbyr3, size, ylow, yhigh,
                          xlow, xhigh) {
  highlighted_points <- delaysbyr2[delaysbyr2$r==30 | delaysbyr2$r==200,]
  
  return(ggplot(delaysbyr1, aes(x=r, y=delay.symponset.peakload, 
                         color = as.character(symp.thresh))) + theme_classic() +
    theme(
          plot.title = element_text(hjust = 0.5, vjust = 3, size=size,
                                    face="bold"), 
          plot.margin = unit(c(0.05, 0, 0, -0.05),"in"),
          axis.title.y = element_text(margin = margin(r=4), size=size),
          axis.title.x = element_text(margin = margin(t=8), size=size),
          axis.text.x = element_text(size=size, margin = margin(t=2)),
          axis.text.y = element_text(size=size, margin = margin(r=2)),
          axis.ticks.length=unit(0.05, "in"),
          legend.key.size = unit(0.001, "in"),
          legend.text = element_text(size=size * .9),
          legend.title = element_text(size=size * .9),
          legend.position = c(0.85, 0.25)) +
      ylim(ylow, yhigh) + 
    xlim(xlow, xhigh) +
  geom_point(size=1) + 
    geom_line() + geom_point(data=delaysbyr2, size=1) + 
    geom_line(data=delaysbyr2) + geom_point(data=delaysbyr3, size=1) +
    geom_line(data=delaysbyr3) +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
    geom_point(data=highlighted_points, aes(x=r, y=delay.symponset.peakload), color="red", size=1)) 
}

tiff("Delays varying r.tiff", res=500, 
     height=3.5, width=3.5, unit="in")

cols <- colorRampPalette(c("deepskyblue", "deepskyblue4"))
cols <- cols(3)

plot.varyingr.nocap <- plot.varyingr(delays.byr.lowsympthresh.nocap, 
              delays.byr.medsympthresh.nocap, 
              delays.byr.highsympthresh.nocap, 
              size=6, 
              ylow=-.02, 
              yhigh=.005, 
              xlow=0, 
              xhigh=350) + 
  ggtitle("A) No Carrying Capacity") + 
  xlab("") +
  ylab("") +
  geom_text(aes(x=315, y=-0.0013, label = "Pre-symptomatic"),
                color="black", size=2) +
  geom_text(aes(x=315, y=0.0017, label = "Post-symptomatic"),
                color="black", size=2) +
  scale_color_manual(name="Symptom Threshold", values = c("5"= cols[1], 
                                                          "10"= cols[2], 
                                                          "15"= cols[3]),
                     limits=c("5", "10", "15"))
  

plot.varyingr.cap <- plot.varyingr(delays.byr.lowsympthresh.cap,
              delays.byr.medsympthresh.cap, 
              delays.byr.highsympthresh.cap, 
              size=6, 
              ylow=-.006, 
              yhigh=.006, 
              xlow=0, 
              xhigh=350) + 
  ggtitle("B) Pathogen Load Carrying Capacity") +
  xlab("Replication rate (r)") + 
  ylab("") +
  geom_text(aes(x=315, y=-0.00065, label = "Pre-symptomatic"),
                color="black", size=2) +
  geom_text(aes(x=315, y=0.0008, label = "Post-symptomatic"),
                color="black", size=2) + 
  scale_color_manual(name="Symptom Threshold", values = c("1"= cols[1], 
                                                          "2"= cols[2], 
                                                          "3"= cols[3]))

grid.arrange(plot.varyingr.nocap, plot.varyingr.cap, nrow=2, 
             left=textGrob("Day of peak viral load \n minus day of symptom onset",
             gp = gpar(fontsize = 6),
             rot=90))
dev.off()
```

We show that for the range of replication rates used, immune response peaks after 
pathogen load.
```{r, fig.height = 3.5, fig.width = 6.5}
peak.immunetime <- function(df) {
  peak.immune <- max(df$immune, na.rm=TRUE)
  return(df[which(df$immune==peak.immune),]$time)
}

get.peaktimes.df <- function(data) {
  
  peaktimes.df <- do.call(rbind, lapply(r_vals, function(x) 
    data.frame(r=x,
               pathogen.peaktime=peak.pathogenloadtime(data[data$r==x,]),
               immune.peaktime=peak.immunetime(data[data$r==x,])))) 
  
  return(peaktimes.df)
}

peaktimes.cap <- get.peaktimes.df(varyingr.cap)
peaktimes.nocap <- get.peaktimes.df(varyingr.nocap)

{
  par(mfrow=c(1, 2), mai=c(0.82, 0.72, 0.52, 0.06), oma=c(0, 1, 0, 0))
  
  plot(pathogen.peaktime ~ r, data=peaktimes.cap, ylim=c(0, 0.15), type="l",
       xlab="", ylab="", yaxt="n", bty="l", 
       mgp=c(2.25, 0.75, 0))
  axis(2, las=2)
  mtext("Time of max abundance", 2, line=3)
  lines(immune.peaktime ~ r, data=peaktimes.cap, col="red", lty=2)
  
  mtext("Exponential growth", side=3, line=1.5)
  
  mtext("Replication rate (r)", side=1, line=2.5, adj=2)
  
  plot(pathogen.peaktime ~ r, data=peaktimes.nocap, ylim=c(0, 0.15), type="l",
       xlab="", ylab="", yaxt="n", bty="l")
  axis(2, las=2)
  lines(immune.peaktime ~ r, data=peaktimes.nocap, col="red", lty=2)
  
  mtext("Pathogen carrying capacity", side=3, line=1.5)
  
  legend(50, 0.11, legend=c('Immune effector',
                             'Pathogen'), 
         lty=c(2, 1), col=c("red", "black"), bty="n")
}
```