---
title: "Supplemental Code"
author: "Kayla Zhang, Damie Pak, Megan Greischar"
date: "2024-05-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

We use the R package metaDigitise to extract data points from figure 1C-E in
Atmar *et al*., 2008.

```{r, echo = TRUE, eval = FALSE}
require(metaDigitise)
 
atmardata <- metaDigitise("Process Images", summary=FALSE)
atmardata <- rbind(atmardata$scatterplot$atmarplotc.jpg, 
                   rbind(atmardata$scatterplot$atmarplotd.jpg,
                         atmardata$scatterplot$atmarplote.jpg))
```

We then fit a statistical model (Li & Handel, 2014, Holder & Beauchemin, 2011) 
that estimates viral load values over time to the shedding data we just 
extracted using the least squares method. There is a typo in 
Li & Handel, 2014 in the denominator, where the term 
exp(p4 * (t-p3)) is written as exp(-p4 * (t-p3)) (Holder & Beauchemin, 2011). 
We make the assumption that viral load is a 1-to-1 mapping to viral shedding. 
We also subtract 19 hours each time we convert from days to hours to set time 
point = 0 as time of inoculation, since Atmar et al., 2008 defined study day 1 
as beginning around 5-6 hours after inoculation.

```{r}
generate.estimates <- function(t, pars) {
  
  p1 <- exp(pars[1])
  p2 <- exp(pars[2])
  p3 <- exp(pars[3])
  p4 <- exp(pars[4])
  
  estimate <- (2 * p1) / (exp(-p2*(t-p3)) + exp(p4*(t-p3)))
  
  return(log10(estimate))
  
} 

sse.viralLoad <- function(pars, data) {
  vl <- log10(data$y)
  t <- data$x * 24 - 19
  
  estimates <- generate.estimates(t, pars)
  
  sse <- sum((estimates - vl)^2)
  
  if (sse > 1e20) {
    return(1e20)
  } else {
    return(sse)
  }
  
}

generate.possible.params <- function(data) {
  numguess <- 1e4
  
  fits <- data.frame(matrix(NA, nrow = numguess, ncol = 6))
  
  colnames(fits) <- c('p1','p2','p3','p4','convergence','sse')
  
  # These are ln(param) values
  for (g in 1:numguess) {
    p1 <- runif(1, -1, log(1e12))
    p2 <- runif(1, -1, 1)
    p3 <- runif(1, -1, log(4*24))
    p4 <- runif(1, -5, 1)
    
    params <- c(p1, p2, p3, p4)
    
    fit <- optim(params, sse.viralLoad, data=data)
    
    fits[g, ] <- c(fit$par, fit$convergence, fit$value)
  }
  
  return(fits)
}

get.params <- function(data) {
  
  fits <- generate.possible.params(data)
  
  minsse <- min(fits[fits$convergence==0, ]$sse)
  
  pars <- unlist((fits[fits$sse == minsse, ])[1:4])
  
  return (c(pars, minsse))
}

```

We define two endpoints for the fit: 1) stopping at the first local minimum
in viral load after the peak, and 2) stopping at the first viral load value 
after the peak that is equal to or less than the inoculum dose/initial viral 
load.

```{r}
make.param.df <- function(data, maxt) {
  
  participants <- unique(data$participant)
  
  participantparams <- matrix(nrow=length(participants), ncol=5)
  
  for (i in 1:length(participants)) {
    pars <- get.params(data[data$participant==participants[i],]
                         [1:(maxt[maxt$participant==participants[i],]$max.point),])
    
    participantparams[i,] <- c(participants[i], pars[1:4])
  }
  
  participantparams <- as.data.frame(participantparams)
  colnames(participantparams) <- c("participant", "p1", "p2",
                               "p3", "p4")
  
  return(participantparams)
}

atmardata <- read.csv("atmardata.csv")
atmarsymptoms <- read.csv('atmarsymptoms.csv')
participants <- unique(atmardata$participant)

# Method 1: Stopping at the first local minimum
maxt.firstmin <- data.frame(participant=participants, 
                            max.point=rep(NA, length(participants)))

for (p in participants) {
  
  vldata <- atmardata[atmardata$participant==p,]
  
  t.peak <- which(vldata$y == max(vldata$y))
  
  points.after.peak <- vldata$y[-1:-t.peak]
  
  next.points <- append(vldata$y[-1:-(t.peak+1)], NA)
  
  local.min <- which(points.after.peak <= next.points)[1] + t.peak
  
  if (is.na(local.min)) {
    
    maxt.firstmin[maxt.firstmin$participant==p,]$max.point <- length(vldata$y)
    
  } else {
    
    maxt.firstmin[maxt.firstmin$participant==p,]$max.point <- local.min
    
  }
}

# Method 2: Stopping when viral load gets back down to the initial viral load
maxt.backtoinitial <- data.frame(participant=participants, 
                                 max.point=rep(NA, length(participants)))

for (p in participants) {
  vldata <- atmardata[atmardata$participant==p,]
  
  t.peak <- which(vldata$y == max(vldata$y))
  
  points.after.peak <- vldata$y[-1:-t.peak]
  
  maxt.backtoinitial[maxt.backtoinitial$participant==p,]$max.point <-  
    which(points.after.peak <= vldata$y[1])[1] + t.peak
}
```

Note that because this model is statistical, not deterministic, the parameters
will fluctuate with each run.

```{r, eval = FALSE}
params.localmin <- make.param.df(atmardata, maxt.firstmin)

params.backtoinitial <- make.param.df(atmardata, maxt.backtoinitial)
```

As a result, we saved the results of our first iteration in a csv file.

```{r}
params.localmin <- read.csv("Atmar Parameters local minimum.csv")
params.backtoinitial <- read.csv("Atmar Parameters back to initial.csv")
```

Because participants 723 and 731 do not have clear peaks in pathogen load,
we exclude them from any future calculations using our conservative measure
of pre-symptomatic transmission, which is that peak pathogen load is reached
before symptom onset. We proceed to calculate the delay between symptom onset and peak pathogen load for
each participant in the Atmar *et al*., 2008 study, barring the excluded ones.

```{r}
participants.conservative <- participants[participants != 723 & 
                                            participants != 731]

delays.conservative <- data.frame("participant"=participants.conservative, 
                    "delay"=rep(0, length(participants.conservative)), 
                    "dose"=rep(0, length(participants.conservative)))

for (p in participants.conservative) {
  
  vldata <- atmardata[atmardata$participant == p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  dose <- unique(vldata$dose)
  
  params <- unlist(params.localmin[params.localmin$participant==p, 2:5])

  t <- seq(min(vldata$x)*24-19, 
        vldata[maxt.firstmin[maxt.firstmin$participant==p,]$max.point,]$x*24-19, 
        by=1)
  
  predictedvl <- generate.estimates(t, params)
  
  peak.vl.time <- t[which(predictedvl==max(predictedvl))]
  
  participantdelay <- peak.vl.time - (sympdata$symp.onset * 24 - 19)
  
  delays.conservative[delays.conservative$participant==p, ]$delay <- participantdelay
   
  delays.conservative[delays.conservative$participant==p, ]$dose <- dose

}
```

We run a linear regression between the delay between symptom onset and peak
pathogen load and each of the 4 parameters from the statistical model as well
as between viral growth and decay rate. Only p2 and p4, which represent pathogen 
replication rate and time of peak pathogen load, respectively, were found to 
have significant correlations with delay. They are also significantly correlated 
 to each other.
 
```{r}
params.and.delays.conservative <- merge(params.localmin, delays.conservative)

lm.fit.p1.delay.conservative <- 
  lm(delay/24 ~ p1, data=params.and.delays.conservative)
p1.delay.conservative.pval <- 
  summary(lm.fit.p1.delay.conservative)$coefficients[2,4]
p1.delay.conservative.pval

lm.fit.p2.delay.conservative <- 
  lm(delay/24 ~ p2, data=params.and.delays.conservative)
p2.delay.conservative.pval <- 
  summary(lm.fit.p2.delay.conservative)$coefficients[2,4]
p2.delay.conservative.pval

lm.fit.p3.delay.conservative <- 
  lm(delay/24 ~ p3, data=params.and.delays.conservative)
p3.delay.conservative.pval <- 
  summary(lm.fit.p3.delay.conservative)$coefficients[2,4]
p3.delay.conservative.pval

lm.fit.p4.delay.conservative <- 
  lm(delay/24 ~ p4, data=params.and.delays.conservative)
p4.delay.conservative.pval <- 
  summary(lm.fit.p4.delay.conservative)$coefficients[2,4]
p4.delay.conservative.pval

# Viral growth vs. decay rate
lm.fit.p4.p2.conservative <- 
  lm(p4 ~ p2, data=params.and.delays.conservative)
p4.p2.conservative.pval <- 
  summary(lm.fit.p4.p2.conservative)$coefficients[2,4]
p4.p2.conservative.pval

# Test for collinearity
lm.fit.p2.p3.conservative <-
  lm(p2 ~ p3, data=params.and.delays.conservative)
p2.p3.conservative.pval <-
  summary(lm.fit.p2.p3.conservative)$coefficients[2,4]
p2.p3.conservative.pval
```

Using a non-conservative measure of pre-symptomatic transmission, ie shedding
onset values as reported in the study, we calculate the delay between
shedding onset and symptom onset (the length of the latency period). For 
simplicity's sake, a negative delay still indicates pre-symptomatic transmission.

```{r}
atmarshedding <- read.csv("atmarshedding.csv")

delays.nonconservative <- data.frame("participant"=participants, 
                    "delay"=rep(0, length(participants)), 
                    "dose"=rep(0, length(participants)))

for (p in participants) {
  
  sheddata <- atmarshedding[atmarshedding$participant == p, ]
  sympdata <- atmarsymptoms[atmarsymptoms$participant == p, ]
  dose <- unique(sheddata$dose)
  
  participantdelay <- sheddata$shed.onset - sympdata$symp.onset
  
  delays.nonconservative[delays.nonconservative$participant==p, ]$delay <- participantdelay
   
  delays.nonconservative[delays.nonconservative$participant==p, ]$dose <- dose

}
```

As before, we test for any significant correlations between the statistical 
model parameters and the duration of the latency period as well as
between viral growth and decay rate. Just as with the
conservative measure of pre-symptomatic transmission, there is a significant
correlation between both viral replication rate and time of peak pathogen load
with the duration of the latency period. We test for collinearity with
another linear regression, which returns positive.

```{r}
params.and.delays.nonconservative <- merge(params.localmin, delays.nonconservative)

lm.fit.p1.delay.nonconservative <- 
  lm(delay ~ p1, data=params.and.delays.nonconservative)
p1.delay.nonconservative.pval <- 
  summary(lm.fit.p1.delay.nonconservative)$coefficients[2,4]
p1.delay.nonconservative.pval

lm.fit.p2.delay.nonconservative <- 
  lm(delay ~ p2, data=params.and.delays.nonconservative)
p2.delay.nonconservative.pval <- 
  summary(lm.fit.p2.delay.nonconservative)$coefficients[2,4]
p2.delay.nonconservative.pval

lm.fit.p3.delay.nonconservative <- 
  lm(delay ~ p3, data=params.and.delays.nonconservative)
p3.delay.nonconservative.pval <- 
  summary(lm.fit.p3.delay.nonconservative)$coefficients[2,4]
p3.delay.nonconservative.pval

lm.fit.p4.delay.nonconservative <- 
  lm(delay ~ p4, data=params.and.delays.nonconservative)
p4.delay.nonconservative.pval <- 
  summary(lm.fit.p4.delay.nonconservative)$coefficients[2,4]
p4.delay.nonconservative.pval

# Viral growth vs. decay rate
lm.fit.p4.p2.nonconservative <- 
  lm(p4 ~ p2, data=params.and.delays.nonconservative)
p4.p2.nonconservative.pval <- 
  summary(lm.fit.p4.p2.nonconservative)$coefficients[2,4]
p4.p2.nonconservative.pval

# Test for collinearity
lm.fit.p2.p3.nonconservative <-
  lm(p2 ~ p3, data=params.and.delays.nonconservative)
p2.p3.nonconservative.pval <-
  summary(lm.fit.p2.p3.nonconservative)$coefficients[2,4]
p2.p3.nonconservative.pval
```

We also run a linear regression between pre-symptomatic transmission potential,
measured as average shedding prior to symptoms, and time of symptom onset. 
There is no correlation.

```{r}
transmission.latent <- data.frame("participant"=participants.conservative, 
                                  "dose"=rep(0, length(participants.conservative)),
                    "transmission_potential"=rep(0, length(participants.conservative)), 
                    "symptom_onset"=rep(0, length(participants.conservative)))

for (i in 1:length(participants.conservative)) {
  p <- participants.conservative[i]
  
  transmission.latent[i,]$participant <- p
  
  symp.onset <- atmarsymptoms[atmarsymptoms$participant == p,]$symp.onset
  
  transmission.latent[i,]$symptom_onset <- symp.onset
  
  params <- unlist(params.localmin[params.localmin$participant == p,][2:5])
  
  vldata <- atmardata[atmardata$participant == p,]
  
  transmission.latent[i,]$dose <- unique(vldata$dose)
  
  shed.onset <- min(vldata$x) * 24 - 19
  
  if(shed.onset > symp.onset * 24 - 19) {
    avg.shedding.latent <- 0
  } else {
    t <- seq(min(vldata$x)*24-19, symp.onset*24-19, by=1)
    vl.latent <- 10^generate.estimates(t, params)
    avg.shedding.latent <- log10(mean(vl.latent))
  }
  
  transmission.latent[i,]$transmission_potential <- avg.shedding.latent
}

transmission.latent.fit <- lm(transmission_potential ~ symptom_onset, 
                              data = transmission.latent)
transmission.latent.pval <- 
  summary(transmission.latent.fit)$coefficients[2,4]

transmission.latent.pval
```

Lastly, we break down the correlation between pre-symptomatic transmission
and viral growth rate by examining the relationships between time of
peak shedding and viral growth rate as well as between symptom onset
and viral growth rate. There is a significant correlation between 
viral growth rate and time of peak shedding, but not between 
viral growth rate and time of symptom onset.

```{r, fig.height = 3, fig.width = 6.5}

participant.peaktimes <- data.frame("participant"=participants.conservative, 
                    "peaktime"=rep(0, length(participants.conservative)), 
                    "dose"=rep(0, length(participants.conservative)))

for (p in participants.conservative) {
  
  vldata <- atmardata[atmardata$participant == p, ]
  
  t <- seq(min(vldata$x)*24-19, 
        vldata[maxt.firstmin[maxt.firstmin$participant==p,]$max.point,]$x*24-19, 
        by=1)
  
  params <- unlist(params.localmin[params.localmin$participant==p, 2:5])
  
  predictedvl <- generate.estimates(t, params)
  
  participant.peaktime <- t[which(predictedvl==max(predictedvl))]
  
  dose <- unique(vldata$dose)
  
  participant.peaktimes[participant.peaktimes$participant==p, ]$dose <- dose
  
  participant.peaktimes[participant.peaktimes$participant==p, ]$peaktime <- 
    participant.peaktime
}

participant.params <- params.localmin[params.localmin$participant != 723 & 
                                        params.localmin$participant != 731,]

params.peaktime.data <- merge(merge(participant.params, participant.peaktimes),
                              delays.conservative)
lm.fit.p2.peaktime <- lm(peaktime/24 ~ p2, data=params.peaktime.data)
p2.peaktime.pval <- summary(lm.fit.p2.peaktime)$coefficients[2,4]
p2.peaktime.pval

params.symp.data <- merge(merge(participant.params, atmarsymptoms),
                          delays.conservative)
lm.fit.p2.symponset <- lm(symp.onset ~ p2, data=params.symp.data)
p2.symponset.pval <- summary(lm.fit.p2.symponset)$coefficients[2,4]
p2.symponset.pval
```

The correlation between peak shedding and viral growth rate is verified
by running a linear regression between the p2 and p3 parameters from the
statistical model.

```{r}

lm.fit.p2.p3 <- 
  lm(p3 ~ p2, data=participant.params)
p2.p3.pval <- 
  summary(lm.fit.p2.p3)$coefficients[2,4]
p2.p3.pval

```

Turning to the second part of our study, i.e. using a deterministic model to
replicate the findings above, here are our packages for ODE model and plotting:

```{r}
require(deSolve)
require(ggplot2)
require(gridExtra)
require(grid)
```

Here, we have our ODE model, based on King *et al*., 2009,  as well as a 
modified model implementing a pathogen carrying capacity.
```{r}
transmission.model <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}

transmission.model.cap <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P*(1 - P/(5*10^5)) - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}
```

We define below some useful functions for defining parameters and creating 
dataframes containing the simulated data from our model. We focus on varying 
pathogen replication rate and immune cell kill rate of the pathogen in the
present study, so we maintain constant values for the remaining parameters
based on values used in King *et al*., 2009. We truncate the 
dataframes when pathogen load reaches 0, assuming pathogen clearance. 

```{r}
get.params <- function(k, r) {
  return(c(y=10^-4, k=k, r=r, a=1, d=.5))
}

desolve_function <- function(tstep, k, r, model, id) {
  time <- seq(0,2, by=tstep)
  params <- get.params(k, r)

  df <- data.frame(lsoda(c(P=id, X=0), time, model, 
                       params))
  
  if(any(df$P <= 0)){
    clearance <- min(which(df$P <= 0))
    df[clearance:length(time),] <- NA
  }
  
  df <- data.frame(time=time, 
                  pathogen=df$P, 
                  immune=df$X, 
                  k=k, r=r, dose=id)
  
  return(df)
}

```

Here we provide functions for calculating time of peak pathogen load and 
symptom onset as well as the delay between them. These values are needed to 
determine whether pre-symptomatic transmission has occurred. If time of peak 
pathogen load comes before time of symptom onset, meaning the calculated delay
is negative, then we have pre-symptomatic transmission. We also include 
a function for creating dataframes that house different values of a chosen
parameter and the corresponding delays between symptom onset and peak pathogen
load, allowing us to assess the correlation between that parameter and 
pre-symptomatic transmission.

```{r}
peak.pathogenloadtime <- function(df) {
  peak.pl <- max(df$pathogen, na.rm=TRUE)
  return(df[which(df$pathogen==peak.pl),]$time)
}

symptom.onset <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(min(symptomatic))
  } else {
    return(NA)
  }
}

symptom.end <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(max(symptomatic))
  } else {
    return(NA)
  }
}

delay.calculation <- function(df, symp.thresh) {
  return(peak.pathogenloadtime(df) - symptom.onset(df, 'immune', symp.thresh))
}

```

The first model we examine is one in which symptom onset and duration is 
dependent only on pathogen load. We are defining pre-symptomatic 
transmission as reaching peak pathogen load before symptom onset. This is a 
conservative method of estimating pre-symptomatic transmission. 

Given this definition, there are two potential outcomes: 1) symptoms 
threshold is below peak pathogen load, signifying post-symptomatic transmission, 
and 2) symptoms threshold is above peak pathogen load, signifying 
asymptomatic infection. The outlier case (symptoms threshold is at peak pathogen 
load) is ignored. 

The second model we examine is one in which symptom onset and duration is 
dependent only on immune response. When immune response rises above a certain 
threshold, the patient is symptomatic.

We produce simulations of these two models to demonstrate their dynamics.

```{r, fig.height = 7.5, fig.width = 6.5}

data.lowreplication.nocap <- desolve_function(tstep=10^-4, 
                            k=3.5, r=30, model=transmission.model, id=10^4)
data.highreplication.nocap <- desolve_function(tstep=10^-4, 
                             k=3.5, r=200, model=transmission.model, id=10^4) 

data.lowreplication.cap <- desolve_function(tstep=10^-4, 
                            k=3.5, r=30, model=transmission.model.cap, id=10^4)
data.highreplication.cap <- desolve_function(tstep=10^-4, 
                             k=3.5, r=200, model=transmission.model.cap, id=10^4) 

path.symp.thresh <- 10^6
immune.sympthresh <- 10
immune.pathcap.sympthresh <- 2
```

Now, we vary pathogen replication rate (r) and plot the delay between
symptom onset and peak pathogen load against various r values for both 
transmission models (with and without pathogen load cap) to
investigate the correlation between replication rate and
pre-symptomatic transmission. 

```{r, fig.height = 3.5, fig.width = 3.5}
get.delays.df <- function(data, param.vals, symp.thresh, var_name) {
  delays.df <- data.frame(placeholder_name=integer(), 
                          delay.symponset.peakload=integer())
  
  names(delays.df)[names(delays.df) == "placeholder_name"] <- var_name
  
  for (v in param.vals) {
    df <- data[data[var_name]==v,]
    delay.symponset.peakload <- delay.calculation(df, symp.thresh)
    dftoadd <- data.frame(v, delay.symponset.peakload)
    names(dftoadd)= c(var_name, "delay.symponset.peakload")
    delays.df <- rbind(delays.df, dftoadd)
  }
  
  delays.df$symp.thresh = symp.thresh
  
  return(delays.df)
}

r_vals <- seq(10, 250, by=10)

varyingr.nocap <- data.frame(k= 3.5, r=r_vals, dose=10^5)
varyingr.nocap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  k=varyingr.nocap$k, 
                                  r=varyingr.nocap$r, 
                                  id=varyingr.nocap$dose,                                  
                                  MoreArgs=list(model=transmission.model),
                                  SIMPLIFY = FALSE))

varyingr.cap <- data.frame(k= 7, r=r_vals, dose=10^5)
varyingr.cap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  k=varyingr.cap$k, 
                                  r=varyingr.cap$r, 
                                  id=varyingr.cap$dose,                                  
                                  MoreArgs=list(model=transmission.model.cap),
                                  SIMPLIFY = FALSE))

delays.byr.lowsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=5, 
                                          var_name="r")

delays.byr.medsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=10, 
                                          var_name="r")

delays.byr.highsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=15, 
                                          var_name="r")

delays.byr.lowsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=1, 
                                          var_name="r")

delays.byr.medsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=2, 
                                          var_name="r")

delays.byr.highsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=3, 
                                          var_name="r")
```

We show that for all replication rates used, immune response (X) peaks after 
pathogen load (P).

```{r, fig.height = 3.5, fig.width = 6.5}
peak.immunetime <- function(df) {
  peak.immune <- max(df$immune, na.rm=TRUE)
  return(df[which(df$immune==peak.immune),]$time)
}

get.peaktimes.df <- function(data) {
  peaktimes.df <- data.frame()
  
  for (r in r_vals) {
    df <- data[data$r==r,]
    pathogen.peaktime <- peak.pathogenloadtime(df)
    immune.peaktime <- peak.immunetime(df)
    dftoadd <- data.frame(r, pathogen.peaktime, immune.peaktime)
    peaktimes.df <- rbind(peaktimes.df, dftoadd)
  }
  
  return(peaktimes.df)
}

peaktimes.cap <- get.peaktimes.df(varyingr.cap)
peaktimes.nocap <- get.peaktimes.df(varyingr.nocap)
```
